### 2020-06-10 Output/manuscript/HI_Model_ms.Rmd
commit files

---
title: "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
bibliography: HI_Model.bib
csl: environmental-research-letters.csl
fontsize: 12pt
geometry: margin=1in
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
latex_engine: pdflatex  
output: pdf_document
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```
Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
Parker C. Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^

**Affiliations:**  
^1^U.S. Geological Survey, Moffett Field, CA, USA   
^2^U.S. Geological Survey, Seattle, WA, USA  
^3^University of Hawaii at Manoa, Honolulu, HI, USA  
^4^The East-West Center, Honolulu, HI, USA  
^5^Arizona State University, Tempe, AZ, USA  
^6^Author to whom correspondence should be addressed

**Email:** pselmants@usgs.gov

**Running title:** Hawaii carbon balance  
**Keywords:** land use, climate change, carbon balance, Hawaii, scenarios, disturbance, ecosystem model

**Date:** `r format(Sys.time(), '%B %d, %Y')`

\pagebreak

## Abstract

The State of Hawaii recently passed legislation setting a goal to be carbon neutral by 2045. Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under four unique scenarios over a 90-year timespan. Net ecosystem carbon balance declined under all four scenarios. Moving from a high to a low radiative forcing scenario reduced net ecosystem carbon loss by ~21%, and net carbon losses were reduced by a total of ~55% under the combined scenario of low radiative forcing and low rates of land‐use change. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to better constrainment of models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction

## Methods

### *Study area*

The study area encompasses the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^ (Figure 1).

```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover map of the seven main Hawaiian Islands, adapted from Jacobi et al. (2017). Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate State Classes in the simulation model."}
knitr::include_graphics("./fig_images/fig1_HI_LC.png")
```

### *States and transitions*

### *Carbon stocks and flows*

### *Initial conditions*

### *Scenario simulations*

## Results

## Discussion

## Conclusion

## Acknowledgements

### 2020-06-10 CarbonStockFlow_IBIS/code/HI_Events_FlowPathways.R

@@ -1,0 +1,171 @@
+## Disturbance Event Flow Pathways
+## for HI LUCAS Model 
+## Soil flux from Don et al. (2010)
+## Other fluxes from Sleeter et al. (2017)
+
+## Paul C. Selmants 
+## 2019-09-10 (ISO 8601) 
+
+## UPDATED 2020-02-06
+
+#load required packages into R
+library(dplyr)
+
+#Flow pathways and multipliers for Agricultural Expansion
+AgEx <- data.frame("FromStratumID" = "",
+	"FromStockTypeID" = c(rep(c("Deadwood", "Litter", "Living Biomass", 
+		"Soil"), 3), "Living Biomass"),
+	"ToStockTypeID" = c(rep("Atmosphere",12), "Harvest Products"),
+	"TransitionGroupID" = c(rep(c("AGRICULTURAL EXPANSION: Forest->Ag [Type]",
+		"AGRICULTURAL EXPANSION: Grassland->Ag [Type]", 
+		"AGRICULTURAL EXPANSION: Shrubland->Ag [Type]"), each = 4), 
+		"AGRICULTURAL EXPANSION: Forest->Ag [Type]"), 
+	"FlowTypeID" = c(rep(c("Emission (deadwood)", "Emission (litter)", 
+		"Emission (biomass)", "Emission (soil)"), 3), "Harvest (Timber)"), 
+	"Multiplier" = c(rep(1,2), 0.45, 0.13, rep(1, 3), 0.10, rep(1, 3),
+		0.10, 0.55),
+	stringsAsFactors = FALSE)
+
+#Flow pathways and multipliers for Urbanization
+Urban <- data.frame("FromStratumID" = "",
+	"FromStockTypeID" = c(rep(c("Deadwood", "Litter", "Living Biomass",
+	 "Soil"), 4), "Living Biomass"),
+	"ToStockTypeID" = c(rep("Atmosphere",16), "Harvest Products"),
+	"TransitionGroupID" = c(rep(c("URBANIZATION: Agriculture->Developed [Type]", 
+		"URBANIZATION: Forest->Developed [Type]",
+		"URBANIZATION: Grassland->Developed [Type]", 
+		"URBANIZATION: Shrubland->Developed [Type]"), each = 4), 
+		"URBANIZATION: Forest->Developed [Type]"), 
[…]

### 2020-06-10 CarbonStockFlow_IBIS/code/HI_FlowPathways.R

@@ -1,0 +1,324 @@
+## Flow pathway multipliers for HI LUCAS model
+## forest (covmax1), ag (covmax2), shrub (covmax3), 
+## grass (covmax4), and woodycrop (covmax5)
+## state classes within each moisture zone (MZ)
+## Paul C. Selmants
+## 2019-02-04 (ISO 8601)
+
+# Set working directory to "CarbonStockFlow_IBIS" folder of HI_Model
+# GitHub repository (https://github.com/selmants/HI_Model)
+
+## This R script is dependent on output from Run HI_SAV.R and HI_Flux.R scripts ##
+
+#load required packages into R
+library(dplyr)
+library(tidyr)
+
+#read State Attribute Value (SAV) file into R,
+#filter out values < 19 years, calculate lag 
+#(previous year) values for each carbon stock, filter out year 19, and 
+#filter out Dry WoodyCrop values (not accurately represented by IBIS). 
+SAV <- read.csv('./data/processed/SAV_Full.csv', stringsAsFactors = FALSE) %>%
+	filter(AgeMin >= 19) %>%
+	group_by(StratumID, StateClassID, StateAttributeTypeID) %>%
+	mutate(lagValue = lag(Value)) %>%
+	filter(AgeMin != 19) %>%
+	filter(StateClassID != 'WoodyCrop:All' | StratumID != 'Dry') %>%
+	filter(StateClassID != 'Plantation:All') %>%
+	as.data.frame(.)
+# Read IBIS flux values .csv file into R
+flux <- read.csv('./data/processed/HI_Flux.csv', stringsAsFactors = FALSE)
+
+				#### All COVER CLASSES ####
+#make separate R objects for State Attribute Values used as input 
+#to Flow Multiplier equations. 
+biomass <- filter(SAV, StateAttributeTypeID == 'Initial Conditions: Living Biomass')		
+litter <- filter(SAV, StateAttributeTypeID == 'Initial Conditions: Litter')	
+soc <- filter(SAV, StateAttributeTypeID == 'Initial Conditions: Soil')
+#make separate R objects for fluxes used as variables in 
+#Flow Multiplier calculations for ALL cover classes
+NPP <- filter(SAV, StateAttributeTypeID == 'NPP')
+rawlitc <- filter(flux, Flux == 'rawlitc')
[…]

### 2020-06-10 CarbonStockFlow_IBIS/code/HI_Flux.R

@@ -1,0 +1,199 @@
+# Format Hawaii IBIS summary files for input as Flow Pathways in
+# LUCAS ST-SIM including only years 20-110 (stable) for each Flow
+# Paul C. Selmants
+# 2017-01-05 (ISO 8601)
+# Updated to include Agriculture (covmax2) and WoodyCrop (covmax5)
+
+# Set working directory to "CarbonStockFlow_IBIS" folder in HI_Model
+# GitHub repository (https://github.com/selmants/HI_Model)
+
+## load required packages into R
+library(dplyr)
+library(tidyr)
+
+## make list of IBIS fallw (mortality, kgC/m2/y) summary output text files  
+fallw_list <- list.files(pattern = 'fallw', recursive = TRUE,
+	full.names = TRUE)
+## read IBIS fallw text files into R and format for ST-SIM
+fallw <- lapply(fallw_list[c(1,3,5)], FUN = read.table, header = TRUE) %>%
+	bind_rows() %>%
+	slice(c(7:9, 23:25, 39:41)) %>%
+	mutate(FromStateClassID = rep(c('Forest:All', 'Shrubland:All', 'WoodyCrop:All'),
+		each = 3)) %>%
+	select(FromStateClassID, region_id, X20:X110) %>%
+	gather(year, Value, X20:X110) %>%
+	arrange(FromStateClassID, region_id) %>%
+	mutate(Age = rep(20:110, 9),
+		Flux = 'fallw',
+		FromStratumID = rep(c('Dry', 'Mesic', 'Wet'), 3,  each = 91)) %>%
+	select(FromStratumID, FromStateClassID, Age, Flux, Value)
+
+## make list of IBIS rawlitc (litterfall, kgC/m2/y) summary output text files
+rawlitc_list <- list.files(pattern = 'rawlitc', recursive = TRUE, 
+	full.names = TRUE)	
+## read IBIS rawlitc text files into R and format for ST-SIM
+rawlitc <- lapply(rawlitc_list[c(1:5)], FUN = read.table, header = TRUE) %>%
+	bind_rows() %>%
+	slice(c(7:9, 23:25, 39:41, 55:57, 71:73)) %>%
+	mutate(FromStateClassID = rep(c('Forest:All', 'Agriculture:All', 'Shrubland:All',
+		'Grassland:All', 'WoodyCrop:All'),
[…]

### 2020-06-10 CarbonStockFlow_IBIS/code/HI_SAV.R

@@ -1,0 +1,208 @@
+## Format Hawaii IBIS summary files for input as 
+## State Attribute Values in LUCAS ST-SIM
+## Paul C. Selmants
+## 2020-04-02 (ISO 8601)
+## Summarized NPP for all State Classes and 
+## SAV for non-tree State Classes
+
+# Set working directory to "CarbonStockFlow_IBIS" folder in HI_Model
+# GitHub repository (https://github.com/selmants/HI_Model)
+
+## load required packages into R
+library(dplyr)
+library(tidyr)
+
+## make a list of IBIS cbiotot (live biomass in kgC/m2) summary output
+## ASCII text files for all covmax classes 
+biotot_list <- list.files(pattern = 'cbiotot', recursive = TRUE,
+	full.names = TRUE)
+## read IBIS cbiotot (live biomass in kgC/m2) summary output ASCII 
+## text files for all covmax classes into R and format for ST-SIM 
+biomass <- lapply(biotot_list[c(1:5)], FUN = read.table, header = TRUE) %>%
+	bind_rows() %>%
+	slice(c(7:9, 23:25, 39:41, 55:57, 71:73)) %>%
+	mutate(StateClassID = rep(c('Forest:All', 'Agriculture:All','Shrubland:All',
+	 'Grassland:All', 'WoodyCrop:All'), each = 3)) %>%
+	select(StateClassID, region_id, X1:X110) %>%
+	gather(year, Value, X1:X110) %>%
+	arrange(StateClassID, region_id) %>%
+	mutate(AgeMin = rep(1:110, 15),
+		AgeMax = as.integer(rep(c(1:109,""), 15)),
+		StratumID = rep(c('Dry', 'Mesic', 'Wet'), 5,  each = 110),
+		StateAttributeTypeID = 'Initial Conditions: Living Biomass') %>%
+	select(StratumID, StateClassID, StateAttributeTypeID, AgeMin, AgeMax, Value)
+
+## make a list of IBIS stdwdc (standing deadwood in kgC/m2) summary output 
+## ASCII text files for covmax 1(forest), 3(shrubland), and 5(woodycrop)
+stdw_list <- list.files(pattern = 'stdwdc', recursive = TRUE,
+	full.names = TRUE)
+## read IBIS stdwdc (standing deadwood C in kgC/m2) summary output ASCII 
[…]

### 2020-06-10 Fire/code/AnnualFire_plot.R

@@ -1,0 +1,41 @@
+## Graph of Annual Area Burned, 
+## State of Hawaii, 1999-2019
+## Paul C. Selmants
+## R version 3.6.1
+## 2019-10-26
+
+# Set working directory to "Fire" folder in HI_Model
+# GitHub repository (https://github.com/selmants/HI_Model)
+
+## This R script is dependent on output from Fire_probabilities.R script ## 
+
+# load ggplot2 version 3.2.0
+library(ggplot2)
+# load dplyr version 0.8.3
+library(dplyr)
+
+# read statewide annual burned area data into R
+data <- read.csv('AnnualFire.csv', stringsAsFactors = FALSE) 
+
+# sum burned area statewide by year
+statewide <- data %>%
+	group_by(year) %>%
+	summarize(burnarea_km2 = sum(burnarea_km2)) %>%
+	as.data.frame()
+
+# make line graph of annual area burned
+p <- ggplot(statewide, aes(year, burnarea_km2)) +
+	geom_line(color = "#D95F02", size = 0.5) +
+	geom_point(color = "#D95F02", size = 2, shape = 25, fill = "#D95F02") +
+	xlab("Year") +
+	ylab(expression(Area~burned~(km^{2}))) +
+	scale_x_continuous(limits = c(1999, 2019), breaks = seq(2000, 2018, 2)) +
+	scale_y_continuous(limits = c(0, 175), breaks = seq(0, 175, 25)) +
+	geom_hline(aes(yintercept=28.75), linetype="dashed", size = 0.3) +
+	theme(panel.grid.minor=element_blank()) +
+	theme(axis.title.x = element_text(size = 16),
+          axis.text.x  = element_text(size = 12)) +
+	theme(axis.title.y = element_text(size = 16),
+          axis.text.y  = element_text(size = 12))
+
+ggsave("./data/processed/HI_AnnualAreaBurned.png", width = 7, height = 3.5)

### 2020-06-10 Fire/code/FireSizeDistribution.R

@@ -1,0 +1,45 @@
+## Fire size distribution
+## Paul C. Selmants
+## 2019-10-18
+
+## Set working directory to "Fire" folder in HI_Model
+## GitHub repository (https://github.com/selmants/HI_Model)
+
+# load required packages into R
+library(raster)
+library(dplyr)
+
+# read in Fire Perimeters shapefile
+fires <- shapefile("./data/base/Hawaii Fire Perimeters 1999-2018.shp")
+
+# calculate area of each polygon in km2
+fires$area_km2 <- area(fires)/1000000
+
+# make dataframe of year and fire size in km2
+size <- data.frame("year" = fires$Year,
+			"area_km2" = fires$area_km2) %>%
+		arrange(year) 
+
+# create function to calculate proportion of fires within a size range
+propsize <- function(x,y){
+	(sum(size$area_km2 > x & size$area_km2 <= y))/269
+}
+
+# calculate proportion of fires within 8 size ranges, from 0-105km2
+p1k <- propsize(0,1)
+p2k <- propsize(1,2)
+p5k <- propsize(2,5)
+p10k <- propsize(5,10)
+p25k <- propsize(10,25)
+p50k <- propsize(25,50)
+p100k <- propsize(50,100)
+p200k <- propsize(100,200)
+
+# create Transition Size Distribution dataframe
+sizedist <- data.frame(
+	"TransitionGroupID" = "FIRE",
+	"MaximumArea" = c(1,2,5,10,25,50,100,200),
+	"RelativeAmount" = c(p1k,p2k,p5k,p10k,p25k,p50k,p100k,p200k))
+
+# write sizedist dataframe to .csv file
+write.csv(sizedist, "./data/processed/FireSizeDistribution.csv", row.names = FALSE)

### 2020-06-10 Fire/code/Fire_probabilities.R

@@ -1,0 +1,311 @@
+## Rasterize shapefile of 1999-2019 
+## annual fire perimeters in Hawaii 
+## and calculate annual area burned 
+## by Island, Moisture Zone, 
+## and State Class 
+## Paul C. Selmants
+## 2020-03-17
+
+# Set working directory to "Fire" folder of HI_Model
+# GitHub repository (https://github.com/selmants/HI_Model)
+
+# load required packages into R
+library(raster)
+library(rgdal)
+library(sf)
+library(fasterize)
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+			#### Read spatial data into R ####
+
+#set working directory to "HI_Model/Fire" in repo. 
+
+# read Island ID GeoTIFF into R from Build_STsim directory, where:
+# 1 = Hawaii		5 = Molokai
+# 2 = Kahoolawe		6 = Oahu
+# 3 = Lanai			7 = Kauai
+# 4 = Maui
+Islands <- raster("../Build_STsim/data/spatial_data/Island_250m.tif")
+# read Moisture Zone GeoTIFF into R, where:
+# 1 = "Dry"
+# 2 = "Mesic"
+# 3 = "Wet"
+MZ <- raster("../Build_STsim/data/spatial_data/MZ_3_250m.tif")
+# read State Class GeoTIFF into R, where: 
+# 1 = Water 		6 = Forest
+# 2 = Urban 		7 = Grassland
+# 3 = Plantation	8 = Agriculture
+# 4 = WoodyCrop		9 = Wetland
+# 5 = Barren	   10 = Shrubland
+StateClass <- raster("../Build_STsim/data/spatial_data/StateClassNew_250m.tif")
+# read in Hawaii statewide annual fire perimeter shapefile (1999-2019) 
+fires <- st_read("./data/base/2019_1999_Hawaii_Fire_Perimeters.shp") %>%
+	dplyr::select(UH_ID, Year, geometry)
+
+			### Reclassify and customize raster data ### 
+
+# reclassification matrix to make all islands values = 0
+islands_recl <- c(0, NA, 1, 0, 2, 0, 3, 0, 4, 0, 5, 0, 6, 0, 
+	7, 0, 255, NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+# reclassify islands raster to all zero values
+zero_islands <- reclassify(Islands, islands_recl)
[…]

### 2020-06-10 FlowMultipliers/code/CFEmultipliers.R

@@ -1,0 +1,110 @@
+## Create dataframes of annual NPP multipliers
+## for different rates of CO2 fertilization (CFE)
+## based on projections of annual atmospheric CO2 
+## concentrations under RCP 4.5 and RCP 8.5
+## Paul C. Selmants
+## R version 3.6.1
+## 2020-04-10
+
+# load dplyr v. 0.8.3 into R
+library(dplyr)
+
+# read in climate-based NPP multipliers for RCP 4.5 and RCP 8.5
+nppmulti45 <- read.csv("RCP45_NPPmultipliers.csv", stringsAsFactors = FALSE)
+nppmulti85 <- read.csv("RCP85_NPPmultipliers.csv", stringsAsFactors = FALSE)
+
+# read in RCP 4.5 mid-year atm CO2 concentrations
+rcp45 <- read.table(
+	"http://www.pik-potsdam.de/~mmalte/rcps/data/RCP45_MIDYEAR_CONCENTRATIONS.DAT",
+	header = FALSE,
+	stringsAsFactors = FALSE, 
+	skip = 39) %>% 
+select(year = V1, rcp45CO2ppm = V4) %>%
+filter(year >2009, year <2101) %>%
+mutate(incr45 = rcp45CO2ppm - .[1,2], 
+		cfe025 = 1+(incr45*0.025)/100, 
+		cfe05 = 1+(incr45*0.05)/100,
+		cfe075 = 1+(incr45*0.075)/100,
+		cfe10 = 1+(incr45*0.10)/100,
+		cfe125 = 1+(incr45*0.125)/100,
+		cfe15 = 1+(incr45*0.15)/100) %>%
+filter(year > 2010) %>%
+select(year, rcp45CO2ppm, cfe025:cfe15)
+# read in RCP 8.5 mid-year atm CO2 concentrations
+rcp85 <- read.table(
+	"http://www.pik-potsdam.de/~mmalte/rcps/data/RCP85_MIDYEAR_CONCENTRATIONS.DAT",
+	header = FALSE,
+	stringsAsFactors = FALSE, 
+	skip = 39) %>% 
+select(year = V1, rcp85CO2ppm = V4) %>%
+filter(year >2009, year <2101) %>%
+mutate(incr85 = rcp85CO2ppm - .[1,2],
+		cfe025 = 1+(incr85*0.025)/100, 
[…]

### 2020-06-10 FlowMultipliers/code/MZSC.R

@@ -1,0 +1,73 @@
+## Create raster of
+## merged Moisture Zones 
+## and State Classes 
+## Paul C. Selmants
+## 2019-10-25
+
+# read required packages into R
+library(raster)
+library(rgdal)
+library(dplyr)
+
+## set working directory at highest level of repo: "HI_Model" ##
+
+# read Island ID TIFF into R, where:
+# 1 = Hawaii		5 = Molokai
+# 2 = Kahoolawe		6 = Oahu
+# 3 = Lanai			7 = Kauai
+# 4 = Maui
+Islands <- raster("./Build_STsim/data/spatial_data/Island_250m.tif")
+# read Moisture Zone TIFF into R, where:
+# 1 = "Dry", 2 = "Mesic", 3 = "Wet"
+MZ <- raster('./Build_STsim/data/spatial_data/MZ_3_250m.tif')
+# read State Class TIFF into R, where: 
+# 1 = Water 		6 = Forest
+# 2 = Urban 		7 = Grassland
+# 3 = Plantation	8 = Agriculture
+# 4 = WoodyCrop		9 = Wetland
+# 5 = Barren	   10 = Shrubland
+StateClass <- raster('./Build_STsim/data/spatial_data/StateClassNew_250m.tif')
+
+# make reclassification matrix to make all islands values = 0
+islands_recl <- c(0, NA, 1, 0, 2, 0, 3, 0, 4, 0, 5, 0, 6, 0, 
+	7, 0, 255, NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+# reclassify islands raster
+zero_islands <- reclassify(Islands, islands_recl)
+
+# make reclassification matrices to subset moisture zones into 
+# individual rasters and set both '0' and '255' to 'NA'
+dry_recl <- c(0, NA, 1, 1, 2, NA, 3, NA, 255, NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+mesic_recl <- c(0, NA, 1, NA, 2, 1, 3, NA, 255, NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+wet_recl <- c(0, NA, 1, NA, 2, NA, 3, 1, 255, NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+#reclassify MZ into individual zones 
[…]

### 2020-06-10 FlowMultipliers/code/NPPmultipliers.R

@@ -1,0 +1,578 @@
+## Calculate NPP spatial multiplier raster
+## and NPP climate flow multiplier datasheets
+## using Del Grosso NCEAS NPP equations and 
+## statistically downscaled CMIP5
+## projections of rainfall and temperature
+## for RCP 4.5 and RCP 8.5. 
+## Paul C. Selmants
+## 2020-03-19 (ISO 8601)
+
+# load required packages into R
+library(raster)
+library(rgdal)
+library(tidyr, warn.conflicts = FALSE)
+library(dplyr, warn.conflicts = FALSE)
+
+# read in State Class raster, where: 
+# 1 = Water 		 6 = Forest		
+# 2 = Urban 		 7 = Grassland		
+# 3 = Plantation 	 8 = Agriculture			 
+# 4 = WoodyCrop 	 9 = Wetland		 
+# 5 = Barren 		10 = Shrubland		
+SC <- raster("StateClassNew_250m.tif")
+
+# reclassification matrix for SC to set Water, Urban, 
+# Barren, & Wetland to 1 and set other classes to NA
+sc_rcl <- c(0,NA,1,1,2,1,3,NA,4,NA,5,1,6,NA,7,NA,8,NA,9,1,10,NA,255,NA) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+# reclassify SC raster 
+SCnoveg <- reclassify(SC, sc_rcl)
+
+# read in raster of combined Moisture Zones and State Classes, where:
+# 1 = Dry Forest	 	 7 = Mesic Forest			13 = Wet Forest
+# 2 = Dry Grass		 	 8 = Mesic Grass			14 = Wet Grass
+# 3 = Dry Shrub		 	 9 = Mesic Shrub			15 = Wet Shrub 
+# 4 = Dry Plantation	10 = Mesic Plantation		16 = Wet Plantation
+# 5 = Dry Ag			11 = Mesic Ag				17 = Wet Ag
+# 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
+MZSC <- raster("MZSC.tif")
+
+# reclassification matrix to create forest MZSC raster
+for_rcl <- c(1,1,2,NA,3,NA,4,4,5,NA,6,6,7,7,8,NA,9,NA,10,10,
+	11,NA,12,12,13,13,14,NA,15,NA,16,16,17,NA,18,18) %>%
[…]

### 2020-06-10 FlowMultipliers/code/Q10DecayMultipliers.R

@@ -1,0 +1,512 @@
+## Calculate future flow multipliers for decay and 
+## CO2 efflux for litter, down dead wood, and soil
+## based on statistically downscaled
+## CMIP5 projections of future temperature
+## for RCP 4.5 and RCP 8.5. 
+## Paul C. Selmants
+## 2020-03-23 (ISO 8601)
+
+# load required packages into R
+library(raster)
+library(rgdal)
+library(tidyr, warn.conflicts = FALSE)
+library(dplyr, warn.conflicts = FALSE)
+
+# read in raster of combined Moisture Zones and State Classes, where:
+# 1 = Dry Forest	 	 7 = Mesic Forest			13 = Wet Forest
+# 2 = Dry Grass		 	 8 = Mesic Grass			14 = Wet Grass
+# 3 = Dry Shrub		 	 9 = Mesic Shrub			15 = Wet Shrub 
+# 4 = Dry Plantation	10 = Mesic Plantation		16 = Wet Plantation
+# 5 = Dry Ag			11 = Mesic Ag				17 = Wet Ag
+# 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
+MZSC <- raster("MZSC.tif")
+
+# create list of individual island .tif files
+islandlist <- list.files('islands', full.names = TRUE)
+#create raster stack of individual island .tif files
+islandstack <- stack(islandlist)
+
+# mask MZ-specific SC raster with raster stack of islands
+mzsc_island <- mask(MZSC, islandstack) 
+names(mzsc_island) <- names(islandstack)
+
+# create list of temperature delta .tif files
+tdeltalist <- list.files(pattern = 'sd_Tdeltas', full.names = TRUE)
+# stack temperature delta rasters and match projection to MZSC
+tdeltastack <- stack(tdeltalist) %>%
+	projectRaster(., MZSC, method = 'bilinear')
+# create a vector of rcp names
+rcpnames <- c('rcp45late', 'rcp45mid', 'rcp85late', 'rcp85mid')
+# rename layers in Tdeltastack
+names(tdeltastack) <- rcpnames
+
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_CFE_CarbonBalance.R

@@ -1,0 +1,148 @@
+# CFE carbon balance 
+# Hawaii LUCAS Model 
+# Paul C. Selmants
+# 2020-04-14
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+
+# read LUCAS CFE C stocks from RCP85High output data into R
+CFE_Cstocks <- read.csv("Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in LUCAS C stocks by scenario and filter on RCP85High scenario 
+Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_rate = 0) %>%
+	select(CFE_rate, Timestep:Amount)
+
+# read in CFE base flow data from RCP85High Scenario
+cfe_baseflow <- read.csv("BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in CFE C loss data (transition, fire, harvest)
+cfe_closs <- read.csv("TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in zero CFE base flow data 
+baseflowzero <- read.csv("BaseFlows.csv", stringsAsFactors = FALSE) %>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_Rate = 0) %>%
+	select(CFE_Rate, Timestep:Amount)
+# read zero CFE non-Rh C Loss data 
+closszero <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_Rate = 0) %>%
+	select(CFE_Rate, Timestep:Amount)
+
+## Summarize C stock data and create TEC by CFE figure ## 
+
+# combine CFE Cstock data together and summarize as Statewide TEC 
+cfetec <- bind_rows(Cstocks, CFE_Cstocks) %>%
+	filter(CFE_rate %in% c(0, 5, 10, 15)) %>%
+	group_by(CFE_rate, Timestep, SecondaryStratumID, StockGroupID) %>%
+	summarize(CstockTg = mean(Amount)/1000, 
+		Cstock_min = min(Amount)/1000, 
+		Cstock_max = max(Amount)/1000) %>%
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_CarbonBalance.R

@@ -1,0 +1,233 @@
+## Processing HI LUCAS model output:
+## summarizing Flows to estimate annual
+## and cumulative net biome productivity (NBP)
+## Paul C. Selmants
+## 2020-05-20
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(cowplot)
+
+
+# read HI LUCAS baseline C flow data into R
+baseflow<- read.csv("BaseFlows.csv", stringsAsFactors = FALSE)
+# read HI LUCAS Transition/Harvest C Loss data into R
+transflow <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE)
+
+# calculate statewide NPP
+NPP <- baseflow %>%
+	filter(FlowGroupID == "Growth [Type]") %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(NPP = sum(Amount)) %>%
+	as.data.frame()	
+# calculate statewide baseline Rh 
+Rh <- baseflow %>%
+	filter(FlowGroupID %in% c("Emission (litter) [Type]",
+		"Emission (soil) [Type]")) %>% 
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(Rh = sum(Amount)) %>%
+	as.data.frame()
+# calculate statewide C loss by leaching/erosion
+Leach <- baseflow %>%
+	filter(FlowGroupID == "Leaching [Type]") %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(Leach = sum(Amount)) %>%
+	as.data.frame()
+# calculate disturbance C losses (∆LULC, fire, ag harvest)
+LUCloss <- transflow %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(LUCloss = sum(Amount)) %>%
+	as.data.frame()
+
+# join NPP and C loss dataframes together
+cflowjoin <- left_join(NPP, Rh) %>%
+	left_join(., Leach) %>%
+	left_join(., LUCloss) 
+
+# convert c flow data to long format, separate Scenario column into 
+# RCP and LandUse, summarize flow data by mean, min, and max in Tg 
+cflows <- cflowjoin %>% 
+	pivot_longer(cols = NPP:LUCloss, 
+		names_to = "flowtype",
+		values_to = "Amount") %>%
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_CarbonStocks.R

@@ -1,0 +1,113 @@
+## HI LUCAS model output:
+## summarize & graph Statewide 
+## Carbon stocks, fluxes, and balance
+## Paul C. Selmants
+## 2020-05-20
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+
+# read Hawaii LUCAS model C stock output data into R
+Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE) %>%
+	group_by(Scenario, Timestep, SecondaryStratumID, StockGroupID) %>%
+	summarize(CstockTg = mean(Amount)/1000,
+		Cstock_min = min(Amount)/1000,
+		Cstock_max = max(Amount)/1000) %>%
+	data.frame()
+# remove " [Type]" at trailing end of all stock types  
+Cstocks$StockGroupID <- gsub('.{7}$', '', Cstocks$StockGroupID) 
+
+# create function to summarize carbon stocks statewide
+stocksum <- function(x) {
+	 x %>%
+	group_by(Scenario, Timestep) %>%
+	summarize(C_Tg = sum(CstockTg), 
+		Cmin = sum(Cstock_min),
+		Cmax = sum(Cstock_max)) %>%
+	select(Scenario, Timestep, C_Tg, Cmin, Cmax) %>%
+	arrange(Scenario, Timestep) %>%
+	as.data.frame() 
+}
+
+# summarize TEC and individual C stocks statewide
+TEC <- stocksum(Cstocks)
+# sum of standing and down dead wood
+Deadwood <- Cstocks %>%
+	filter(grepl("Deadwood", StockGroupID)) %>%
+	stocksum(.)
+# Living biomass
+LB <- Cstocks %>%
+	filter(StockGroupID == "Living Biomass") %>%
+	stocksum(.)
+# Litter
+Litter <- Cstocks %>%
+	filter(StockGroupID == "Litter") %>%
+	stocksum(.)
+# Soil organic carbon
+SOC <- Cstocks %>%
+	filter(StockGroupID == "Soil") %>%
+	stocksum(.)
+
+# merge individual C stocks into single dataframe
+indCstocks <- bind_rows(Deadwood, LB, Litter, SOC) %>%
+	mutate(StockGroupID = rep(c("Dead wood", "Live biomass", "Litter",
+		"Soil organic matter"), each = 364)) %>%
+	arrange(Scenario, Timestep) %>%
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_ClimateFigs.R

@@ -1,0 +1,146 @@
+# Use ggplot to visualize
+# current and end-of-century
+# rainfall and air temperature  
+# Paul C. Selmants
+# 2020-02-24
+
+# load required R packages
+library(raster)
+library(rgdal)
+library(ggplot2)
+library(dplyr)
+library(viridis)
+library(RColorBrewer)
+library(cowplot) 
+
+# create list of annual rainfall .tif files
+rainlist <- list.files('rainfall_tiff_annual', full.names = TRUE)
+#create raster stack of annual rainfall (mm) .tif files
+rainstack <- stack(rainlist)
+# make raster layer of 30-year mean annual rainfall and re-sample
+# to change projection to UTM (this takes ~30 seconds)
+rain30ymean <- calc(rainstack, mean)
+# read in rasters of late century rainfall anomolies for RCPs 4.5 & 8.5
+rcp85anom <- raster("sd_rfanom_rcp85_late.tif")
+rcp45anom <- raster("sd_rfanom_rcp45_late.tif")
+# make rasters of late century rainfall differences from present
+raindiff85 <- rain30ymean*(rcp85anom/100)
+raindiff45 <- rain30ymean*(rcp45anom/100)
+
+# read in current mean annual temperature raster
+MAT <- raster("MAT_annual.tif") 
+# read in rasters of late century temperature differences for RCPs 4.5 & 8.5
+Tdelta_rcp45 <- raster("sd_Tdeltas_rcp45_late.tif")
+Tdelta_rcp85 <- raster("sd_Tdeltas_rcp85_late.tif")
+
+#prepare rain30ymean and raindiff data to plot with geom_tile in ggplot
+rain_df <- rasterToPoints(rain30ymean) %>%
+	data.frame(.)
+
+r45_df <- rasterToPoints(raindiff45) %>%
+	data.frame(.)
+
+r85_df <- rasterToPoints(raindiff85) %>%
+	data.frame(.)
+#prepare MAT and dT data to plot with geom_tile in ggplot2 
+MAT_df <- rasterToPoints(MAT) %>%
+	data.frame(.) %>%
+	select(x, y, layer = MAT_annual)
+
+t45_df <- rasterToPoints(Tdelta_rcp45) %>%
+	data.frame(.) %>%
+	select(x, y, layer = sd_Tdeltas_rcp45_late)
+
+t85_df <- rasterToPoints(Tdelta_rcp85) %>%
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_HistoricLULC.R

@@ -1,0 +1,33 @@
+# Historic rates of land cover change
+# for Hawaii LUCAS model
+# Paul C. Selmants
+# 2020-03-10
+
+# load packages
+library(dplyr)
+library(ggplot2)
+
+# read in historic land cover change data
+ccap <- read.csv("lulc_historic.csv", stringsAsFactors = FALSE) %>%
+	group_by(LC_change) %>%
+	summarize(median = sum(mid), 
+		min = sum(minimum), 
+		max = sum(maximum)) %>%
+	as.data.frame()
+
+urbagex <- ccap %>%
+	filter(LC_change != "Ag contraction")
+# make historic dLULC graph
+urbagexfig <- ggplot(urbagex, aes(LC_change, median)) +
+	geom_point(size = 4, shape = 15) +
+	geom_errorbar(aes(ymin = min, ymax = max), width = 0) +
+	scale_y_continuous(limits = c(0, 10), breaks = seq(0,10,1)) +
+	ylab(expression(Annual~rate~(km^2~y^{-1}))) +
+	theme(panel.grid.minor = element_blank(),
+		axis.title.y = element_blank(), 
+		axis.text.y = element_text(size = 12), 
+		axis.title.x = element_text(size = 14),
+		axis.text.x = element_text(size=12))+
+	coord_flip() 
+
+ggsave("histLULC.png", width = 9, height = 3)	

### 2020-06-10 Output/manuscript/fig_code/HI_LandCoverFig.R

@@ -1,0 +1,80 @@
+# Using ggplot2 to visualize
+# Hawaiʻi land cover classes
+# Paul C. Selmants
+# 2020-02-24
+
+# load required packages
+library(raster)
+library(rgdal)
+library(dplyr)
+library(ggplot2)
+
+# set working directory to HI_Model GitHub repository
+
+# read State Class TIFF into R, where: 
+# 1 = Water 		6 = Forest
+# 2 = Urban 		7 = Grassland
+# 3 = Plantation	8 = Agriculture
+# 4 = WoodyCrop		9 = Wetland
+# 5 = Barren	   10 = Shrubland
+StateClass <- raster('./Build_STsim/data/spatial_data/StateClassNew_250m.tif')
+
+# read annual MAT Tiff into R
+MAT <- raster("./FlowMultipliers/data/base/MAT_annual.tif")
+
+# make reclassification matrix for simplified land cover raster
+lc_rcl <- c(1, NA, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6,
+	7, 7, 8, 4, 9, NA, 10, 10) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+# reclassify StateClass to create simplified land cover raster, where: 
+# 2 = Urban 		 6 = Forest
+# 3 = Plantation	 7 = Grassland
+# 4 = Agriculture	10 = Shrubland
+# 5 = Barren	
+lc <- reclassify(StateClass, lc_rcl)
+# re-sample lc projection to Lat. Long. using rainstack
+lcdeg <- projectRaster(lc, MAT, method = 'ngb')
+# convert lc raster to points dataframe for usage in ggplot 
+lc_df <- rasterToPoints(lcdeg) %>%
+	data.frame(.) %>%
+	select(x, y, layer = StateClassNew_250m) 
+# convert layer column to factor 
+lc_df$layer <- as.factor(lc_df$layer)
+# Create custom colorblind palette for land cover classes
+lcpal <- c("#000000", "#CC79A7", "#A6761D", "#969696", "#006D2C", 
+	"#FFD92F", "#B2DF8A") 
+# create statewide land cover map for Hawaii
+lcmap <- ggplot(data=lc_df) +
+	geom_tile(aes(x=x, y=y, fill=layer)) +
+	xlab(expression("Longitude " ( degree*W))) +
+	ylab(expression("Latitude " ( degree*N))) +
[…]

### 2020-06-10 Output/manuscript/fig_code/HI_StateClassArea.R

@@ -1,0 +1,79 @@
+## Processing HI LUCAS model output:
+## summarizing State Class areas over time
+## by land use scenario
+## Paul C. Selmants
+## 2020-04-22
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(tidyr) 
+
+# read State Class model output data into R 
+sc <- read.csv("StateClassArea.csv", stringsAsFactors = FALSE) %>%
+	mutate(StateClassID = str_remove(.$StateClassID, ":All"))
+
+# summarize State Class Area statewide by Land Use Scenario
+StateClass <- sc %>%
+	group_by(Scenario, Timestep, Iteration, StateClassID) %>%
+	summarize(area = sum(Area)) %>%
+	separate(Scenario, c("Climate", "LandUse"), sep = 5) %>%
+	group_by(LandUse, Timestep, StateClassID) %>%
+	summarize(area_mean = mean(area), 
+		area_min = min(area), 
+		area_max = max(area)) %>%
+	data.frame() 
+
+# ggplot faceted land cover change graph
+scfig <- StateClass %>%
+filter(StateClassID %in% c("Agriculture",
+	"Developed", "Forest", "Grassland", "Shrubland")) %>% 
+	{ggplot(., aes(x = Timestep, y = area_mean, color = LandUse)) +
+	facet_wrap(~ StateClassID, scales = "free_y") +
+	geom_ribbon(aes(ymin = area_min, ymax = area_max, fill = LandUse),
+		alpha = 0.4, color = NA) +
+	geom_line(size = 0.75) +
+	scale_x_continuous(limits = c(2000,2100), breaks = seq(2000,2100,20)) +
+	scale_y_continuous(breaks = scales::pretty_breaks(5), limits = c(NA, NA)) +
+	labs(x = "Year", 
+		 y = expression(Land~area~(km^{2})), 
+		color = "Land Use\nScenario",
+		fill = "Land Use\nScenario") +
+	theme_bw() +
+	theme(panel.grid.minor = element_blank(), 
+		legend.position = c(0.825,0.25),
+		legend.text = element_text(size=10))}
+	
+
+# write function to create ggplot graphs of land cover change
+fig <- function(a,b,c) {
+	ggplot(filter(StateClass, StateClassID == a), 
[…]

### 2020-06-10 Transitions/code/CCAP_LULCC.R

@@ -1,0 +1,206 @@
+# HI CCAP land cover change, 2001-2005
+# Urbanization rate and Ag change rate
+# Paul C. Selmants 
+# 2019-08-19
+
+#load packages into R
+library(dplyr)
+library(tidyr)
+
+#make a list of NOAA CCAP 1992-2001 land cover change files
+ccap01_list <- list.files(pattern = "LCchange_1992")
+#make a list of NOAA CCAP 2001-2005 land cover change files
+ccap05_list <- list.files(pattern = "LCchange_2001")
+#make a list of NOAA CCAP 2001-2005 land cover change files
+ccap10_list <- list.files(pattern = "LCchange_2005") 
+
+#read in ccap 1992-2001 files, bind together, filter out zero values
+LC92_01 <- lapply(ccap01_list[c(1:4)], FUN = read.csv, header = TRUE,
+ 	stringsAsFactors = FALSE) %>%
+	bind_rows() %>%
+	mutate(County = c(rep(c("Hawai'i", "Kaua'i", "Maui"), each = 626),
+		rep("O'ahu", 548))) %>%
+	filter(Count > 0) %>%
+	select(County, Count, Code_1992 = X1992_Code, Class_1992 = X1992_Class, 
+		Code_2001 = X2001_Code, Class_2001 = X2001_Class, Class_Name)
+#read in ccap 2001-2005 files, bind together, filter out zero values
+LC01_05 <- lapply(ccap05_list[c(1:4)], FUN = read.csv, header = TRUE,
+ 	stringsAsFactors = FALSE) %>%
+ 	bind_rows() %>%
+ 	mutate(County = c(rep(c("Hawai'i", "Kaua'i"), each = 626), 
+ 		rep(c("Maui", "O'ahu"), each = 548))) %>%
+ 	filter(Count > 0) %>%
+ 	select(County, Count, Code_2001 = X2001_Code, Class_2001 = X2001_Class, 
+ 		Code_2005 = X2005_Code, Class_2005 = X2005_Class, Class_Name)
+#read in ccap 2001-2005 files, bind together, rename column headers
+LC05_10 <- lapply(ccap10_list, FUN=read.csv, header = TRUE,
+	stringsAsFactors = FALSE) %>%
+	bind_rows() %>%
[…]

### 2020-06-10 Transitions/code/HI_Transition_Pathways_Fire.R

@@ -1,0 +1,147 @@
+## Transition Pathways for Urbanization,
+## Ag Contraction, Ag Expansion, and Fire
+## in Hawaii Model
+## Paul C. Selmants
+## 2018-08-30 
+## UPDATED with new fire data 2019-10-31
+## R version 3.6.1
+
+# load dplyr package v. 0.8.3
+library(dplyr)
+
+# Plantation Harvest transition pathways by island   
+Plant_Harvest <- data.frame("Iteration" = "",
+		"Timestep" = "",
+		"StratumIDSource" = rep(c("Dry", "Mesic", "Wet"), each = 4, 4),
+		"StateClassIDSource" = "Plantation:All",
+		"StratumIDDest" = "", 
+		"StateClassIDDest" = rep(c("Plantation:All", "Agriculture:All",
+			"Grassland:All", "WoodyCrop:All"), each = 12),
+		"SecondaryStratumID" = rep(c("Hawai'i", "Moloka'i", "O'ahu", "Kaua'i"),
+			12),
+		"TertiaryStratumID" = "",
+		"TransitionTypeID" = rep(c("MANAGEMENT: Plantation Harvest", 
+		"MANAGEMENT: Plantation->Ag", "MANAGEMENT: Plantation->Grassland",
+		"MANAGEMENT: Plantation->WoodyCrop"), each = 12), 
+		"Probability" = 1, 
+		"Proportion" = NA,
+		"AgeMin" = 5, 
+		stringsAsFactors = FALSE) 
+
+# Urbanization transition pathways by island 
+Urban <- data.frame("Iteration" = "",
+		"Timestep" = "",
+		"StratumIDSource" = rep(c("Dry", "Mesic", "Wet"), each = 6, 5),  
+		"StateClassIDSource" = rep(c("Agriculture:All", "Barren:All",
+			"Forest:All", "Grassland:All", "Shrubland:All"), each = 18), 
+		"StratumIDDest" = "",
+		"StateClassIDDest" = "Developed:All",
+		"SecondaryStratumID" = rep(c("Hawai'i", "Kaua'i",
+			"Lana'i", "Maui", "Moloka'i", "O'ahu"), 15),
+		"TertiaryStratumID" = "",
+		"TransitionTypeID" = rep(c("URBANIZATION: Agriculture->Developed",
+			"URBANIZATION: Barren->Developed", 
+			"URBANIZATION: Forest->Developed",
[…]

### 2020-06-10 Transitions/code/HI_Transition_Targets.R

@@ -1,0 +1,358 @@
+## Transition Targets for two Urbanization/Ag
+## Change scenarios in Hawaii LUCAS Model 
+## Paul C. Selmants
+## R version 3.6.1
+## 2019-09-09 ISO-8601
+## updated 2019-11-05 ISO-8601
+## updated 2020-04-23 ISO-8601
+
+# load dplyr package (version 0.8.5)
+library(dplyr) 
+
+# read .csv baseline data into R 
+lulc <- read.csv("lulc_historic.csv", stringsAsFactors = FALSE)
+ddev <- read.csv("HI_ddev.csv", stringsAsFactors = FALSE)
+
+# Historical urbanization transition targets by island for 2010-2019, 
+# based on statewide population projections from HI Office of Planning
+UrbanHist <- ddev %>%
+	filter(end_year <= 2020) %>%
+	mutate( 
+		Iteration = "",
+		StratumID = "",
+		TransitionGroupID = "URBANIZATION",
+		DistributionType = "",
+		DistributionFrequencyID = "",
+		DistributionSD = as.numeric(NA),
+		DistributionMin = as.numeric(NA), 
+		DistributionMax = as.numeric(NA)) %>%
+	select(Iteration, Timestep = start_year, StratumID, 
+		SecondaryStratumID = island, TransitionGroupID, 
+		Amount = ddev_km2y, DistributionType, DistributionFrequencyID, 
+		DistributionSD, DistributionMin, DistributionMax)
+# Low urbanization transition targets by island for 2020-2045 
+# based on statewide population projections from HI Office of Planning
+UrbanLow_early <- ddev %>%
+	filter(start_year >= 2020) %>%
+	mutate(min = ddev_km2y*0.25, 
+		Iteration = "",
+		StratumID = "",
+		TransitionGroupID = "URBANIZATION",
+		Amount = as.numeric(NA),
+		DistributionType = "Uniform",
+		DistributionFrequencyID = "Iteration and Timestep",
+		DistributionSD = as.numeric(NA),
+		DistributionMin = as.numeric(NA), 
+		DistributionMax = as.numeric(NA)) %>%
+	select(Iteration, Timestep = start_year, StratumID, 
+		SecondaryStratumID = island, TransitionGroupID, 
+		Amount, DistributionType, DistributionFrequencyID, 
+		DistributionSD, DistributionMin = min, DistributionMax = ddev_km2y) 
[…]

### 2020-06-10 Transitions/code/HI_pop_urban.R

@@ -1,0 +1,67 @@
+#Hawaii population and urban area projections
+#Paul C. Selmants
+#2019-08-21 (ISO-8601)
+
+library(dplyr)
+library(tidyr)
+
+# read in projected de facto population by county (includes residents & tourists)
+# data from State of Hawaii Dept. of Business, Economic Development & Tourism
+proj_pop <- read.csv('HI_proj_pop.csv', stringsAsFactors = FALSE)
+
+#filter by Maui County and calculate population by island based on 2010
+#population distribution and convert from wide to long format
+MC <- proj_pop %>%
+	filter(county == 'Maui') %>%
+	mutate('Maui' = 0.93*population,
+		"Lana'i" = 0.02*population,
+		"Moloka'i" = 0.05*population,
+		"Kaho'olawe" = 0) %>%
+	select(year, "Kaho'olawe", "Lana'i", "Maui", "Moloka'i") %>%
+	gather(island, population, "Kaho'olawe":"Moloka'i")
+#round Maui County population values to nearest whole number
+MC$population <- round(MC$population,0) 
+	
+#combine county projections with island-specific Maui county projections to
+#get statewide population projections by island
+HI_proj <- proj_pop %>%
+	filter(county != "Maui") %>%
+	mutate(island = rep(c("Hawai'i", "O'ahu", "Kaua'i"), each = 8)) %>%
+	select(year, island, population) %>%
+	bind_rows(MC) %>%
+	arrange(island) 
+
+#calculate 2010 population density using urban area (in km2) from
+#2010 NOAA CCAP map (Impervious Surface)
+Pd <- HI_proj %>%
+	filter(year == 2010) %>%
+	mutate(urban_area = c(123.13, 0, 39.49, 4.38, 62.37, 7.05, 223.39),
+		Pd = population/urban_area) %>%
+	select(island, Pd)
+#convert 'NaN' to zero for Kaho'olawe Pd	
+Pd[is.na(Pd)] <- 0 
+
+#combine statewide population projections by island with population density
+#values by island, then calculate annual rate of change in developed area by island
+HI_ddev <- left_join(HI_proj, Pd, by = 'island') %>%
+	group_by(island) %>%
+	mutate(lagpop = lag(population),
[…]

### 2020-06-11 Output/manuscript/HI_Model_ms.Rmd
re-arranged file paths and didirectory structure

@@
 
-The study area encompasses the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^ (Figure 1).
+The study area encompassed the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^ (Figure 1).
+This land area was subdivided into a grid of 250 x 250 m simulation cells.
+The state type of each simulation cell was based on combinations of three Moisture Zones, seven Islands, and ten discrete land cover classes (Figure 1).
 
@@
 (2017).
-Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate State Classes in the simulation model."}
-knitr::include_graphics("./fig_images/fig1_HI_LC.png")
+Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
+Land cover classes not shown: Water, Wetland."}
+knitr::include_graphics("./fig_images/fig1_HI_LC.png").
 ```

### 2020-06-11 Fire/code/Fire_probabilities.R

@@ -20,26 +20,24 @@ library(stringr)
 
 			#### Read spatial data into R ####
 
-#set working directory to "HI_Model/Fire" in repo. 
-
 # read Island ID GeoTIFF into R from Build_STsim directory, where:
 # 1 = Hawaii		5 = Molokai
 # 2 = Kahoolawe		6 = Oahu
 # 3 = Lanai			7 = Kauai
 # 4 = Maui
-Islands <- raster("../Build_STsim/data/spatial_data/Island_250m.tif")
+Islands <- raster("../Model_InputData/spatial_data/Island_250m.tif")
 # read Moisture Zone GeoTIFF into R, where:
 # 1 = "Dry"
 # 2 = "Mesic"
 # 3 = "Wet"
-MZ <- raster("../Build_STsim/data/spatial_data/MZ_3_250m.tif")
+MZ <- raster("../Model_InputData/spatial_data/MZ_3_250m.tif")
 # read State Class GeoTIFF into R, where: 
 # 1 = Water 		6 = Forest
 # 2 = Urban 		7 = Grassland
 # 3 = Plantation	8 = Agriculture
 # 4 = WoodyCrop		9 = Wetland
 # 5 = Barren	   10 = Shrubland
-StateClass <- raster("../Build_STsim/data/spatial_data/StateClassNew_250m.tif")
+StateClass <- raster("../Model_InputData/spatial_data/StateClassNew_250m.tif")
 # read in Hawaii statewide annual fire perimeter shapefile (1999-2019) 
 fires <- st_read("./data/base/2019_1999_Hawaii_Fire_Perimeters.shp") %>%
 	dplyr::select(UH_ID, Year, geometry)
@@ -293,19 +291,25 @@ high_fire <- firedistribution %>%
 		ExternalVariableMin, ExternalVariableMax, Value) %>%
 	as.data.frame()
 	
-	### Write data to files, save to HI_Model/Fire/data/processed ###
+	### Write data to HI_Model/Model_InputData/spatial_data ###
 
 # write individual fire rasters to GeoTIFF files 
 writeRaster(firebrick, names(firebrick), 
 	bylayer = TRUE, format = 'GTiff', overwrite = TRUE)
 # write fire_2020 raster to GeoTIFF file
 writeRaster(fire_2020, "fire_2020.tif", overwrite = TRUE)
+
+## Write data to HI_Model/Model_InputData ##
+
 # write firedistribution to .csv file
[…]

### 2020-06-11 FlowMultipliers/code/NPPmultipliers.R

@@ -13,13 +13,17 @@ library(rgdal)
 library(tidyr, warn.conflicts = FALSE)
 library(dplyr, warn.conflicts = FALSE)
 
+
+## Set working directory at highest level in 
+## GitHub repository https://github.com/selmants/HI_Model 
+
 # read in State Class raster, where: 
 # 1 = Water 		 6 = Forest		
 # 2 = Urban 		 7 = Grassland		
 # 3 = Plantation 	 8 = Agriculture			 
 # 4 = WoodyCrop 	 9 = Wetland		 
 # 5 = Barren 		10 = Shrubland		
-SC <- raster("StateClassNew_250m.tif")
+SC <- raster("./Model_InputData/spatial_data/StateClassNew_250m.tif")
 
 # reclassification matrix for SC to set Water, Urban, 
 # Barren, & Wetland to 1 and set other classes to NA
@@ -35,7 +39,7 @@ SCnoveg <- reclassify(SC, sc_rcl)
 # 4 = Dry Plantation	10 = Mesic Plantation		16 = Wet Plantation
 # 5 = Dry Ag			11 = Mesic Ag				17 = Wet Ag
 # 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
-MZSC <- raster("MZSC.tif")
+MZSC <- raster("./Fire/data/processed/MZSC.tif")
 
 # reclassification matrix to create forest MZSC raster
 for_rcl <- c(1,1,2,NA,3,NA,4,4,5,NA,6,6,7,7,8,NA,9,NA,10,10,
@@ -51,7 +55,7 @@ shrubgrass_rcl <- c(1,NA,2,2,3,3,4,NA,5,5,6,NA,7,NA,8,8,9,9,10,NA,
 shrubgrassMZSC <- reclassify(MZSC, shrubgrass_rcl)
 
 # create list of individual island .tif files
-islandlist <- list.files('islands', full.names = TRUE)
+islandlist <- list.files('./FlowMultipliers/data/islands', full.names = TRUE)
 #create raster stack of individual island .tif files
 islandstack <- stack(islandlist)
 
@@ -64,7 +68,7 @@ names(mzsc_island) <- names(islandstack)
 ## late-century rainfall anomalies for RCP 4.5 and RCP 8.5
 
 # create list of annual rainfall .tif files
-rainlist <- list.files('rainfall_tiff_annual', full.names = TRUE)
[…]

### 2020-06-11 CarbonStockFlow_IBIS/code/HI_FlowPathways.R
Edited R script file pathways and changed folder name

@@ -8,7 +8,7 @@
 # Set working directory to "CarbonStockFlow_IBIS" folder of HI_Model
 # GitHub repository (https://github.com/selmants/HI_Model)
 
-## This R script is dependent on output from Run HI_SAV.R and HI_Flux.R scripts ##
+## This R script is dependent on output from HI_SAV.R and HI_Flux.R scripts ##
 
 #load required packages into R
 library(dplyr)

### 2020-06-11 FlowMultipliers/code/CFEmultipliers.R

@@ -6,6 +6,10 @@
 ## R version 3.6.1
 ## 2020-04-10
 
+
+## set working directory to "Model_InputData" sub-directory of 
+## HI_Model GitHub repository, https://github.com/selmants/HI_Model
+
 # load dplyr v. 0.8.3 into R
 library(dplyr)
 

### 2020-06-11 FlowMultipliers/code/NPPmultipliers.R

@@ -13,9 +13,8 @@ library(rgdal)
 library(tidyr, warn.conflicts = FALSE)
 library(dplyr, warn.conflicts = FALSE)
 
-
-## Set working directory at highest level in 
-## GitHub repository https://github.com/selmants/HI_Model 
+## Set working directory at highest level in HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model 
 
 # read in State Class raster, where: 
 # 1 = Water 		 6 = Forest		

### 2020-06-11 FlowMultipliers/code/Q10DecayMultipliers.R

@@ -6,6 +6,9 @@
 ## Paul C. Selmants
 ## 2020-03-23 (ISO 8601)
 
+## Set working directory at highest level in HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model 
+
 # load required packages into R
 library(raster)
 library(rgdal)
@@ -19,10 +22,10 @@ library(dplyr, warn.conflicts = FALSE)
 # 4 = Dry Plantation	10 = Mesic Plantation		16 = Wet Plantation
 # 5 = Dry Ag			11 = Mesic Ag				17 = Wet Ag
 # 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
-MZSC <- raster("MZSC.tif")
+MZSC <- raster("./Fire/data/processed/MZSC.tif")
 
 # create list of individual island .tif files
-islandlist <- list.files('islands', full.names = TRUE)
+islandlist <- list.files('./FlowMultipliers/data/islands', full.names = TRUE)
 #create raster stack of individual island .tif files
 islandstack <- stack(islandlist)
 
@@ -31,7 +34,8 @@ mzsc_island <- mask(MZSC, islandstack)
 names(mzsc_island) <- names(islandstack)
 
 # create list of temperature delta .tif files
-tdeltalist <- list.files(pattern = 'sd_Tdeltas', full.names = TRUE)
+tdeltalist <- list.files(path = "./FlowMultipliers/data", pattern = 'sd_Tdeltas', 
+	full.names = TRUE)
 # stack temperature delta rasters and match projection to MZSC
 tdeltastack <- stack(tdeltalist) %>%
 	projectRaster(., MZSC, method = 'bilinear')

### 2020-06-11 Transitions/code/HI_Transition_Targets.R

@@ -346,13 +346,12 @@ LowLULC_Targets <- bind_rows(list(Low_Urban, High_AgCon,
 HighLULC_Targets <- bind_rows(list(High_Urban, AgCon_Hist, 
 	High_AgEx, Zero_PlantHarvest, High_PlantHarvest))
 
-# write Low LULC Scenario transition targets to .csv file,
-# save in HI_Model/Build_STsim/data directory
-write.csv(LowLULC_Targets, "TransitionTargets_LowLULC.csv", 
+# write Low LULC Scenario transition targets to .csv file
+write.csv(LowLULC_Targets, "./Model_InputData/TransitionTargets_LowLULC.csv", 
 	row.names = FALSE, na = "") 
  
 # write High Urbanization Scenario transition targets to .csv file,
 # save in HI_Model/Build_STsim/data directory
-write.csv(HighLULC_Targets, "TransitionTargets_HighLULC.csv", 
+write.csv(HighLULC_Targets, "./Model_InputData/TransitionTargets_HighLULC.csv", 
 	row.names = FALSE, na = "") 
 

### 2020-06-11 Output/manuscript/HI_Model_ms.Rmd
edited directory pathway in R script

@@
 This land area was subdivided into a grid of 250 x 250 m simulation cells.
-The state type of each simulation cell was based on combinations of three Moisture Zones, seven Islands, and ten discrete land cover classes (Figure 1).
+Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three Moisture Zones (Figure S1), seven Islands, and ten discrete land cover classes (Figure 1).
 
@@
 Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
-Land cover classes not shown: Water, Wetland."}
+Water and Wetland land cover classes are not shown."}
 knitr::include_graphics("./fig_images/fig1_HI_LC.png").

### 2020-06-11 CarbonStockFlow_IBIS/code/HI_Events_FlowPathways.R

@@ -165,7 +165,6 @@ Events <- bind_rows(list(AgEx, Fire_Forest, Fire_Grassland, Fire_Plantation,
 		FromStockTypeID, ToStratumID, ToStateClassID, ToAgeMin, ToStockTypeID, 
 		TransitionGroupID, StateAttributeTypeID, FlowTypeID, Multiplier)
 
-# write Events dataframe to .csv file, save in 
-# HI_Model/Build_STsim/data
-write.csv(Events, "./data/FlowPathways_Events.csv", row.names = FALSE)
+# write Events dataframe to .csv file
+write.csv(Events, "./InputData/FlowPathways_Events.csv", row.names = FALSE)
 

### 2020-06-11 FlowMultipliers/code/CFEmultipliers.R

@@ -7,7 +7,7 @@
 ## 2020-04-10
 
 
-## set working directory to "Model_InputData" sub-directory of 
+## set working directory to "InputData" folder of 
 ## HI_Model GitHub repository, https://github.com/selmants/HI_Model
 
 # load dplyr v. 0.8.3 into R

### 2020-06-11 FlowMultipliers/code/MZSC.R

@@ -16,17 +16,17 @@ library(dplyr)
 # 2 = Kahoolawe		6 = Oahu
 # 3 = Lanai			7 = Kauai
 # 4 = Maui
-Islands <- raster("./Build_STsim/data/spatial_data/Island_250m.tif")
+Islands <- raster("./InputData/spatial_data/Island_250m.tif")
 # read Moisture Zone TIFF into R, where:
 # 1 = "Dry", 2 = "Mesic", 3 = "Wet"
-MZ <- raster('./Build_STsim/data/spatial_data/MZ_3_250m.tif')
+MZ <- raster('./InputData/spatial_data/MZ_3_250m.tif')
 # read State Class TIFF into R, where: 
 # 1 = Water 		6 = Forest
 # 2 = Urban 		7 = Grassland
 # 3 = Plantation	8 = Agriculture
 # 4 = WoodyCrop		9 = Wetland
 # 5 = Barren	   10 = Shrubland
-StateClass <- raster('./Build_STsim/data/spatial_data/StateClassNew_250m.tif')
+StateClass <- raster('./InputData/spatial_data/StateClassNew_250m.tif')
 
 # make reclassification matrix to make all islands values = 0
 islands_recl <- c(0, NA, 1, 0, 2, 0, 3, 0, 4, 0, 5, 0, 6, 0, 

### 2020-06-11 FlowMultipliers/code/NPPmultipliers.R

@@ -22,7 +22,7 @@ library(dplyr, warn.conflicts = FALSE)
 # 3 = Plantation 	 8 = Agriculture			 
 # 4 = WoodyCrop 	 9 = Wetland		 
 # 5 = Barren 		10 = Shrubland		
-SC <- raster("./Model_InputData/spatial_data/StateClassNew_250m.tif")
+SC <- raster("./InputData/spatial_data/StateClassNew_250m.tif")
 
 # reclassification matrix for SC to set Water, Urban, 
 # Barren, & Wetland to 1 and set other classes to NA
@@ -230,7 +230,7 @@ growthmultipliers <- overlay(presentNPP, NPPMZSC,
 # merge growthmultipliers with non-veg State Class raster
 gm <- merge(growthmultipliers, SCnoveg)
 # save NPP multipliers raster as NCEAS_NPP_sm
-writeRaster(gm, "../processed/NCEAS_NPP_sm.tif", overwrite = TRUE)
+writeRaster(gm, "./InputData/spatial_data/NCEAS_NPP_sm.tif", overwrite = TRUE)
 
 ## Create future NPP flow multiplier tables for RCPs 4.5 & 8.5 ##
 
@@ -576,6 +576,6 @@ NPPmulti45 <- gmall(gm45list)
 NPPmulti85 <- gmall(gm85list)
 
 # write NPP multiplier data frames to .csv files
-write.csv(NPPmulti45, "./Model_InputData/RCP45_NPPmultipliers.csv", row.names = FALSE)
-write.csv(NPPmulti85, "./Model_InputData/RCP85_NPPmultipliers.csv", row.names = FALSE)
+write.csv(NPPmulti45, "./InputData/RCP45_NPPmultipliers.csv", row.names = FALSE)
+write.csv(NPPmulti85, "./InputData/RCP85_NPPmultipliers.csv", row.names = FALSE)
 

### 2020-06-11 FlowMultipliers/code/Q10DecayMultipliers.R

@@ -504,8 +504,8 @@ Q1085NoForKaho <- filter(Q10FlowMulti85,
 	SecondaryStratumID != "Kaho'olawe" | StateClassID != "Forest:All")
 
 #write modified Q10 Flow Multiplier Tables to .csv files
-write.csv(Q1045NoForKaho, "../processed/RCP45Q10multi_NoKahoFor.csv", row.names = FALSE)
-write.csv(Q1085NoForKaho, "../processed/RCP85Q10multi_NoKahoFor.csv", row.names = FALSE)
+write.csv(Q1045NoForKaho, "./InputData/RCP45Q10multi_NoKahoFor.csv", row.names = FALSE)
+write.csv(Q1085NoForKaho, "./InputData/RCP85Q10multi_NoKahoFor.csv", row.names = FALSE)
 
 
 

### 2020-06-11 Transitions/code/CCAP_LULCC.R

@@ -7,12 +7,18 @@
 library(dplyr)
 library(tidyr)
 
+## Set R working directory to highest level of HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model
+
 #make a list of NOAA CCAP 1992-2001 land cover change files
-ccap01_list <- list.files(pattern = "LCchange_1992")
+ccap01_list <- list.files(path = "./Transitions/data/base/", 
+	pattern = "LCchange_1992")
 #make a list of NOAA CCAP 2001-2005 land cover change files
-ccap05_list <- list.files(pattern = "LCchange_2001")
+ccap05_list <- list.files(path = "./Transitions/data/base/", 
+	pattern = "LCchange_2001")
 #make a list of NOAA CCAP 2001-2005 land cover change files
-ccap10_list <- list.files(pattern = "LCchange_2005") 
+ccap10_list <- list.files(path = "./Transitions/data/base/", 
+	pattern = "LCchange_2005") 
 
 #read in ccap 1992-2001 files, bind together, filter out zero values
 LC92_01 <- lapply(ccap01_list[c(1:4)], FUN = read.csv, header = TRUE,
@@ -199,8 +205,9 @@ lulc_minmax <- bind_rows(lulc1992_2001, lulc2001_2005, lulc2005_2010) %>%
 		maximum = max(rate_km2y)) %>%
 	data.frame(.) 
 
-#write lulcminmax to .csv file, save in Transistions/data/processed
-write.csv(lulc_minmax, "lulc_historic.csv", row.names = FALSE)
+#write lulcminmax to .csv file
+write.csv(lulc_minmax, "./Transitions/data/processed/lulc_historic.csv", 
+	row.names = FALSE)
 	
 
 

### 2020-06-11 Transitions/code/HI_Transition_Pathways_Fire.R

@@ -6,6 +6,9 @@
 ## UPDATED with new fire data 2019-10-31
 ## R version 3.6.1
 
+## Set R working directory to highest level of HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model
+
 # load dplyr package v. 0.8.3
 library(dplyr)
 
@@ -142,6 +145,6 @@ LULC <- bind_rows(list(Urban, AgCon,
 	mutate(TSTMin = NA,
 		TSTMax = NA, 
 		TSTRelative = NA)
-# write Transition Pathways to csv file save in 
-write.csv(LULC, "TransitionPathways_Fire.csv", row.names = FALSE, na = "")
+# write Transition Pathways to csv file 
+write.csv(LULC, "./InputData/TransitionPathways_Fire.csv", row.names = FALSE, na = "")
 

### 2020-06-11 Transitions/code/HI_Transition_Targets.R

@@ -347,11 +347,11 @@ HighLULC_Targets <- bind_rows(list(High_Urban, AgCon_Hist,
 	High_AgEx, Zero_PlantHarvest, High_PlantHarvest))
 
 # write Low LULC Scenario transition targets to .csv file
-write.csv(LowLULC_Targets, "./Model_InputData/TransitionTargets_LowLULC.csv", 
+write.csv(LowLULC_Targets, "./InputData/TransitionTargets_LowLULC.csv", 
 	row.names = FALSE, na = "") 
  
 # write High Urbanization Scenario transition targets to .csv file,
 # save in HI_Model/Build_STsim/data directory
-write.csv(HighLULC_Targets, "./Model_InputData/TransitionTargets_HighLULC.csv", 
+write.csv(HighLULC_Targets, "./InputData/TransitionTargets_HighLULC.csv", 
 	row.names = FALSE, na = "") 
 

### 2020-06-11 Transitions/code/HI_pop_urban.R

@@ -5,9 +5,12 @@
 library(dplyr)
 library(tidyr)
 
+# set R working directory to highest level of HI_Model
+# GitHub repository, https://github.com/selmants/HI_Model
+
 # read in projected de facto population by county (includes residents & tourists)
 # data from State of Hawaii Dept. of Business, Economic Development & Tourism
-proj_pop <- read.csv('HI_proj_pop.csv', stringsAsFactors = FALSE)
+proj_pop <- read.csv('./Transitions/data/base/HI_proj_pop.csv', stringsAsFactors = FALSE)
 
 #filter by Maui County and calculate population by island based on 2010
 #population distribution and convert from wide to long format
@@ -62,6 +65,6 @@ HI_ddev$ddev_km2y[29] <- 1.1399 #Maui
 HI_ddev$ddev_km2y[36] <- 0.1357 #Moloka'i
 HI_ddev$ddev_km2y[43] <- 1.8844 #O'ahu
 
-#write ddev_range to .csv file, save in Transistions/data/processed
-write.csv(HI_ddev, "HI_ddev.csv", row.names = FALSE)
+#write ddev_range to .csv file
+write.csv(HI_ddev, "./Transitions/data/processed/HI_ddev.csv", row.names = FALSE)
 

### 2020-06-18 Output/manuscript/HI_Model_ms.Rmd
added age and transition spatial multiplier rasters

@@
    - \doublespacing
-latex_engine: pdflatex  
-output: pdf_document
+output: 
+  pdf_document: 
+    latex_engine: xelatex
 ---
@@
 ^2^U.S. Geological Survey, Seattle, WA, USA  
-^3^University of Hawaii at Manoa, Honolulu, HI, USA  
+^3^University of Hawai`i at Mānoa, Honolulu, HI, USA  
 ^4^The East-West Center, Honolulu, HI, USA  
@@
 
-The State of Hawaii recently passed legislation setting a goal to be carbon neutral by 2045.
+The State of Hawai`i recently passed legislation setting a goal to be carbon neutral by 2045.
 Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain.
@@
 Water and Wetland land cover classes are not shown."}
-knitr::include_graphics("./fig_images/fig1_HI_LC.png").
+knitr::include_graphics("./fig_images/fig1_HI_LC.png")
 ```
@@
 ## Acknowledgements
+
+This study was funded by the Biological Sequestration Program of the U.S.
+Geological Survey.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.

### 2020-06-18 CarbonStockFlow_IBIS/code/HI_Events_FlowPathways.R

@@ -1,6 +1,6 @@
 ## Disturbance Event Flow Pathways
 ## for HI LUCAS Model 
-## Soil flux from Don et al. (2010)
+## Soil flux from Don et al. (2011)
 ## Other fluxes from Sleeter et al. (2017)
 
 ## Paul C. Selmants 

### 2020-06-18 Output/manuscript/fig_code/HI_LandCoverFig.R

@@ -17,10 +17,10 @@ library(ggplot2)
 # 3 = Plantation	8 = Agriculture
 # 4 = WoodyCrop		9 = Wetland
 # 5 = Barren	   10 = Shrubland
-StateClass <- raster('./Build_STsim/data/spatial_data/StateClassNew_250m.tif')
+StateClass <- raster('./InputData/spatial_data/StateClassNew_250m.tif')
 
 # read annual MAT Tiff into R
-MAT <- raster("./FlowMultipliers/data/base/MAT_annual.tif")
+MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
 
 # make reclassification matrix for simplified land cover raster
 lc_rcl <- c(1, NA, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6,
@@ -61,20 +61,20 @@ lcmap <- ggplot(data=lc_df) +
     legend.background = element_rect(fill="white", 
     	linetype = "solid", size = 0.5, color = "white")) +
     coord_fixed() +
-    annotate(geom = "text", x= -155, y = 20.15, label = "Hawaii Island", 
+    annotate(geom = "text", x= -155, y = 20.15, label = "Hawaiʻi Island", 
         size = 3.5) +
     annotate(geom = "text", x= -156.025, y = 20.95, label = "Maui",
         size = 3.5) +
-    annotate(geom = "text", x= -157, y = 20.55, label = "Kahoolawe",
+    annotate(geom = "text", x= -157, y = 20.55, label = "Kahoʻolawe",
         size = 3.5) +
-    annotate(geom = "text", x= -157.2, y = 20.8, label = "Lanai",
+    annotate(geom = "text", x= -157.2, y = 20.8, label = "Lānaʻi",
         size = 3.5) +
-    annotate(geom = "text", x= -156.78, y = 21.275, label = "Molokai",
+    annotate(geom = "text", x= -156.78, y = 21.275, label = "Molokaʻi",
         size = 3.5) +
-    annotate(geom = "text", x= -157.7, y = 21.65, label = "Oahu",
+    annotate(geom = "text", x= -157.7, y = 21.65, label = "Oʻahu",
         size = 3.5) +
-    annotate(geom = "text", x= -159.1, y = 22.2, label = "Kauai",
+    annotate(geom = "text", x= -159.1, y = 22.2, label = "Kauaʻi",
         size = 3.5)
 
[…]

### 2020-06-19 Output/manuscript/fig_code/HI_LandCoverFig.R
edits to manuscript & figures

@@ -65,7 +65,7 @@ lcmap <- ggplot(data=lc_df) +
         size = 3.5) +
     annotate(geom = "text", x= -156.025, y = 20.95, label = "Maui",
         size = 3.5) +
-    annotate(geom = "text", x= -157, y = 20.55, label = "Kahoʻolawe",
+    annotate(geom = "text", x= -157.025, y = 20.55, label = "Kahoʻolawe",
         size = 3.5) +
     annotate(geom = "text", x= -157.2, y = 20.8, label = "Lānaʻi",
         size = 3.5) +

### 2020-06-19 Output/manuscript/fig_code/HI_StateClassArea.R

@@ -1,5 +1,6 @@
 ## Processing HI LUCAS model output:
-## summarizing State Class areas over time
+## summarizing area of different  
+## State Classes over time
 ## by land use scenario
 ## Paul C. Selmants
 ## 2020-04-22
@@ -11,12 +12,10 @@ library(cowplot)
 library(stringr)
 library(tidyr) 
 
-# read State Class model output data into R 
+# read State Class model output data into R and summarize 
+# statewide by land use scenario
 sc <- read.csv("StateClassArea.csv", stringsAsFactors = FALSE) %>%
-	mutate(StateClassID = str_remove(.$StateClassID, ":All"))
-
-# summarize State Class Area statewide by Land Use Scenario
-StateClass <- sc %>%
+	mutate(StateClassID = str_remove(.$StateClassID, ":All")) %>%
 	group_by(Scenario, Timestep, Iteration, StateClassID) %>%
 	summarize(area = sum(Area)) %>%
 	separate(Scenario, c("Climate", "LandUse"), sep = 5) %>%
@@ -25,9 +24,9 @@ StateClass <- sc %>%
 		area_min = min(area), 
 		area_max = max(area)) %>%
 	data.frame() 
-
+ 
 # ggplot faceted land cover change graph
-scfig <- StateClass %>%
+scfig <- sc %>%
 filter(StateClassID %in% c("Agriculture",
 	"Developed", "Forest", "Grassland", "Shrubland")) %>% 
 	{ggplot(., aes(x = Timestep, y = area_mean, color = LandUse)) +
@@ -43,37 +42,9 @@ filter(StateClassID %in% c("Agriculture",
 		fill = "Land Use\nScenario") +
 	theme_bw() +
 	theme(panel.grid.minor = element_blank(), 
-		legend.position = c(0.825,0.25),
+		legend.position = c(0.85,0.25),
 		legend.text = element_text(size=10))}
-	
 
-# write function to create ggplot graphs of land cover change
-fig <- function(a,b,c) {
-	ggplot(filter(StateClass, StateClassID == a), 
-	aes(x = Timestep, y = area_mean, color = LandUse)) +
-	geom_ribbon(aes(ymin = area_min, ymax = area_max, fill=LandUse),
-		alpha = 0.4, color = NA) +
-	geom_line(size = 0.75) +
[…]

### 2020-07-01 Manuscript/HI_Model_ms.Rmd
re-organized directory structure

--- 2020-06-18 Output/manuscript/HI_Model_ms.Rmd
@@
 ## Abstract
-
 The State of Hawai`i recently passed legislation setting a goal to be carbon neutral by 2045.
@@
 ## Methods
+We used the Land Use and Carbon Scenario Simulator (LUCAS) to project changes in ecosystem carbon balance for the main Hawaiian Islands under four unique scenarios incorporating different levels of climate and land-use change.
+The LUCAS model is a state-and-transition simulation model ...
 

### 2020-07-01 Manuscript/fig_code/AnnualFire_plot.R

@@ -1,0 +1,58 @@
+## Graph of Annual Area Burned, 
+## State of Hawaii, 1999-2019
+## Paul C. Selmants
+## R version 3.6.1
+## 2019-10-26
+
+## Data in GitHub repository: https://github.com/selmants/HI_Model
+
+## This R script is dependent on output from Fire_probabilities.R script ## 
+
+# load required R packages
+library(ggplot2)
+library(dplyr)
+library(cowplot)
+
+# read statewide annual burned area data into R
+data <- read.csv('./Fire/data/processed/AnnualFire.csv', stringsAsFactors = FALSE) 
+
+# line graph of annual area burned statewide
+stateplot <- data %>%
+	group_by(year) %>%
+	summarize(burnarea_km2 = sum(burnarea_km2)) %>%
+	as.data.frame()%>%
+		{ggplot(., aes(year, burnarea_km2)) +
+		geom_line(color = "#67000D", size = 0.6) +
+		geom_point(color = "#67000D", size = 1.5, shape = 25, fill = "#67000D") +
+		labs(x = "Year",
+		y = expression(Area~burned~(km^{2}))) +
+		scale_x_continuous(limits = c(1998, 2019), breaks = seq(1998, 2018, 4)) +
+		scale_y_continuous(limits = c(0, 175), breaks = seq(0, 175, 25)) +
+		geom_hline(aes(yintercept=28.75), linetype="dashed", size = 0.35) +
+		theme_bw() +
+		theme(panel.grid.minor=element_blank(),
+		axis.title = element_text(size = 10),
+		axis.text = element_text(size = 8))}
+
+# facet plot of annual area burned by island (Hawai, Maui, Oahu, Kauai)
+islandplot <- data %>%
+	filter(island %in% c("Hawaii", "Maui", "Oahu", "Kauai")) %>%
+	{ggplot(., aes(year, burnarea_km2)) +
+		facet_wrap(~ island) +
+		geom_line(color = "#67000D", size = 0.35) +
+		geom_point(color = "#67000D", size = 0.45, shape = 25, fill = "#67000D") +
+		labs(x = "Year",
+			y = expression(Area~burned~(km^{2}))) +
+		scale_x_continuous(limits = c(1998, 2019), breaks = seq(1998, 2018, 5)) +
+		scale_y_continuous(limits = c(0, 138.875), breaks = seq(0, 125, 25)) +
+		theme_bw() +
[…]

### 2020-07-01 Manuscript/fig_code/HI_CFE_CarbonBalance.R

@@ -1,0 +1,152 @@
+# CFE carbon balance 
+# Hawaii LUCAS Model 
+# Paul C. Selmants
+# 2020-04-14
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+
+## Download LUCAS model output data from USGS ScienceBase: 
+## https://doi.org/10.5066/P9AWLFKZ
+
+# read LUCAS CFE C stocks from RCP85High output data into R
+CFE_Cstocks <- read.csv("Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in LUCAS C stocks by scenario and filter on RCP85High scenario 
+Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_rate = 0) %>%
+	select(CFE_rate, Timestep:Amount)
+
+# read in CFE base flow data from RCP85High Scenario
+cfe_baseflow <- read.csv("BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in CFE C loss data (transition, fire, harvest)
+cfe_closs <- read.csv("TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+	select(-Scenario)
+# read in zero CFE base flow data 
+baseflowzero <- read.csv("BaseFlows.csv", stringsAsFactors = FALSE) %>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_Rate = 0) %>%
+	select(CFE_Rate, Timestep:Amount)
+# read zero CFE non-Rh C Loss data 
+closszero <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
+	filter(Scenario == "RCP85High") %>%
+	mutate(CFE_Rate = 0) %>%
+	select(CFE_Rate, Timestep:Amount)
+
+## Summarize C stock data and create TEC by CFE figure ## 
+
+# combine CFE Cstock data together and summarize as Statewide TEC 
+cfetec <- bind_rows(Cstocks, CFE_Cstocks) %>%
+	filter(CFE_rate %in% c(0, 5, 10, 15)) %>%
+	group_by(CFE_rate, Timestep, SecondaryStratumID, StockGroupID) %>%
[…]

### 2020-07-01 Manuscript/fig_code/HI_CarbonBalance.R

@@ -1,0 +1,235 @@
+## Processing HI LUCAS model output:
+## summarizing Flows to estimate annual
+## and cumulative net biome productivity (NBP)
+## Paul C. Selmants
+## 2020-05-20
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(cowplot)
+
+## Download LUCAS model output data from USGS ScienceBase: 
+## https://doi.org/10.5066/P9AWLFKZ
+
+# read HI LUCAS baseline C flow data into R
+baseflow <- read.csv("BaseFlows.csv", stringsAsFactors = FALSE)
+# read HI LUCAS Transition/Harvest C Loss data into R
+transflow <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE)
+
+# calculate statewide NPP
+NPP <- baseflow %>%
+	filter(FlowGroupID == "Growth [Type]") %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(NPP = sum(Amount)) %>%
+	as.data.frame()	
+# calculate statewide baseline Rh 
+Rh <- baseflow %>%
+	filter(FlowGroupID %in% c("Emission (litter) [Type]",
+		"Emission (soil) [Type]")) %>% 
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(Rh = sum(Amount)) %>%
+	as.data.frame()
+# calculate statewide C loss by leaching/erosion
+Leach <- baseflow %>%
+	filter(FlowGroupID == "Leaching [Type]") %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(Leach = sum(Amount)) %>%
+	as.data.frame()
+# calculate disturbance C losses (∆LULC, fire, ag harvest)
+LUCloss <- transflow %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(LUCloss = sum(Amount)) %>%
+	as.data.frame()
+
+# join NPP and C loss dataframes together
+cflowjoin <- left_join(NPP, Rh) %>%
+	left_join(., Leach) %>%
+	left_join(., LUCloss) 
+
+# convert c flow data to long format, separate Scenario column into 
+# RCP and LandUse, summarize flow data by mean, min, and max in Tg 
+cflows <- cflowjoin %>% 
+	pivot_longer(cols = NPP:LUCloss, 
[…]

### 2020-07-01 Manuscript/fig_code/HI_CarbonStocks.R

@@ -1,0 +1,116 @@
+## HI LUCAS model output:
+## summarize & graph Statewide 
+## Carbon stocks, fluxes, and balance
+## Paul C. Selmants
+## 2020-05-20
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+
+## Download LUCAS model output data from USGS ScienceBase: 
+## https://doi.org/10.5066/P9AWLFKZ
+
+# read Hawaii LUCAS model C stock output data into R
+Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE) %>%
+	group_by(Scenario, Timestep, SecondaryStratumID, StockGroupID) %>%
+	summarize(CstockTg = mean(Amount)/1000,
+		Cstock_min = min(Amount)/1000,
+		Cstock_max = max(Amount)/1000) %>%
+	data.frame()
+# remove " [Type]" at trailing end of all stock types  
+Cstocks$StockGroupID <- gsub('.{7}$', '', Cstocks$StockGroupID) 
+
+# create function to summarize carbon stocks statewide
+stocksum <- function(x) {
+	 x %>%
+	group_by(Scenario, Timestep) %>%
+	summarize(C_Tg = sum(CstockTg), 
+		Cmin = sum(Cstock_min),
+		Cmax = sum(Cstock_max)) %>%
+	select(Scenario, Timestep, C_Tg, Cmin, Cmax) %>%
+	arrange(Scenario, Timestep) %>%
+	as.data.frame() 
+}
+
+# summarize TEC and individual C stocks statewide
+TEC <- stocksum(Cstocks)
+# sum of standing and down dead wood
+Deadwood <- Cstocks %>%
+	filter(grepl("Deadwood", StockGroupID)) %>%
+	stocksum(.)
+# Living biomass
+LB <- Cstocks %>%
+	filter(StockGroupID == "Living Biomass") %>%
+	stocksum(.)
+# Litter
+Litter <- Cstocks %>%
+	filter(StockGroupID == "Litter") %>%
+	stocksum(.)
+# Soil organic carbon
+SOC <- Cstocks %>%
+	filter(StockGroupID == "Soil") %>%
+	stocksum(.)
+
+# merge individual C stocks into single dataframe
+indCstocks <- bind_rows(Deadwood, LB, Litter, SOC) %>%
[…]

### 2020-07-01 Manuscript/fig_code/HI_ClimateFigs.R

@@ -1,0 +1,161 @@
+# Use ggplot to visualize
+# current and end-of-century
+# rainfall and air temperature 
+# in the Hawaiian Islands 
+# Paul C. Selmants
+# 2020-02-24
+
+## Data in GitHub repository: https://github.com/selmants/HI_Model
+
+# load required R packages
+library(raster)
+library(rgdal)
+library(ggplot2)
+library(dplyr)
+library(viridis)
+library(RColorBrewer)
+library(cowplot) 
+
+# create list of annual rainfall .tif files
+rainlist <- list.files('rainfall_tiff_annual', full.names = TRUE)
+#create raster stack of annual rainfall (mm) .tif files
+rainstack <- stack(rainlist)
+# make raster layer of 30-year mean annual rainfall and re-sample
+# to change projection to UTM (this takes ~30 seconds)
+rain30ymean <- calc(rainstack, mean)
+# read in rasters of late century rainfall anomolies for RCPs 4.5 & 8.5
+rcp85anom <- raster("sd_rfanom_rcp85_late.tif")
+rcp45anom <- raster("sd_rfanom_rcp45_late.tif")
+# make rasters of late century rainfall differences from present
+raindiff85 <- rain30ymean*(rcp85anom/100)
+raindiff45 <- rain30ymean*(rcp45anom/100)
+
+# read in current mean annual temperature raster
+MAT <- raster("MAT_annual.tif") 
+# read in rasters of late century temperature differences for RCPs 4.5 & 8.5
+Tdelta_rcp45 <- raster("sd_Tdeltas_rcp45_late.tif")
+Tdelta_rcp85 <- raster("sd_Tdeltas_rcp85_late.tif")
+
+#prepare rain30ymean and raindiff data to plot with geom_tile in ggplot
+rain_df <- rasterToPoints(rain30ymean) %>%
+	data.frame(.)
+
+r45_df <- rasterToPoints(raindiff45) %>%
+	data.frame(.)
+
+r85_df <- rasterToPoints(raindiff85) %>%
+	data.frame(.)
+#prepare MAT and dT data to plot with geom_tile in ggplot2 
+MAT_df <- rasterToPoints(MAT) %>%
+	data.frame(.) %>%
+	select(x, y, layer = MAT_annual)
+
+t45_df <- rasterToPoints(Tdelta_rcp45) %>%
[…]

### 2020-07-01 Manuscript/fig_code/HI_LandCoverFig.R

@@ -1,0 +1,80 @@
+# Using ggplot2 to visualize
+# Hawaiʻi land cover classes
+# Paul C. Selmants
+# 2020-02-24
+
+# load required packages
+library(raster)
+library(rgdal)
+library(dplyr)
+library(ggplot2)
+
+## Data in GitHub repository: https://github.com/selmants/HI_Model
+
+# read State Class TIFF into R, where: 
+# 1 = Water 		6 = Forest
+# 2 = Urban 		7 = Grassland
+# 3 = Plantation	8 = Agriculture
+# 4 = WoodyCrop		9 = Wetland
+# 5 = Barren	   10 = Shrubland
+StateClass <- raster('./InputData/spatial_data/StateClassNew_250m.tif')
+
+# read annual MAT Tiff into R
+MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
+
+# make reclassification matrix for simplified land cover raster
+lc_rcl <- c(1, NA, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6,
+	7, 7, 8, 4, 9, NA, 10, 10) %>%
+	matrix(., ncol = 2, byrow = TRUE)
+# reclassify StateClass to create simplified land cover raster, where: 
+# 2 = Urban 		 6 = Forest
+# 3 = Plantation	 7 = Grassland
+# 4 = Agriculture	10 = Shrubland
+# 5 = Barren	
+lc <- reclassify(StateClass, lc_rcl)
+# re-sample lc projection to Lat. Long. using MAT
+lcdeg <- projectRaster(lc, MAT, method = 'ngb')
+# convert lc raster to points dataframe for usage in ggplot 
+lc_df <- rasterToPoints(lcdeg) %>%
+	data.frame(.) %>%
+	select(x, y, layer = StateClassNew_250m) 
+# convert layer column to factor 
+lc_df$layer <- as.factor(lc_df$layer)
+# Create custom colorblind palette for land cover classes
+lcpal <- c("#000000", "#CC79A7", "#A6761D", "#969696", "#006D2C", 
+	"#FFD92F", "#B2DF8A") 
+# create statewide land cover map for Hawaii
+lcmap <- ggplot(data=lc_df) +
+	geom_tile(aes(x=x, y=y, fill=layer)) +
+	xlab(expression("Longitude " ( degree*W))) +
+	ylab(expression("Latitude " ( degree*N))) +
[…]

### 2020-07-01 Manuscript/fig_code/HI_StateClassArea.R

@@ -1,0 +1,60 @@
+## Processing HI LUCAS model output:
+## summarizing area of different  
+## State Classes over time
+## by land use scenario
+## Paul C. Selmants
+## 2020-04-22
+
+# load required R packages
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(tidyr) 
+
+## Download LUCAS model output data from USGS ScienceBase: 
+## https://doi.org/10.5066/P9AWLFKZ
+
+# read State Class model output data into R and summarize 
+# statewide by land use scenario
+sc <- read.csv("StateClassArea.csv", stringsAsFactors = FALSE) %>%
+	mutate(StateClassID = str_remove(.$StateClassID, ":All")) %>%
+	group_by(Scenario, Timestep, Iteration, StateClassID) %>%
+	summarize(area = sum(Area)) %>%
+	separate(Scenario, c("Climate", "LandUse"), sep = 5) %>%
+	group_by(LandUse, Timestep, StateClassID) %>%
+	summarize(area_mean = mean(area), 
+		area_min = min(area), 
+		area_max = max(area)) %>%
+	data.frame() 
+
+#Make Land Use color palette
+lu_pal <- c("#A6761D","#1B9E77")
+
+# ggplot faceted land cover change graph
+scfig <- sc %>%
+filter(StateClassID %in% c("Agriculture",
+	"Developed", "Forest", "Grassland", "Shrubland")) %>% 
+	{ggplot(., aes(x = Timestep, y = area_mean, color = LandUse)) +
+	facet_wrap(~ StateClassID, scales = "free_y") +
+	geom_ribbon(aes(ymin = area_min, ymax = area_max, fill = LandUse),
+		alpha = 0.4, color = NA) +
+	geom_line(size = 0.5) +
+	scale_fill_manual(values = lu_pal, labels=c("High", "Low"),
+			name = "Land use\nscenario") +
+	scale_colour_manual(values = lu_pal, labels=c("High", "Low"), 
+			name = "Land use\nscenario") +
+	scale_x_continuous(limits = c(2000,2100), breaks = seq(2000,2100,20)) +
+	scale_y_continuous(breaks = scales::pretty_breaks(5), limits = c(NA, NA)) +
+	labs(x = "Year", 
+		 y = expression(Land~area~(km^{2})), 
+		color = "Land Use\nScenario",
+		fill = "Land Use\nScenario") +
+	theme_bw() +
[…]

### 2020-07-01 Manuscript/fig_code/MZfig.R

@@ -1,0 +1,55 @@
+# Use ggplot to visualize
+# Moisture Zones in the 
+# Hawaiian Islands
+# Paul C. Selmants
+# 2020-06-23
+
+# load required packages
+library(raster)
+library(rgdal)
+library(dplyr)
+library(ggplot2)
+
+## Data in GitHub repository: https://github.com/selmants/HI_Model
+
+# read Moisture Zone TIFF into R, where: 
+# 1 = Dry 		
+# 2 = Mesic 		
+# 3 = Wet	
+MZ <- raster('./InputData/spatial_data/MZ_3_250m.tif')
+
+# read annual MAT Tiff into R
+MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
+
+# re-sample MZ projection to Lat. Long. using MAT
+MZdeg <- projectRaster(MZ, MAT, method = 'ngb')
+# convert lc raster to points dataframe for usage in ggplot 
+MZ_df <- rasterToPoints(MZdeg) %>%
+	data.frame(.) %>%
+	select(x, y, layer = MZ_3_250m) 
+# convert layer column to factor 
+MZ_df$layer <- as.factor(MZ_df$layer)
+# Create custom palette for MZ classes
+mzpal <- c("#FDBF6F", "#B2DF8A", "#1F78B4")
+
+# create statewide Moisture Zone map for Hawaii
+mzmap <- ggplot(data=MZ_df) +
+	geom_tile(aes(x=x, y=y, fill=layer)) +
+	xlab(expression("Longitude " ( degree*W))) +
+	ylab(expression("Latitude " ( degree*N))) +
+	scale_fill_manual("Moisture Zone", values = mzpal,
+		 labels=c("Dry", "Mesic", "Wet")) +
+	theme_bw() + 
+	theme(axis.title.x = element_text(size=10),
+        axis.text.x  = element_text(size=8),
+        axis.title.y = element_text(size=10),
+        axis.text.y = element_text(size=8),
+    legend.position = c(0.188, 0.465),
+    legend.justification = c("left", "top"),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.background = element_rect(fill="white", 
+    	linetype = "solid", size = 0.5, color = "white"))
+
+# save Moisture Zone map as .png file 
+ggsave("figS1_MZ.png", height = 5, width = 7, dpi = 300)

### 2020-07-01 Output/manuscript/fig_code/AnnualFire_plot.R

@@ -1,60 +1,0 @@
-## Graph of Annual Area Burned, 
-## State of Hawaii, 1999-2019
-## Paul C. Selmants
-## R version 3.6.1
-## 2019-10-26
-
-# GitHub repository (https://github.com/selmants/HI_Model)
-
-## This R script is dependent on output from Fire_probabilities.R script ## 
-
-# load required R packages
-library(ggplot2)
-library(dplyr)
-library(cowplot)
-
-
-# read statewide annual burned area data into R
-data <- read.csv('./Fire/data/processed/AnnualFire.csv', stringsAsFactors = FALSE) 
-
-# line graph of annual area burned statewide
-stateplot <- data %>%
-	group_by(year) %>%
-	summarize(burnarea_km2 = sum(burnarea_km2)) %>%
-	as.data.frame()%>%
-		{ggplot(., aes(year, burnarea_km2)) +
-		geom_line(color = "#67000D", size = 0.6) +
-		geom_point(color = "#67000D", size = 1.5, shape = 25, fill = "#67000D") +
-		labs(x = "Year",
-		y = expression(Area~burned~(km^{2}))) +
-		scale_x_continuous(limits = c(1998, 2019), breaks = seq(1998, 2018, 4)) +
-		scale_y_continuous(limits = c(0, 175), breaks = seq(0, 175, 25)) +
-		geom_hline(aes(yintercept=28.75), linetype="dashed", size = 0.35) +
-		theme_bw() +
-		theme(panel.grid.minor=element_blank(),
-		axis.title = element_text(size = 10),
-		axis.text = element_text(size = 8))}
-
-# facet plot of annual area burned by island (Hawai, Maui, Oahu, Kauai)
-islandplot <- data %>%
-	filter(island %in% c("Hawaii", "Maui", "Oahu", "Kauai")) %>%
-	{ggplot(., aes(year, burnarea_km2)) +
-		facet_wrap(~ island) +
-		geom_line(color = "#67000D", size = 0.35) +
-		geom_point(color = "#67000D", size = 0.45, shape = 25, fill = "#67000D") +
-		labs(x = "Year",
-			y = expression(Area~burned~(km^{2}))) +
-		scale_x_continuous(limits = c(1998, 2019), breaks = seq(1998, 2018, 5)) +
-		scale_y_continuous(limits = c(0, 138.875), breaks = seq(0, 125, 25)) +
-		theme_bw() +
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_CFE_CarbonBalance.R

@@ -1,148 +1,0 @@
-# CFE carbon balance 
-# Hawaii LUCAS Model 
-# Paul C. Selmants
-# 2020-04-14
-
-# load required R packages
-library(dplyr)
-library(ggplot2)
-library(cowplot)
-
-# read LUCAS CFE C stocks from RCP85High output data into R
-CFE_Cstocks <- read.csv("Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
-	select(-Scenario)
-# read in LUCAS C stocks by scenario and filter on RCP85High scenario 
-Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
-	filter(Scenario == "RCP85High") %>%
-	mutate(CFE_rate = 0) %>%
-	select(CFE_rate, Timestep:Amount)
-
-# read in CFE base flow data from RCP85High Scenario
-cfe_baseflow <- read.csv("BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
-	select(-Scenario)
-# read in CFE C loss data (transition, fire, harvest)
-cfe_closs <- read.csv("TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
-	select(-Scenario)
-# read in zero CFE base flow data 
-baseflowzero <- read.csv("BaseFlows.csv", stringsAsFactors = FALSE) %>%
-	filter(Scenario == "RCP85High") %>%
-	mutate(CFE_Rate = 0) %>%
-	select(CFE_Rate, Timestep:Amount)
-# read zero CFE non-Rh C Loss data 
-closszero <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
-	filter(Scenario == "RCP85High") %>%
-	mutate(CFE_Rate = 0) %>%
-	select(CFE_Rate, Timestep:Amount)
-
-## Summarize C stock data and create TEC by CFE figure ## 
-
-# combine CFE Cstock data together and summarize as Statewide TEC 
-cfetec <- bind_rows(Cstocks, CFE_Cstocks) %>%
-	filter(CFE_rate %in% c(0, 5, 10, 15)) %>%
-	group_by(CFE_rate, Timestep, SecondaryStratumID, StockGroupID) %>%
-	summarize(CstockTg = mean(Amount)/1000, 
-		Cstock_min = min(Amount)/1000, 
-		Cstock_max = max(Amount)/1000) %>%
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_CarbonBalance.R

@@ -1,233 +1,0 @@
-## Processing HI LUCAS model output:
-## summarizing Flows to estimate annual
-## and cumulative net biome productivity (NBP)
-## Paul C. Selmants
-## 2020-05-20
-
-# load packages
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(cowplot)
-
-
-# read HI LUCAS baseline C flow data into R
-baseflow<- read.csv("BaseFlows.csv", stringsAsFactors = FALSE)
-# read HI LUCAS Transition/Harvest C Loss data into R
-transflow <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE)
-
-# calculate statewide NPP
-NPP <- baseflow %>%
-	filter(FlowGroupID == "Growth [Type]") %>%
-	group_by(Scenario, Timestep, Iteration) %>%
-	summarize(NPP = sum(Amount)) %>%
-	as.data.frame()	
-# calculate statewide baseline Rh 
-Rh <- baseflow %>%
-	filter(FlowGroupID %in% c("Emission (litter) [Type]",
-		"Emission (soil) [Type]")) %>% 
-	group_by(Scenario, Timestep, Iteration) %>%
-	summarize(Rh = sum(Amount)) %>%
-	as.data.frame()
-# calculate statewide C loss by leaching/erosion
-Leach <- baseflow %>%
-	filter(FlowGroupID == "Leaching [Type]") %>%
-	group_by(Scenario, Timestep, Iteration) %>%
-	summarize(Leach = sum(Amount)) %>%
-	as.data.frame()
-# calculate disturbance C losses (∆LULC, fire, ag harvest)
-LUCloss <- transflow %>%
-	group_by(Scenario, Timestep, Iteration) %>%
-	summarize(LUCloss = sum(Amount)) %>%
-	as.data.frame()
-
-# join NPP and C loss dataframes together
-cflowjoin <- left_join(NPP, Rh) %>%
-	left_join(., Leach) %>%
-	left_join(., LUCloss) 
-
-# convert c flow data to long format, separate Scenario column into 
-# RCP and LandUse, summarize flow data by mean, min, and max in Tg 
-cflows <- cflowjoin %>% 
-	pivot_longer(cols = NPP:LUCloss, 
-		names_to = "flowtype",
-		values_to = "Amount") %>%
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_CarbonStocks.R

@@ -1,113 +1,0 @@
-## HI LUCAS model output:
-## summarize & graph Statewide 
-## Carbon stocks, fluxes, and balance
-## Paul C. Selmants
-## 2020-05-20
-
-# load required R packages
-library(dplyr)
-library(ggplot2)
-library(cowplot)
-
-# read Hawaii LUCAS model C stock output data into R
-Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE) %>%
-	group_by(Scenario, Timestep, SecondaryStratumID, StockGroupID) %>%
-	summarize(CstockTg = mean(Amount)/1000,
-		Cstock_min = min(Amount)/1000,
-		Cstock_max = max(Amount)/1000) %>%
-	data.frame()
-# remove " [Type]" at trailing end of all stock types  
-Cstocks$StockGroupID <- gsub('.{7}$', '', Cstocks$StockGroupID) 
-
-# create function to summarize carbon stocks statewide
-stocksum <- function(x) {
-	 x %>%
-	group_by(Scenario, Timestep) %>%
-	summarize(C_Tg = sum(CstockTg), 
-		Cmin = sum(Cstock_min),
-		Cmax = sum(Cstock_max)) %>%
-	select(Scenario, Timestep, C_Tg, Cmin, Cmax) %>%
-	arrange(Scenario, Timestep) %>%
-	as.data.frame() 
-}
-
-# summarize TEC and individual C stocks statewide
-TEC <- stocksum(Cstocks)
-# sum of standing and down dead wood
-Deadwood <- Cstocks %>%
-	filter(grepl("Deadwood", StockGroupID)) %>%
-	stocksum(.)
-# Living biomass
-LB <- Cstocks %>%
-	filter(StockGroupID == "Living Biomass") %>%
-	stocksum(.)
-# Litter
-Litter <- Cstocks %>%
-	filter(StockGroupID == "Litter") %>%
-	stocksum(.)
-# Soil organic carbon
-SOC <- Cstocks %>%
-	filter(StockGroupID == "Soil") %>%
-	stocksum(.)
-
-# merge individual C stocks into single dataframe
-indCstocks <- bind_rows(Deadwood, LB, Litter, SOC) %>%
-	mutate(StockGroupID = rep(c("Dead wood", "Live biomass", "Litter",
-		"Soil organic matter"), each = 364)) %>%
-	arrange(Scenario, Timestep) %>%
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_ClimateFigs.R

@@ -1,159 +1,0 @@
-# Use ggplot to visualize
-# current and end-of-century
-# rainfall and air temperature 
-# in the Hawaiian Islands 
-# Paul C. Selmants
-# 2020-02-24
-
-# load required R packages
-library(raster)
-library(rgdal)
-library(ggplot2)
-library(dplyr)
-library(viridis)
-library(RColorBrewer)
-library(cowplot) 
-
-# create list of annual rainfall .tif files
-rainlist <- list.files('rainfall_tiff_annual', full.names = TRUE)
-#create raster stack of annual rainfall (mm) .tif files
-rainstack <- stack(rainlist)
-# make raster layer of 30-year mean annual rainfall and re-sample
-# to change projection to UTM (this takes ~30 seconds)
-rain30ymean <- calc(rainstack, mean)
-# read in rasters of late century rainfall anomolies for RCPs 4.5 & 8.5
-rcp85anom <- raster("sd_rfanom_rcp85_late.tif")
-rcp45anom <- raster("sd_rfanom_rcp45_late.tif")
-# make rasters of late century rainfall differences from present
-raindiff85 <- rain30ymean*(rcp85anom/100)
-raindiff45 <- rain30ymean*(rcp45anom/100)
-
-# read in current mean annual temperature raster
-MAT <- raster("MAT_annual.tif") 
-# read in rasters of late century temperature differences for RCPs 4.5 & 8.5
-Tdelta_rcp45 <- raster("sd_Tdeltas_rcp45_late.tif")
-Tdelta_rcp85 <- raster("sd_Tdeltas_rcp85_late.tif")
-
-#prepare rain30ymean and raindiff data to plot with geom_tile in ggplot
-rain_df <- rasterToPoints(rain30ymean) %>%
-	data.frame(.)
-
-r45_df <- rasterToPoints(raindiff45) %>%
-	data.frame(.)
-
-r85_df <- rasterToPoints(raindiff85) %>%
-	data.frame(.)
-#prepare MAT and dT data to plot with geom_tile in ggplot2 
-MAT_df <- rasterToPoints(MAT) %>%
-	data.frame(.) %>%
-	select(x, y, layer = MAT_annual)
-
-t45_df <- rasterToPoints(Tdelta_rcp45) %>%
-	data.frame(.) %>%
-	select(x, y, layer = sd_Tdeltas_rcp45_late)
-
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_LandCoverFig.R

@@ -1,80 +1,0 @@
-# Using ggplot2 to visualize
-# Hawaiʻi land cover classes
-# Paul C. Selmants
-# 2020-02-24
-
-# load required packages
-library(raster)
-library(rgdal)
-library(dplyr)
-library(ggplot2)
-
-# set working directory to HI_Model GitHub repository
-
-# read State Class TIFF into R, where: 
-# 1 = Water 		6 = Forest
-# 2 = Urban 		7 = Grassland
-# 3 = Plantation	8 = Agriculture
-# 4 = WoodyCrop		9 = Wetland
-# 5 = Barren	   10 = Shrubland
-StateClass <- raster('./InputData/spatial_data/StateClassNew_250m.tif')
-
-# read annual MAT Tiff into R
-MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
-
-# make reclassification matrix for simplified land cover raster
-lc_rcl <- c(1, NA, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6,
-	7, 7, 8, 4, 9, NA, 10, 10) %>%
-	matrix(., ncol = 2, byrow = TRUE)
-# reclassify StateClass to create simplified land cover raster, where: 
-# 2 = Urban 		 6 = Forest
-# 3 = Plantation	 7 = Grassland
-# 4 = Agriculture	10 = Shrubland
-# 5 = Barren	
-lc <- reclassify(StateClass, lc_rcl)
-# re-sample lc projection to Lat. Long. using MAT
-lcdeg <- projectRaster(lc, MAT, method = 'ngb')
-# convert lc raster to points dataframe for usage in ggplot 
-lc_df <- rasterToPoints(lcdeg) %>%
-	data.frame(.) %>%
-	select(x, y, layer = StateClassNew_250m) 
-# convert layer column to factor 
-lc_df$layer <- as.factor(lc_df$layer)
-# Create custom colorblind palette for land cover classes
-lcpal <- c("#000000", "#CC79A7", "#A6761D", "#969696", "#006D2C", 
-	"#FFD92F", "#B2DF8A") 
-# create statewide land cover map for Hawaii
-lcmap <- ggplot(data=lc_df) +
-	geom_tile(aes(x=x, y=y, fill=layer)) +
-	xlab(expression("Longitude " ( degree*W))) +
-	ylab(expression("Latitude " ( degree*N))) +
-	scale_fill_manual("Land Cover", values = lcpal,
-		 labels=c("Developed", "Tree Plantation", "Agriculture",
[…]

### 2020-07-01 Output/manuscript/fig_code/HI_StateClassArea.R

@@ -1,57 +1,0 @@
-## Processing HI LUCAS model output:
-## summarizing area of different  
-## State Classes over time
-## by land use scenario
-## Paul C. Selmants
-## 2020-04-22
-
-# load required R packages
-library(dplyr)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-library(tidyr) 
-
-# read State Class model output data into R and summarize 
-# statewide by land use scenario
-sc <- read.csv("StateClassArea.csv", stringsAsFactors = FALSE) %>%
-	mutate(StateClassID = str_remove(.$StateClassID, ":All")) %>%
-	group_by(Scenario, Timestep, Iteration, StateClassID) %>%
-	summarize(area = sum(Area)) %>%
-	separate(Scenario, c("Climate", "LandUse"), sep = 5) %>%
-	group_by(LandUse, Timestep, StateClassID) %>%
-	summarize(area_mean = mean(area), 
-		area_min = min(area), 
-		area_max = max(area)) %>%
-	data.frame() 
-
-#Make Land Use color palette
-lu_pal <- c("#A6761D","#1B9E77")
-
-# ggplot faceted land cover change graph
-scfig <- sc %>%
-filter(StateClassID %in% c("Agriculture",
-	"Developed", "Forest", "Grassland", "Shrubland")) %>% 
-	{ggplot(., aes(x = Timestep, y = area_mean, color = LandUse)) +
-	facet_wrap(~ StateClassID, scales = "free_y") +
-	geom_ribbon(aes(ymin = area_min, ymax = area_max, fill = LandUse),
-		alpha = 0.4, color = NA) +
-	geom_line(size = 0.5) +
-	scale_fill_manual(values = lu_pal, labels=c("High", "Low"),
-			name = "Land use\nscenario") +
-	scale_colour_manual(values = lu_pal, labels=c("High", "Low"), 
-			name = "Land use\nscenario") +
-	scale_x_continuous(limits = c(2000,2100), breaks = seq(2000,2100,20)) +
-	scale_y_continuous(breaks = scales::pretty_breaks(5), limits = c(NA, NA)) +
-	labs(x = "Year", 
-		 y = expression(Land~area~(km^{2})), 
-		color = "Land Use\nScenario",
-		fill = "Land Use\nScenario") +
-	theme_bw() +
-	theme(panel.grid.minor = element_blank(), 
-		legend.position = c(0.85,0.25),
[…]

### 2020-07-01 Output/manuscript/fig_code/MZfig.R

@@ -1,56 +1,0 @@
-# Use ggplot to visualize
-# Moisture Zones in the 
-# Hawaiian Islands
-# Paul C. Selmants
-# 2020-06-23
-
-# load required packages
-library(raster)
-library(rgdal)
-library(dplyr)
-library(ggplot2)
-
-# set working directory to HI_Model GitHub repository
-
-# read Moisture Zone TIFF into R, where: 
-# 1 = Dry 		
-# 2 = Mesic 		
-# 3 = Wet	
-MZ <- raster('./InputData/spatial_data/MZ_3_250m.tif')
-
-# read annual MAT Tiff into R
-MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
-
-# re-sample MZ projection to Lat. Long. using MAT
-MZdeg <- projectRaster(MZ, MAT, method = 'ngb')
-# convert lc raster to points dataframe for usage in ggplot 
-MZ_df <- rasterToPoints(MZdeg) %>%
-	data.frame(.) %>%
-	select(x, y, layer = MZ_3_250m) 
-# convert layer column to factor 
-MZ_df$layer <- as.factor(MZ_df$layer)
-# Create custom palette for MZ classes
-mzpal <- c("#FDBF6F", "#B2DF8A", "#1F78B4")
-
-# create statewide Moisture Zone map for Hawaii
-mzmap <- ggplot(data=MZ_df) +
-	geom_tile(aes(x=x, y=y, fill=layer)) +
-	xlab(expression("Longitude " ( degree*W))) +
-	ylab(expression("Latitude " ( degree*N))) +
-	scale_fill_manual("Moisture Zone", values = mzpal,
-		 labels=c("Dry", "Mesic", "Wet")) +
-	theme_bw() + 
-	theme(axis.title.x = element_text(size=10),
-        axis.text.x  = element_text(size=8),
-        axis.title.y = element_text(size=10),
-        axis.text.y = element_text(size=8),
-    legend.position = c(0.188, 0.465),
-    legend.justification = c("left", "top"),
-    legend.title = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.background = element_rect(fill="white", 
-    	linetype = "solid", size = 0.5, color = "white"))
-
-# save Moisture Zone map as .png file 
-ggsave("./Output/manuscript/fig_images/figS1_MZ.png", 
-	height = 5, width = 7, dpi = 300)

### 2020-07-03 Manuscript/fig_code/HI_CFE_CarbonBalance.R
updated CFE figure

@@ -12,27 +12,27 @@ library(cowplot)
 ## https://doi.org/10.5066/P9AWLFKZ
 
 # read LUCAS CFE C stocks from RCP85High output data into R
-CFE_Cstocks <- read.csv("Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
+CFE_Cstocks <- read.csv("./HI_Cstocks/Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in LUCAS C stocks by scenario and filter on RCP85High scenario 
-Cstocks <- read.csv("Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
+Cstocks <- read.csv("./HI_Cstocks/Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
 	filter(Scenario == "RCP85High") %>%
 	mutate(CFE_rate = 0) %>%
 	select(CFE_rate, Timestep:Amount)
 
 # read in CFE base flow data from RCP85High Scenario
-cfe_baseflow <- read.csv("BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+cfe_baseflow <- read.csv("./HI_Cflows/BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in CFE C loss data (transition, fire, harvest)
-cfe_closs <- read.csv("TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+cfe_closs <- read.csv("./HI_Cflows/TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in zero CFE base flow data 
-baseflowzero <- read.csv("BaseFlows.csv", stringsAsFactors = FALSE) %>%
+baseflowzero <- read.csv("./HI_Cflows/BaseFlows.csv", stringsAsFactors = FALSE) %>%
 	filter(Scenario == "RCP85High") %>%
 	mutate(CFE_Rate = 0) %>%
 	select(CFE_Rate, Timestep:Amount)
 # read zero CFE non-Rh C Loss data 
-closszero <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
+closszero <- read.csv("./HI_Cflows/TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
[…]

### 2020-07-20 Manuscript/HI_Model_ms.Rmd
updated manuscript and BibTex files

--- 2020-06-18 Output/manuscript/HI_Model_ms.Rmd
@@
 ## Methods
+We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under four scenarios incorporating different levels of climate and land-use change.
 
@@
 This land area was subdivided into a grid of 250 x 250 m simulation cells.
-Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three Moisture Zones (Figure S1), seven Islands, and ten discrete land cover classes (Figure 1).
+Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; Figure S1), seven islands, and ten discrete land cover classes (Figure 1).
 
-```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover map of the seven main Hawaiian Islands, adapted from Jacobi et al.
+```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al.
 (2017).

### 2020-07-22 Manuscript/HI_Model_ms.Rmd
modified bibtex, .csl, and manuscript files

@@
 bibliography: HI_Model.bib
-csl: environmental-research-letters.csl
+csl: iop.csl
 fontsize: 12pt
@@
     latex_engine: xelatex
+urlcolor: blue
 ---
@@
 ## Methods
-We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under four scenarios incorporating different levels of climate and land-use change.
+We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under four scenarios of climate and land-use change.
+The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
+The carbon gain-loss portion of LUCAS tracks carbon stocks within each simulation cell over time as continuous state variables, as well as a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018].
+The Hawai`i LUCAS model estimated the annual change in carbon stocks and fluxes in response to land use changes associated with urbanization, agricultural expansion and contraction, forest and agricultural harvesting, and wildland fire.
+We also estimated the impact of short- and long-term climate variability on net primary productivity and the turnover of dead organic matter [@sleeter_effects_2019].
 
@@
 
-```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al.
-(2017).
+```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
 Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
@@
 Government.
+
+## Data Availability
+
+Tabular model output data are available from the USGS ScienceBase data repository at: https://doi.org/10.5066/P9AWLFKZ.
+Model input data and R code used to generate input data, summarize output data, and compile this manuscript are available from the HI_Model GitHub repository at: https://github.com/selmants/HI_Model.
+
+## ORCID
+
+Paul C. Selmants https://orcid.org/0000-0001-6211-3957
+
+Benjamin M. Sleeter https://orcid.org/0000-0003-2371-9571
+
+Jinxun Liu https://orcid.org/0000-0003-0561-8988
+
+Tamara S. Wilson https://orcid.org/0000-0001-7399-7532
+
+Abby G. Frazier https://orcid.org/0000-0003-4076-4577
+
+Gregory P. Asner https://orcid.org/0000-0001-7893-6421
+
+## References

### 2020-07-23 Manuscript/HI_Model_ms.Rmd
edited Rmd manuscript

@@
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
-The carbon gain-loss portion of LUCAS tracks carbon stocks within each simulation cell over time as continuous state variables, as well as a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018].
-The Hawai`i LUCAS model estimated the annual change in carbon stocks and fluxes in response to land use changes associated with urbanization, agricultural expansion and contraction, forest and agricultural harvesting, and wildland fire.
-We also estimated the impact of short- and long-term climate variability on net primary productivity and the turnover of dead organic matter [@sleeter_effects_2019].
+The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018].
+We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability [@sleeter_effects_2019].
 
@@
 
-The study area encompassed the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^ (Figure 1).
-This land area was subdivided into a grid of 250 x 250 m simulation cells.
+We subdivided the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^, into a grid of 250 x 250 m simulation cells.
 Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; Figure S1), seven islands, and ten discrete land cover classes (Figure 1).

### 2020-07-28 Manuscript/HI_Model_ms.Rmd
updated manuscript file and eited comments in transition targets script

@@
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
-The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018].
-We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability [@sleeter_effects_2019].
+The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019].
+We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability.
+Simulations were run for 90 years at an annual timestep, using initial conditions corresponding to the year 2010.
+All simulations were repeated for 30 Monte Carlo realizations.
 
 ### *Study area*
-
-We subdivided the terrestrial portion of the seven main Hawaiian Islands, a total land area of 16,464 km^2^, into a grid of 250 x 250 m simulation cells.
+The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (Figure 1), a total land area of 16,554 km^2^.
+We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size.
 Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; Figure S1), seven islands, and ten discrete land cover classes (Figure 1).
@@
 ### *States and transitions*
+Transition pathways were based on a model described in @daniel_state-and-transition_2016 and @selmants_baseline_2017.
+Specifically, transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
+Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban [@daniel_state-and-transition_2016].
+Although several state types could transition to urban, there was no transition pathway out of the urban state type.
+Water and wetland state types remained static throughout the simulation period.
+We developed a high and a low land-use scenario, with transition targets for agricultural contraction, agricultural expansion, and urbanization based on historical trends and population projections.
+
+For the high land-use scenario,
 
@@
 Geological Survey.
+Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.

### 2020-07-28 Transitions/code/HI_Transition_Targets.R

@@ -92,7 +92,7 @@ Low_Urban <- bind_rows(UrbanHist, UrbanLow_early, UrbanLow_mid, UrbanLow_late) %
 	filter(SecondaryStratumID != "Kaho'olawe")
 
 # High urbanization transition targets by island for 2010-2045
-# based on mid to max range of historical annual rates (1992-2010 NOAA CCAP)  
+# based on mid- to max range of historical annual rates (1992-2010 NOAA CCAP)  
 UrbanHigh_early <- lulc %>%
 	filter(LC_change == "Urbanization") %>%
 	mutate(Iteration = "",
@@ -107,7 +107,7 @@ UrbanHigh_early <- lulc %>%
 		Amount, DistributionType, DistributionFrequencyID, DistributionSD, 
 		DistributionMin = mid, DistributionMax = maximum)
 # High urbanization transition targets for 2045-2100
-# based on mid-max range of historical annual rates (1992-2010) 
+# based on minimum to maximum range of historical annual rates (1992-2010) 
 UrbanHigh_mid <- lulc %>%
 	filter(LC_change == "Urbanization") %>%
 	mutate(Iteration = "",

### 2020-07-29 Manuscript/HI_Model_ms.Rmd
updated manuscript files and edited transition targets R script

@@
 urlcolor: blue
+nocite: |
+  @jacobi_map_2017
 ---
@@
 ## Abstract
-The State of Hawai`i recently passed legislation setting a goal to be carbon neutral by 2045.
+The State of Hawai`i passed legislation in 2018 setting a goal to be carbon neutral by 2045.
 Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain.
@@
 ### *States and transitions*
-Transition pathways were based on a model described in @daniel_state-and-transition_2016 and @selmants_baseline_2017.
-Specifically, transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban [@daniel_state-and-transition_2016].
-Although several state types could transition to urban, there was no transition pathway out of the urban state type.
+Transition pathways were modified from @daniel_state-and-transition_2016 and @selmants_baseline_2017.
+Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
+Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban.
+Although several state types could transition to urban, there was no transition pathway from urban to any other state type.
 Water and wetland state types remained static throughout the simulation period.
-We developed a high and a low land-use scenario, with transition targets for agricultural contraction, agricultural expansion, and urbanization based on historical trends and population projections.
+We developed a high and a low land-use scenario, with transition targets based on historical trends of land use change and population projections [@kim_population_2018].
 
-For the high land-use scenario,
+For the high land-use scenario, annual transition targets for agricultural contraction, agricultural expansion, and urbanization were based on
 
@@
 Geological Survey.
-Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
+Thanks to Nicholas Koch of Forest Solutions, Inc.
+for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with the SyncroSim software.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
@@
 Tabular model output data are available from the USGS ScienceBase data repository at: https://doi.org/10.5066/P9AWLFKZ.
-Model input data and R code used to generate input data, summarize output data, and compile this manuscript are available from the HI_Model GitHub repository at: https://github.com/selmants/HI_Model.
+Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from the HI_Model GitHub repository at: https://github.com/selmants/HI_Model.
 

### 2020-07-29 Transitions/code/HI_Transition_Targets.R

@@ -50,7 +50,7 @@ UrbanLow_early <- ddev %>%
 		Amount, DistributionType, DistributionFrequencyID, 
 		DistributionSD, DistributionMin = min, DistributionMax = ddev_km2y) 
 # Low urbanization transition targes by island for 2045-2070
-# using zero to 25% of minimum annual rate population projection per island
+# using zero to 25% of minimum annual population projection per island
 UrbanLow_mid <- ddev %>%
 	group_by(island) %>%
 	summarise(DistributionMax = min(ddev_km2y)*0.25) %>%
@@ -69,7 +69,7 @@ UrbanLow_mid <- ddev %>%
 		DistributionSD, DistributionMin, DistributionMax) %>%
 	data.frame(.)
 # Low urbanization transition targes by island for 2070-2100 using
-# zero to 0.10 x minimum annual rate population projection per island
+# zero to 0.10 x minimum annual population projection per island
 UrbanLow_late <- ddev %>%
 	group_by(island) %>%
 	summarise(DistributionMax = min(ddev_km2y)*0.10) %>%

### 2020-07-31 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ### *States and transitions*
-Transition pathways were modified from @daniel_state-and-transition_2016 and @selmants_baseline_2017.
+We developed a high and a low land-use scenario, with transition pathways modified from @daniel_state-and-transition_2016.
 Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
 Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban.
-Although several state types could transition to urban, there was no transition pathway from urban to any other state type.
+Although several state types had urbanization transition pathways, there was no transition pathway from urban to any other state type.
 Water and wetland state types remained static throughout the simulation period.
-We developed a high and a low land-use scenario, with transition targets based on historical trends of land use change and population projections [@kim_population_2018].
 
-For the high land-use scenario, annual transition targets for agricultural contraction, agricultural expansion, and urbanization were based on
+ Transition targets were based on historical trends of land use change and population projections.
+For the high land-use scenario, rates of agricultural contraction, agricultural expansion, and urbanization were derived from remote sensing-based time series maps of land use change in the Hawaiian Islands from 1992-2011 developed by the National Oceanic and Atmospheric Administration Office of Coastal Planning [@noaa_coastal_2020].
+We sampled transition rates for each timestep and Monte Carlo realization from uniform distributions bounded by the median and maximum historical values of agricultural contraction, agricultural expansion, and urbanization for each island.
+For the low land-use scenario, rates of agricultural contraction and expansion were derived from the same NOAA time series by sampling from uniform distributions bounded by zero and the minimum historical values for each island.
+Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
+We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population.
+The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
+
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 using a new wildfire spatial database for the Hawaiian Islands from 1999-2019 (Figure S4).
+Wildfire size distribution and ignition probabilities for each year were calculated from this 21-year time series for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1).
+The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization.
+Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide, and the high land-use scenario sampled from historical fire years above this 21-year median (Fig.
+S4a).
+Burn severity probabilities (high, medium, low) for each combination of moisture zone and state type were taken from @selmants_baseline_2017.
+Wildfire transitions were restricted to previously burned areas using a map of all wildfires from 1999-2018.
 

### 2020-07-31 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Wildfire size distribution and ignition probabilities for each year were calculated from this 21-year time series for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1).
-The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization.
-Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide, and the high land-use scenario sampled from historical fire years above this 21-year median (Fig.
+The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
+Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide and the high land-use scenario sampled from historical fire years above this 21-year median (Fig.
 S4a).
-Burn severity probabilities (high, medium, low) for each combination of moisture zone and state type were taken from @selmants_baseline_2017.
-Wildfire transitions were restricted to previously burned areas using a map of all wildfires from 1999-2018.
+
+Tree plantation harvesting
 

### 2020-08-03 Manuscript/HI_Model_ms.Rmd
updated manuscript files and retitled supplementary figures

--- 2020-07-29 Manuscript/HI_Model_ms.Rmd
@@
 ### *States and transitions*
-Transition pathways were modified from @daniel_state-and-transition_2016 and @selmants_baseline_2017.
+We developed a high and a low land-use scenario, with transition pathways modified from @daniel_state-and-transition_2016.
 Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban.
-Although several state types could transition to urban, there was no transition pathway from urban to any other state type.
+Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed.
+Although several state types had urbanization transition pathways, but there was no transition pathway from developed to any other state type.
 Water and wetland state types remained static throughout the simulation period.
-We developed a high and a low land-use scenario, with transition targets based on historical trends of land use change and population projections [@kim_population_2018].
 
-For the high land-use scenario, annual transition targets for agricultural contraction, agricultural expansion, and urbanization were based on
+ Transition targets were based on historical trends of land use change from 1992-2011 [@noaa_coastal_2020] and from population projections [@kim_population_2018].
+For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
+For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
+Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
+We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population.
+The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
+Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016].
+In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion of *Eucalyptus* spp.
+plantations to agriculture.
+The remaining 40% were rotation harvests.
+
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016.
+Historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) were based on a new 21-year historical wildfire spatial database of the Hawaiian Islands covering the years 1999 to 2019 (Figure S2).
+The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
+Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide; the high land-use scenario sampled from historical fire years above this 21-year median (Fig.
+S2a).
 

### 2020-08-03 Manuscript/fig_code/AnnualFire_plot.R

@@ -55,4 +55,4 @@ firepanel <- plot_grid(stateplot, islandplot,
 	labels = c("A", "B"), rel_widths = c(1.15, 1))
 
 # save as .png file 
-ggsave("FigS4_AnnualAreaBurned.png", width = 7.5, height = 3)
+ggsave("FigS2_AnnualAreaBurned.png", width = 7.5, height = 3)

### 2020-08-04 FlowMultipliers/code/NPPmultipliers.R
modified NPPmultipliers code and output

@@ -7,11 +7,13 @@
 ## Paul C. Selmants
 ## 2020-03-19 (ISO 8601)
 
+## R version 3.6.1
+
 # load required packages into R
-library(raster)
-library(rgdal)
-library(tidyr, warn.conflicts = FALSE)
-library(dplyr, warn.conflicts = FALSE)
+library(raster) #version 3.1-5
+library(rgdal) #version 1.4-8
+library(tidyr, warn.conflicts = FALSE) #version 1.0.2
+library(dplyr, warn.conflicts = FALSE) #version 0.8.5
 
 ## Set working directory at highest level in HI_Model
 ## GitHub repository, https://github.com/selmants/HI_Model 
@@ -22,7 +24,7 @@ library(dplyr, warn.conflicts = FALSE)
 # 3 = Plantation 	 8 = Agriculture			 
 # 4 = WoodyCrop 	 9 = Wetland		 
 # 5 = Barren 		10 = Shrubland		
-SC <- raster("./InputData/spatial_data/StateClassNew_250m.tif")
+SC <- raster("./InputData/spatial_data/initial_conditions/StateClassNew_250m.tif")
 
 # reclassification matrix for SC to set Water, Urban, 
 # Barren, & Wetland to 1 and set other classes to NA

### 2020-08-05 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,
+
 ## Methods
-We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under four scenarios of climate and land-use change.
+We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
@@
 
- Transition targets were based on historical trends of land use change from 1992-2011 [@noaa_coastal_2020] and from population projections [@kim_population_2018].
+ Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018].
 For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
@@
 plantations to agriculture.
-The remaining 40% were rotation harvests.
+The remaining 40% were rotation harvests with replanting of *Eucalyptus* spp.
 
-The wildfire transition sub-model was modified from @daniel_state-and-transition_2016.
-Historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) were based on a new 21-year historical wildfire spatial database of the Hawaiian Islands covering the years 1999 to 2019 (Figure S2).
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2) to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1).
 The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
-Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide; the high land-use scenario sampled from historical fire years above this 21-year median (Fig.
+Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019, while the high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig.
 S2a).

### 2020-08-07 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ### *States and transitions*
-We developed a high and a low land-use scenario, with transition pathways modified from @daniel_state-and-transition_2016.
+We developed two land-use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
 Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed.
-Although several state types had urbanization transition pathways, but there was no transition pathway from developed to any other state type.
+Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
+There was no transition pathway out of an urbanized (developed) state.
 Water and wetland state types remained static throughout the simulation period.
@@
 Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016].
-In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion of *Eucalyptus* spp.
-plantations to agriculture.
-The remaining 40% were rotation harvests with replanting of *Eucalyptus* spp.
+In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
+The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
 
-The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2) to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1).
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2).
+We used this new Hawai`i fire database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019.
 The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
@@
 ### *Carbon stocks and flows*
+ The fate of continuous carbon stocks was tracked for each simulation cell based on a suite of continuous flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
 

### 2020-08-18 Manuscript/HI_Model_ms.Rmd
changes to manuscript files

@@
 The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
-Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019, while the high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig.
+Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
+The high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig.
 S2a).
@@
 ### *Carbon stocks and flows*
- The fate of continuous carbon stocks was tracked for each simulation cell based on a suite of continuous flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
+ The fate of carbon stocks was tracked for each simulation cell based on a suite of carbon flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
+We defined carbon stocks as continuous state variables for each simulation cell, including live biomass, standing dead wood, down dead wood, litter, and soil organic carbon.
+We also included and tracked carbon in atmospheric, aquatic, and harvest product pools to enforce carbon mass balance [@daniel_integrating_2018].
+To transfer carbon between stocks, we defined baseline carbon flows as continuous variables resulting from growth, mortality, deadfall, woody decay, litter decomposition, and leaching (which includes runoff).
+We also defined carbon flows resulting from land use, land use change, and wildfire [@selmants_baseline_2017; @daniel_integrating_2018].
+
+ Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (Figure S1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
+We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016].
+We also calibrated gross photosynthesis in IBIS using a Hawai`i-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
+A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate.
+This spatial growth multiplier was the NPP anomaly for each cell relative the mean value for each combination of moisture zone and state type [@sleeter_effects_2019] calculated using empirical relationships between total NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
+Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under RCPs 4.5 and 8.5 [@timm_statistical_2015; @timm_future_2017].
+For each RCP, temporal growth multipliers were calculated using empirical equations relating NPP to climate [@schuur_productivity_2003; @del_grosso_global_2008] and temporal decay multipliers were calculated using Q10 functions [@sleeter_effects_2019].
+Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.
 
@@
 Thanks to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with the SyncroSim software.
+for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.

### 2020-08-22 Manuscript/HI_Model_ms.Rmd
edited fire code and updated manuscript files

@@
 
-```{r fig.align = 'center', out.width = "80%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
 Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
@@
 The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2).
-We used this new Hawai`i fire database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019.
-The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
+We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019.
+Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
 Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
@@
 We also calibrated gross photosynthesis in IBIS using a Hawai`i-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
+
+ Carbon flow rates for each state type and moisture zone were estimated as the ratio of the IBIS-derived flux to the size of the originating carbon stock at each age [@sleeter_effects_2018].
 A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate.
-This spatial growth multiplier was the NPP anomaly for each cell relative the mean value for each combination of moisture zone and state type [@sleeter_effects_2019] calculated using empirical relationships between total NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
-Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under RCPs 4.5 and 8.5 [@timm_statistical_2015; @timm_future_2017].
-For each RCP, temporal growth multipliers were calculated using empirical equations relating NPP to climate [@schuur_productivity_2003; @del_grosso_global_2008] and temporal decay multipliers were calculated using Q10 functions [@sleeter_effects_2019].
+This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
+Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of the two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
+The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
+The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario.
+We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014].
 Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.
 
-### *Initial conditions*
+### *CO~2~ fertilization effect*
+
+Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well.
+However, Following @sleeter_effects_2019, we developed a set of separate scenarios designed to test the influence of a CFE on ecosystem carbon balance.
+We tested five $\beta$ levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
 
@@
 
-Tabular model output data are available from the USGS ScienceBase data repository at: https://doi.org/10.5066/P9AWLFKZ.
-Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from the HI_Model GitHub repository at: https://github.com/selmants/HI_Model.
+Tabular model output data are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ.
+Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.
 

### 2020-08-22 Fire/code/Fire_probabilities.R

@@ -25,19 +25,19 @@ library(stringr)
 # 2 = Kahoolawe		6 = Oahu
 # 3 = Lanai			7 = Kauai
 # 4 = Maui
-Islands <- raster("./InputData/spatial_data/initial_conditions/Island_250m.tif")
+Islands <- raster("../InputData/spatial_data/initial_conditions/Island_250m.tif")
 # read Moisture Zone GeoTIFF into R, where:
 # 1 = "Dry"
 # 2 = "Mesic"
 # 3 = "Wet"
-MZ <- raster("./InputData/spatial_data/initial_conditions/MZ_3_250m.tif")
+MZ <- raster("../InputData/spatial_data/initial_conditions/MZ_3_250m.tif")
 # read State Class GeoTIFF into R, where: 
 # 1 = Water 		6 = Forest
 # 2 = Urban 		7 = Grassland
 # 3 = Plantation	8 = Agriculture
 # 4 = WoodyCrop		9 = Wetland
 # 5 = Barren	   10 = Shrubland
-StateClass <- raster("./InputData/spatial_data/intital_conditions/StateClassNew_250m.tif")
+StateClass <- raster("../InputData/spatial_data/initial_conditions/StateClassNew_250m.tif")
 # read in Hawaii statewide annual fire perimeter shapefile (1999-2019) 
 fires <- st_read("./data/base/2019_1999_Hawaii_Fire_Perimeters.shp") %>%
 	dplyr::select(UH_ID, Year, geometry)
@@ -80,7 +80,7 @@ Maui <- reclassify(Islands, Maui_recl)
 Molokai <- reclassify(Islands, Molokai_recl)
 Oahu <- reclassify(Islands, Oahu_recl)
 Kauai <- reclassify(Islands, Kauai_recl)
-#make raster stack of invidual island rasters
+# make raster stack of invidual island rasters
 islandstack <- stack(Hawaii, Kahoolawe, Lanai, Maui, Molokai, Oahu, Kauai)
 # make vector of island names
 islandnames <- c("Hawaii", "Kahoolawe", "Lanai", "Maui", "Molokai", "Oahu", "Kauai")
@@ -149,7 +149,7 @@ mzsc_area <- zonal(mzsc_island, MZSC, fun = 'count') %>%
 		StratumID = rep(c(rep(c("Dry", "Mesic", "Wet"), each = 6)), 7), 
[…]

### 2020-08-27 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
- Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018].
+ Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawai`i [@kim_population_2018].
 For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
@@
 Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well.
-However, Following @sleeter_effects_2019, we developed a set of separate scenarios designed to test the influence of a CFE on ecosystem carbon balance.
-We tested five $\beta$ levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
+However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans.
+Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE).
+We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario.
+We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
+For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
+This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
 
-### *Scenario simulations*
+### *Scenario simulations and analysis*
 
@@
 for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
+Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
@@
 
-Tabular model output data are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ.
+Tabular model output data in machine readable format are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ.
 Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.

### 2020-09-03 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ## Introduction
-
 The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,
@@
 We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
+We also developed a separate set of scenarios to test model sensitivity to different levels of a CO~2~ fertilization effect (CFE).
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
 The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019].
-We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability.
-Simulations were run for 90 years at an annual timestep, using initial conditions corresponding to the year 2010.
-All simulations were repeated for 30 Monte Carlo realizations.
+We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
 
@@
 ### *CO~2~ fertilization effect*
-
 Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well.
@@
 ### *Scenario simulations and analysis*
+Each of the four unique scenarios were run for 90 years at an annual timestep and repeated for 30 Monte Carlo realizations, using initial conditions corresponding to the year 2010.
+All simulations were performed within the SyncroSim (version 2.2.4) software framework with ST-Sim (version 3.2.13) and SF (version 3.2.10) add-on modules [@daniel_state-and-transition_2016; @daniel_integrating_2018].
+Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
 
 ## Results
+ Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; Fig.
+2a).
+The combination of low radiative forcing and low land use change yielded the highest and most consistent projections of ecosystem carbon accumulation, with a gain of ~15% to an average of 363 Tg of carbon by 2100 (Fig.
+2a).
+
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands.
+Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
+knitr::include_graphics("./fig_images/fig2_Cstocks.png")
+```
 

### 2020-09-08 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
 Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
@@
 2a).
-The combination of low radiative forcing and low land use change yielded the highest and most consistent projections of ecosystem carbon accumulation, with a gain of ~15% to an average of 363 Tg of carbon by 2100 (Fig.
+Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land-use change.
+The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (Fig.
 2a).
+In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 but declining to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (Fig.
+2a).
+Ecosystem carbon accumulation was driven exclusively by increasing soil organic carbon across all four scenarios, all other stocks declined over time (Fig.
+2b).
 
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands.
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands.
 Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}

### 2020-09-10 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
+Net primary production (NPP) declined across all four scenarios (Fig.
+3).
+The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (Fig.
+S4) and reduction of forest and shrubland land area in the high land use scenario (Fig.
+S5).
+In contrast to NPP, heterotrophic respiration (Rh) was higher under RCP 8.5 than under RCP 4.5 (Fig.
+3) because of more intense warming over time (Fig.
+S4).
+High land-use change lowered Rh (Fig.
+3) because of reduced forest and shrubland land area (Fig.
+S5), similar to trends in NPP.
+Carbon fluxes to the atmosphere triggered by land use, land-use change, and fire were largely independent of changes in climate, with greater variability in these fluxes across Monte Carlo iterations in the high land-use scenario (Fig.
+3).
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net primary production (NPP), heterotrophic respiration (Rh) and carbon fluxes induced by land use and land-use change for the seven main Hawaiian Islands.
+Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
+knitr::include_graphics("./fig_images/fig3_Cflows.png")
+```
+
 ## Discussion

### 2020-09-14 Manuscript/HI_Model_ms.Rmd
updated manuscript files and added supplementary document file

--- 2020-09-08 Manuscript/HI_Model_ms.Rmd
@@
 ### *Study area*
-The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (Figure 1), a total land area of 16,554 km^2^.
+The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (figure 1), a total land area of 16,554 km^2^.
 We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size.
-Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; Figure S1), seven islands, and ten discrete land cover classes (Figure 1).
+Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; figure S1), seven islands, and ten discrete land cover classes (figure 1).
 
@@
 
-The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2).
-We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019.
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (figure S2).
+We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (figure S1), island, and state type (figure 1) for the years 1999-2019.
 Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
@@
 
- Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (Figure S1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+ Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (figure S1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
@@
 Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
+Carbon stocks and fluxes for the seven main Hawaiian Islands were calculated for each scenario by summing within each of the 30 Monte Carlo realizations on an annual basis and then calculating annual means as well as the annual upper and lower limits of the 30 Monte Carlo realizations.
+Carbon balance for the seven main Hawaiian Islands was calculated on annual basis for each scenario and Monte Carlo realization as net biome productivity (NBP), which was equal to annual carbon input in the form of net primary productivity (NPP) minus the annual sum of all carbon losses from terrestrial ecosystems, including heterotrophic respiration (R~h~) from litter and soil, carbon fluxes to the atmosphere triggered by land use and land-use change, wildfire emissions, and aquatic carbon losses through leaching and overland flow.
+Positive NBP values indicated ecosystems of the seven main Hawaiian Islands were acting as a net sink for atmospheric CO~2~, while negative NBP values indicated that these ecosystems were acting as a net carbon source to the atmosphere [@chapin_reconciling_2006].
 
 ## Results
- Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; Fig.
-2a).
+ Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2a).
 Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land-use change.
-The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (Fig.
-2a).
-In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 but declining to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (Fig.
-2a).
-Ecosystem carbon accumulation was driven exclusively by increasing soil organic carbon across all four scenarios, all other stocks declined over time (Fig.
-2b).
+The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (figure 2a).
+In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 but declining to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (figure 2a).
+Ecosystem carbon accumulation was driven exclusively by increasing soil organic carbon across all four scenarios, all other stocks declined over time (figure 2b).
 
@@
 knitr::include_graphics("./fig_images/fig2_Cstocks.png")
+```
+
+Net primary production (NPP) for the seven main Hawaiian Islands declined across all four scenarios, driven primarily by climate change and to a lesser extent by land use change (Fig.
+3).
+The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (figure S4) and sustained losses of forest and shrubland land area in the high land-use scenario (figure S5).
+In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (figure S4) resulted in R~h~ being ~3% higher by 2100 than under RCP 4.5 (figure 3).
+Land-use change substantially reduced R~h~ in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (figure S5), similar to trends in NPP.
+Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and an average of ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
+Uncertainty around transition-triggered carbon fluxes were higher in the high land-use scenario, driven by greater variability in wildland fire probabilities.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net primary production (NPP), heterotrophic respiration (Rh) and carbon fluxes induced by land use and land-use change for the seven main Hawaiian Islands.
+Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
+knitr::include_graphics("./fig_images/fig3_Cflows.png")
+```
+Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined under all four scenarios over time (figure 4).
+On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under RCP 4.5, with carbon sink strength ~60% greater in the low land-use scenario than in the high land-use scenaro by the year 2100 (figure 4).
+In contrast, ecosystems of the Hawaiian Islands became a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, although the transition from sink to source occured 15 years earlier on average under the high land-use scenario than under the low land-use scenario (figure 4).
+Net biome productivity was also ~40% more negative under the high land-use scenario by the end of the simulation period.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net biome productivity (NBP) for the seven main Hawaiian Islands.
+Values above zero indicate terrestrial ecosystems are acting as a net carbon sink for atmospheric carbon and values below zero indicate ecosystems are acting as a net carbon source to the atmosphere.
+Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
+knitr::include_graphics("./fig_images/fig4_NBPpanels.png")
 ```
@@
 
-Tabular model output data in machine readable format are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ.
-Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.
+Tabular model output data and metadata  are available in machine readable format from a USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
+Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository at https://github.com/selmants/HI_Model.
 

### 2020-09-14 Manuscript/HI_Model_supp.Rmd
updated manuscript files and added supplementary document file

--- 2020-09-08 Manuscript/HI_Model_ms.Rmd
@@
 ---
-title: "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
+title: "Supplemental material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
 bibliography: HI_Model.bib
@@
 urlcolor: blue
-nocite: |
-  @jacobi_map_2017
+
 ---
@@
 \pagebreak
-
-## Abstract
-The State of Hawai`i passed legislation in 2018 setting a goal to be carbon neutral by 2045.
-Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under four unique scenarios over a 90-year timespan.
-Net ecosystem carbon balance declined under all four scenarios.
-Moving from a high to a low radiative forcing scenario reduced net ecosystem carbon loss by ~21%, and net carbon losses were reduced by a total of ~55% under the combined scenario of low radiative forcing and low rates of land‐use change.
-A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to better constrainment of models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
-
-## Introduction
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,
-
-## Methods
-We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
-We also developed a separate set of scenarios to test model sensitivity to different levels of a CO~2~ fertilization effect (CFE).
-The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
-The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019].
-We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
-
-### *Study area*
-The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (Figure 1), a total land area of 16,554 km^2^.
-We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size.
-Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; Figure S1), seven islands, and ten discrete land cover classes (Figure 1).
-
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
-Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
-Water and Wetland land cover classes are not shown."}
-knitr::include_graphics("./fig_images/fig1_HI_LC.png")
-```
-
-### *States and transitions*
-We developed two land-use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
-Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
-There was no transition pathway out of an urbanized (developed) state.
-Water and wetland state types remained static throughout the simulation period.
-
- Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawai`i [@kim_population_2018].
-For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
-For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
-Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
-We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population.
-The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
-Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016].
-In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
-The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
-
-The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2).
-We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019.
-Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
-Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
-The high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig.
-S2a).
-
-### *Carbon stocks and flows*
- The fate of carbon stocks was tracked for each simulation cell based on a suite of carbon flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
-We defined carbon stocks as continuous state variables for each simulation cell, including live biomass, standing dead wood, down dead wood, litter, and soil organic carbon.
-We also included and tracked carbon in atmospheric, aquatic, and harvest product pools to enforce carbon mass balance [@daniel_integrating_2018].
-To transfer carbon between stocks, we defined baseline carbon flows as continuous variables resulting from growth, mortality, deadfall, woody decay, litter decomposition, and leaching (which includes runoff).
-We also defined carbon flows resulting from land use, land use change, and wildfire [@selmants_baseline_2017; @daniel_integrating_2018].
-
- Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (Figure S1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
-We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
-We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016].
-We also calibrated gross photosynthesis in IBIS using a Hawai`i-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
-
- Carbon flow rates for each state type and moisture zone were estimated as the ratio of the IBIS-derived flux to the size of the originating carbon stock at each age [@sleeter_effects_2018].
-A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate.
-This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
-Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of the two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
-The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
-The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario.
-We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014].
-Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.
-
-### *CO~2~ fertilization effect*
-Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well.
-However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans.
-Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE).
-We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario.
-We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
-For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
-This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
-
-### *Scenario simulations and analysis*
-Each of the four unique scenarios were run for 90 years at an annual timestep and repeated for 30 Monte Carlo realizations, using initial conditions corresponding to the year 2010.
-All simulations were performed within the SyncroSim (version 2.2.4) software framework with ST-Sim (version 3.2.13) and SF (version 3.2.10) add-on modules [@daniel_state-and-transition_2016; @daniel_integrating_2018].
-Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
-
-## Results
- Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; Fig.
-2a).
-Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land-use change.
-The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (Fig.
-2a).
-In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 but declining to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (Fig.
-2a).
-Ecosystem carbon accumulation was driven exclusively by increasing soil organic carbon across all four scenarios, all other stocks declined over time (Fig.
-2b).
-
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands.
-Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
-knitr::include_graphics("./fig_images/fig2_Cstocks.png")
-```
-
-## Discussion
-
-## Conclusion
-
-## Acknowledgements
-
-This study was funded by the Biological Sequestration Program of the U.S.
-Geological Survey.
-Thanks to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
-Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.
-
-## Data Availability
-
-Tabular model output data in machine readable format are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ.
-Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.
-
-## ORCID
-
-Paul C. Selmants https://orcid.org/0000-0001-6211-3957
-
-Benjamin M. Sleeter https://orcid.org/0000-0003-2371-9571
-
-Jinxun Liu https://orcid.org/0000-0003-0561-8988
-
-Tamara S. Wilson https://orcid.org/0000-0001-7399-7532
-
-Abby G. Frazier https://orcid.org/0000-0003-4076-4577
-
-Gregory P. Asner https://orcid.org/0000-0001-7893-6421
-
-## References

### 2020-09-22 Manuscript/HI_Model_ms.Rmd
updated manuscript files, edited CarbonBalance R script

@@
 Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
-Parker C. Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^
+Clay Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^
 
@@
 ### *CO~2~ fertilization effect*
-Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well.
-However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans.
+Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well [@Walker_integrating_2020].
+However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans [@Walker_integrating_2020].
 Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE).
@@
 We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
-For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
+For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further stimulation of NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
 This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
@@
 ## Results
+
+### *Carbon stocks and fluxes*
 Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2a).
@@
 ```
+
+### *Ecosystem carbon balance*
 Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined under all four scenarios over time (figure 4).
-On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under RCP 4.5, with carbon sink strength ~60% greater in the low land-use scenario than in the high land-use scenaro by the year 2100 (figure 4).
-In contrast, ecosystems of the Hawaiian Islands became a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, although the transition from sink to source occured 15 years earlier on average under the high land-use scenario than under the low land-use scenario (figure 4).
-Net biome productivity was also ~40% more negative under the high land-use scenario by the end of the simulation period.
+On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under the RCP 4.5 radiative forcing scenario, but carbon sink strength was ~40% lower in the high land-use scenario compared to the low land-use scenario by the end of the simulation period (figure 4).
+In contrast, ecosystems of the Hawaiian Islands acted as a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, with the transition from sink to source occuring 15 years earlier on average in the high land-use scenario than in the low land-use scenario (figure 4).
+In addition, the high land-use scenario under RCP 8.5 represented a ~40% larger net source of carbon to the atmosphere by the year 2100 than the low-land use scenario under the same radiative forcing.
+Over the entire simulation period, both global emissions reduction and avoided land conversion resulted in substantial increases in cumulative NBP (figure 5).
+However, switching from RCP 8.5 to RCP 4.5 increased cumulative NBP in the Hawaiian Islands more than twice as much as reducing emissions from local land-use change and wildfire disturbance (figure 5).
+Switching from RCP 8.5 to RCP 4.5 under the low land-use scenario yielded the greatest cumulative increase in NBP, resulting in a median gain of 26.5 Tg of carbon over the entire 90-year simulation period.
 
@@
 Values above zero indicate terrestrial ecosystems are acting as a net carbon sink for atmospheric carbon and values below zero indicate ecosystems are acting as a net carbon source to the atmosphere.
-Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
+Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations.
+The dashed horizontal line in each panel represents the boundary between ecosystems acting as a net carbon sink (positive NBP values) and a net carbon source (negative NBP values)."}
 knitr::include_graphics("./fig_images/fig4_NBPpanels.png") 
+```
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from high to low land-use change scenario under each radiative forcing scenario (top panel) and when switching from RCP 8.5 to RCP 4.5 radiative forcing scenarios under each land-use scenario (bottom panel).
+Box plots indicate the median (vertical black line), 25th and 75th percentiles (colored boxes), 10th and 90th percentiles (thin horizontal lines), and values outside of this range (black circles).
+Note the different scales for the x-axis in each panel."}
+knitr::include_graphics("./fig_images/fig5_cumdNBP.png") 
+```
+
+### *CO~2~ fertilization effect*
+Projected estimates of both total ecosystem carbon storage and ecosystem carbon balance were highly sensitive to differing rates of a CFE on plant productivity.
+Under the high radiative forcing (RCP 8.5) and high land-use scenario, the inclusion of a CFE ranging from 5-15% led to ~33-98 Tg of additional carbon storage in ecosystems by the end of the century, a ~10-30% increase (figure 6a).
+Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian Island ecosystems from a net carbon source to the atmosphere for the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
+Net carbon sink strength was further enhanced at higher CFE rates, with NBP increasing by an average of 0.07 Tg C y^-1^ for each 1% increase in CFE (figure 6b).
+When compared to other scenarios, the high radiative forcing and high land-use scenario with a 5% CFE resulted in a mean annual NBP of 0.46 ± 0.3 Tg C y^-1^,  roughly equivalent to mean annual NBP of the low radiative forcing and low land-use scenario with no CFE (0.52 ± 0.12).
+A 15% CFE applied to the high radiative forcing and high land-use scenario resulted in a mean annual NBP  of 1.18 ± 0.29 Tg C y^-1^, more than double that of the low radiative forcing and low land-use scenario with no CFE.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of carbon dioxide fertilization in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing scenario and high land use scenario.
+The carbon dioxide fertilization effect (CFE) is the percent change in net primary productivity (NPP) for every 100 ppm increase in atmospheric carbon dioxide.
+The CFE for all rates is capped at 600 ppm, which is achieved around the year 2060.
+Solid lines in (a) indicate the mean total ecosystem carbon storage across 30 Monte Carlo realizations for each CFE rate, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations.
+Solid circles in (b) represent mean annual net biome productivity averaged across all years and Monte Carlo realizations for each CFE rate, with vertical lines indicating the standard deviation of the mean.
+The dashed horizontal line in (b) represents the boundary between ecosystems acting as a net carbon sink (positive NBP values) and a net carbon source (negative NBP values)."}
+knitr::include_graphics("./fig_images/fig6_cfe.png") 
 ```
@@
 
-This study was funded by the Biological Sequestration Program of the U.S.
-Geological Survey.
+This study was funded by the U.S.
+Geological Survey Biological Carbon Sequestration Program.
 Thanks to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
+for information on eucalyptus harvesting in Hawaii and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
 Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.

### 2020-09-22 Manuscript/fig_code/HI_CFE_CarbonBalance.R

@@ -15,7 +15,7 @@ library(cowplot)
 CFE_Cstocks <- read.csv("./HI_Cstocks/Cstocks_CFE.csv", stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in LUCAS C stocks by scenario and filter on RCP85High scenario 
-Cstocks <- read.csv("./HI_Cstocks/Cstocks_Island.csv", stringsAsFactors = FALSE)	%>%
+Cstocks <- read.csv("./HI_Cstocks/Cstocks_Island.csv", stringsAsFactors = FALSE) %>%
 	filter(Scenario == "RCP85High") %>%
 	mutate(CFE_rate = 0) %>%
 	select(CFE_rate, Timestep:Amount)
@@ -24,7 +24,8 @@ Cstocks <- read.csv("./HI_Cstocks/Cstocks_Island.csv", stringsAsFactors = FALSE)
 cfe_baseflow <- read.csv("./HI_Cflows/BaseFlows_CFE.csv", stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in CFE C loss data (transition, fire, harvest)
-cfe_closs <- read.csv("./HI_Cflows/TransitHarvFlows_CFE.csv", stringsAsFactors = FALSE) %>%
+cfe_closs <- read.csv("./HI_Cflows/TransitHarvFlows_CFE.csv", 
+	stringsAsFactors = FALSE) %>%
 	select(-Scenario)
 # read in zero CFE base flow data 
 baseflowzero <- read.csv("./HI_Cflows/BaseFlows.csv", stringsAsFactors = FALSE) %>%
@@ -111,6 +112,7 @@ LUCloss <- cfeCloss %>%
 	group_by(CFE_Rate, Timestep, Iteration) %>%
 	summarize(LUCloss = sum(Amount)/1000) %>%
 	as.data.frame()
+	
 #Join NPP and C loss dataframes together
 cbal <- left_join(NPP, Rh) %>%
 	left_join(., Leach) %>%

### 2020-09-22 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans [@Walker_integrating_2020].
-Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE).
-We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario.
+Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to  different rates of a CO~2~ fertilization effect (CFE).
+We incorporated a CFE multiplier for NPP that represented the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario.
 We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
@@
 Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
-Carbon stocks and fluxes for the seven main Hawaiian Islands were calculated for each scenario by summing within each of the 30 Monte Carlo realizations on an annual basis and then calculating annual means as well as the annual upper and lower limits of the 30 Monte Carlo realizations.
-Carbon balance for the seven main Hawaiian Islands was calculated on annual basis for each scenario and Monte Carlo realization as net biome productivity (NBP), which was equal to annual carbon input in the form of net primary productivity (NPP) minus the annual sum of all carbon losses from terrestrial ecosystems, including heterotrophic respiration (R~h~) from litter and soil, carbon fluxes to the atmosphere triggered by land use and land-use change, wildfire emissions, and aquatic carbon losses through leaching and overland flow.
+Carbon stocks and fluxes for the seven main Hawaiian Islands were calculated for each scenario by summing within each Monte Carlo realization on an annual basis and then calculating annual means as well as the annual upper and lower limits of the 30 Monte Carlo realizations.
+Carbon balance for the seven main Hawaiian Islands was calculated on annual basis for each scenario and Monte Carlo realization as net biome productivity (NBP), which was equal to annual carbon input in the form of NPP minus the annual sum of all carbon losses from terrestrial ecosystems, including heterotrophic respiration (R~h~) from litter and soil, carbon fluxes to the atmosphere triggered by land use and land-use change, wildfire emissions, and aquatic carbon losses through leaching and overland flow.
 Positive NBP values indicated ecosystems of the seven main Hawaiian Islands were acting as a net sink for atmospheric CO~2~, while negative NBP values indicated that these ecosystems were acting as a net carbon source to the atmosphere [@chapin_reconciling_2006].
@@
 Land-use change substantially reduced R~h~ in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (figure S5), similar to trends in NPP.
-Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and an average of ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
-Uncertainty around transition-triggered carbon fluxes were higher in the high land-use scenario, driven by greater variability in wildland fire probabilities.
+Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
+Uncertainty around transition-triggered carbon fluxes were higher in the high land-use scenario, driven primarily by greater variability in wildland fire probabilities.
 
@@
 ### *Ecosystem carbon balance*
-Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined under all four scenarios over time (figure 4).
+Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined over time in all four scenarios  (figure 4).
 On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under the RCP 4.5 radiative forcing scenario, but carbon sink strength was ~40% lower in the high land-use scenario compared to the low land-use scenario by the end of the simulation period (figure 4).
@@
 Net carbon sink strength was further enhanced at higher CFE rates, with NBP increasing by an average of 0.07 Tg C y^-1^ for each 1% increase in CFE (figure 6b).
-When compared to other scenarios, the high radiative forcing and high land-use scenario with a 5% CFE resulted in a mean annual NBP of 0.46 ± 0.3 Tg C y^-1^,  roughly equivalent to mean annual NBP of the low radiative forcing and low land-use scenario with no CFE (0.52 ± 0.12).
+When compared to other scenarios, applying a 5% CFE to the high radiative forcing and high land-use scenario resulted in a mean annual NBP of 0.46 ± 0.3 Tg C y^-1^, roughly equivalent to mean annual NBP in the low radiative forcing and low land-use scenario with no CFE (0.52 ± 0.12).
 A 15% CFE applied to the high radiative forcing and high land-use scenario resulted in a mean annual NBP  of 1.18 ± 0.29 Tg C y^-1^, more than double that of the low radiative forcing and low land-use scenario with no CFE.

### 2020-09-30 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ## Abstract
-The State of Hawai`i passed legislation in 2018 setting a goal to be carbon neutral by 2045.
-Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under four unique scenarios over a 90-year timespan.
-Net ecosystem carbon balance declined under all four scenarios.
-Moving from a high to a low radiative forcing scenario reduced net ecosystem carbon loss by ~21%, and net carbon losses were reduced by a total of ~55% under the combined scenario of low radiative forcing and low rates of land‐use change.
+The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that partly depends on carbon sequestration by terrestrial ecosystems.
+However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
+We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCPs 4.5 and 8.5) over a 90-year timespan from 2010-2100.
+Collectively, ecocystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing throughout the 90-year simulation period, but carbon sink strength capacity was greatest under the combination of low radiative forcing and low land-use change.
+In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing, with high land-use accelerating this transition and exacerbating net carbon loss.
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to better constrainment of models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
+Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to constrain models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
 
 ## Introduction
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,
+Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [Friedlingstein_global_2019].
+
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
 

### 2020-10-14 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ## Abstract
-The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that partly depends on carbon sequestration by terrestrial ecosystems.
+The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depends on carbon sequestration by terrestrial ecosystems.
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCPs 4.5 and 8.5) over a 90-year timespan from 2010-2100.
-Collectively, ecocystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing throughout the 90-year simulation period, but carbon sink strength capacity was greatest under the combination of low radiative forcing and low land-use change.
+We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land-use scenarios (low and high) over a 90-year timespan from 2010-2100.
+Collectively, ecoystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing throughout the 90-year simulation period, but carbon sink strength was greatest under the combination of low radiative forcing and low land-use change.
 In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing, with high land-use accelerating this transition and exacerbating net carbon loss.
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to constrain models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
+Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
 
 ## Introduction
-Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [Friedlingstein_global_2019].
+Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [@friedlingstein_global_2019].
 
@@
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from high to low land-use change scenario under each radiative forcing scenario (top panel) and when switching from RCP 8.5 to RCP 4.5 radiative forcing scenarios under each land-use scenario (bottom panel).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from the high to low land-use change scenario under each radiative forcing scenario (top panel) and when switching from the high (RCP 8.5) to low (RCP 4.5) radiative forcing scenario under each land-use scenario (bottom panel).
 Box plots indicate the median (vertical black line), 25th and 75th percentiles (colored boxes), 10th and 90th percentiles (thin horizontal lines), and values outside of this range (black circles).
-Note the different scales for the x-axis in each panel."}
+Note the different x-axis scales in each panel."}
 knitr::include_graphics("./fig_images/fig5_cumdNBP.png") 
@@
 
-Tabular model output data and metadata are available in machine readable format from a USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
+Tabular model output data and metadata are available in machine readable format from the USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
 Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository at https://github.com/selmants/HI_Model.

### 2020-11-10 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [@friedlingstein_global_2019].
+However, there is a high degree of uncertainty surrounding the strength of the land carbon sink because of its sensitivity to annual variability in climate and human land use and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+These uncertainties are so high that annual estimates of carbon sequestration by terrestrial ecosystems are often estimated indirectly by subtracting annual atmospheric and ocean carbon sink estimates from annual human CO~2~ emissions [@le_quere_global_2013; @friedlingstein_global_2019].
 

### 2020-11-16 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2020-10-14 Manuscript/HI_Model_ms.Rmd
@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [@friedlingstein_global_2019].
+However, there is a high degree of uncertainty surrounding the strength of the land carbon sink because of its sensitivity to annual variability in climate and human land use and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+These climate and land-use driven uncertainties are so high that annual estimates of carbon sequestration by terrestrial ecosystems are often estimated indirectly by subtracting annual atmospheric and ocean carbon sink estimates from annual human CO~2~ emissions [@le_quere_global_2013; @friedlingstein_global_2019].
 

### 2020-11-20 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2020-10-14 Manuscript/HI_Model_ms.Rmd
@@
 We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land-use scenarios (low and high) over a 90-year timespan from 2010-2100.
-Collectively, ecoystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing throughout the 90-year simulation period, but carbon sink strength was greatest under the combination of low radiative forcing and low land-use change.
-In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing, with high land-use accelerating this transition and exacerbating net carbon loss.
+Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land-use change further enhancing carbon sink strength.
+In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land-use accelerating this transition and exacerbating net carbon loss.
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [@friedlingstein_global_2019].
+However, the strength of the land carbon sink remains highly uncertain because of its sensitivity to interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+Accounting for climate and land-use variability has proven so difficult that annual estimates of terrestrial ecosystem carbon sequestration are often estimated indirectly by subtracting annual atmospheric and ocean carbon sink estimates from annual human CO~2~ emissions [@le_quere_global_2013; @friedlingstein_global_2019].
 
@@
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
-The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019].
+The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019].
 We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
@@
 However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans [@Walker_integrating_2020].
-Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to  different rates of a CO~2~ fertilization effect (CFE).
+Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model projections of ecosystem carbon balance to different rates of a CO~2~ fertilization effect (CFE).
 We incorporated a CFE multiplier for NPP that represented the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario.
@@
 In contrast, ecosystems of the Hawaiian Islands acted as a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, with the transition from sink to source occuring 15 years earlier on average in the high land-use scenario than in the low land-use scenario (figure 4).
-In addition, the high land-use scenario under RCP 8.5 represented a ~40% larger net source of carbon to the atmosphere by the year 2100 than the low-land use scenario under the same radiative forcing.
-Over the entire simulation period, both global emissions reduction and avoided land conversion resulted in substantial increases in cumulative NBP (figure 5).
+The high land-use scenario under RCP 8.5 represented a ~40% larger net source of carbon to the atmosphere by the year 2100 than the low-land use scenario under the same radiative forcing.
+Over the entire simulation period, both global emissions reductions and local avoided land conversion resulted in substantial increases in cumulative NBP (figure 5).
 However, switching from RCP 8.5 to RCP 4.5 increased cumulative NBP in the Hawaiian Islands more than twice as much as reducing emissions from local land-use change and wildfire disturbance (figure 5).
@@
 Geological Survey Biological Carbon Sequestration Program.
-Thanks to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting in Hawaii and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software.
+Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software and to Nicholas Koch of Forest Solutions, Inc.
+for information on eucalyptus harvesting in Hawaii.
 Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.

### 2020-12-21 Manuscript/HI_Model_ms.Rmd
updated manuscript file

@@
 ## Introduction
-Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [@friedlingstein_global_2019].
-However, the strength of the land carbon sink remains highly uncertain because of its sensitivity to interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
-Accounting for climate and land-use variability has proven so difficult that annual estimates of terrestrial ecosystem carbon sequestration are often estimated indirectly by subtracting annual atmospheric and ocean carbon sink estimates from annual human CO~2~ emissions [@le_quere_global_2013; @friedlingstein_global_2019].
+Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019].
+However, the strength of the land carbon sink remains highly uncertain because of its sensitivity to interannual variability in climate, land use and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+The uncertainty surrounding estimates of terrestrial carbon sequestration complicates efforts to employ "natural carbon solutions", defined as conservation, restoration, and land management interventions aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017].
+A growing number of sub-national jurisdictions have pledged to incorporate natural climate solutions as part of their stated goal to achieve carbon neutrality by mid-century [].
 

### 2020-12-30 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ## Abstract
-The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depends on carbon sequestration by terrestrial ecosystems.
+The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems.
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019].
-However, the strength of the land carbon sink remains highly uncertain because of its sensitivity to interannual variability in climate, land use and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
-The uncertainty surrounding estimates of terrestrial carbon sequestration complicates efforts to employ "natural carbon solutions", defined as conservation, restoration, and land management interventions aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017].
-A growing number of sub-national jurisdictions have pledged to incorporate natural climate solutions as part of their stated goal to achieve carbon neutrality by mid-century [].
+There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
+Natural climate solutions are defined as conservation and land management efforts aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017].
+A growing number of sub-national jurisdictions have set goals to achieve carbon neutrality by mid-century, and plan to incorporate natural climate solutions because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
+However, effective implementation is complicated by the considerable uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
+
+Interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+
+Hawaii was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement.
 

### 2021-01-05 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Natural climate solutions are defined as conservation and land management efforts aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017].
-A growing number of sub-national jurisdictions have set goals to achieve carbon neutrality by mid-century, and plan to incorporate natural climate solutions because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
+A growing number of sub-national jurisdictions have set goals to achieve carbon neutrality by mid-century, with most planning to incorporate natural climate solutions because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
 However, effective implementation is complicated by the considerable uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.

### 2021-01-05 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2020-12-30 Manuscript/HI_Model_ms.Rmd
@@
 There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
-Natural climate solutions are defined as conservation and land management efforts aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017].
-A growing number of sub-national jurisdictions have set goals to achieve carbon neutrality by mid-century, and plan to incorporate natural climate solutions because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
-However, effective implementation is complicated by the considerable uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
+Natural climate solutions, defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
+A growing number of sub-national jurisdictions plan to incorporate natural climate solutions as part of a suite of emissions reduction efforts to achieve carbon neutrality by mid-century.
+However, uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale, complicates efforts to effectively implement natural climate solutions.
 
@@
 
-Hawaii was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement.
+Hawaii was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement. Specifically, Hawaii passed Act
 

### 2021-01-12 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ## Introduction
-Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019].
+Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018].
+Interannual fluctuations in carbon uptake by land ecosystems is the primary driver of year-to-year variation in atmosphereric CO~2~ growth rate @friedlingstein_global_2019].
 There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
@@
 A growing number of sub-national jurisdictions plan to incorporate natural climate solutions as part of a suite of emissions reduction efforts to achieve carbon neutrality by mid-century.
-However, uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale, complicates efforts to effectively implement natural climate solutions.
+However, uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale, complicates efforts to effectively implement ecosystem-based carbon climate mitigation strategies.
 

### 2021-01-13 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2021-01-05 Manuscript/HI_Model_ms.Rmd
@@
 ## Introduction
-Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019].
+Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019].
 There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
-Natural climate solutions, defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
-A growing number of sub-national jurisdictions plan to incorporate natural climate solutions as part of a suite of emissions reduction efforts to achieve carbon neutrality by mid-century.
-However, uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale, complicates efforts to effectively implement natural climate solutions.
+Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
+A growing number of sub-national jurisdictions have incorporated  as part of a suite of emissions reduction efforts to achieve carbon neutrality.
+However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
 
+Interannual fluctuations in carbon uptake by land ecosystems is the primary driver of year-to-year variation in atmosphereric CO~2~ growth rate
 Interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].

### 2021-01-14 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
-A growing number of sub-national jurisdictions have incorporated  as part of a suite of emissions reduction efforts to achieve carbon neutrality.
 However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
+Despite this uncertainty, evidence indicates that variability in carbon uptake by land ecosystems, both year-to-year and over the long term, is primarily driven by fluctuations in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017; @friedlingstein_global_2019].
+Spatially explicit future projections of ecosystem carbon balance incorporating the interactive effects of land use and climate could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
+Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
-Interannual fluctuations in carbon uptake by land ecosystems is the primary driver of year-to-year variation in atmosphereric CO~2~ growth rate
-Interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+The State of Hawai`i exemplifies the challenges and complexity of projecting the interactive effects of climate and land use change on ecosystem carbon balance.
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
+Mean annual temperature ranges from ~4 - 24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180 - 9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate.
+Ecosystem carbon stocks across the Hawaiian Islands are still strongly influenced by the legacy of past land-use change [@asner_high-resolution_2011].
+Thousands of hectares of land were deforested beginning in the late 19th century to make way for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
+Beginning in the latter half of the 20th century, this agricultural land has since been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
 
@@
 
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
+We used a fully coupled consider four unique future scenarios
 

### 2021-01-14 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2021-01-13 Manuscript/HI_Model_ms.Rmd
@@
 Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
-A growing number of sub-national jurisdictions have incorporated  as part of a suite of emissions reduction efforts to achieve carbon neutrality.
 However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
+Despite this uncertainty, evidence indicates that both interannual and long-term variability in carbon uptake by land ecosystems is driven primarily by fluctuations in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017; @friedlingstein_global_2019].
+Spatially explicit future projections of ecosystem carbon balance incorporating the interactive effects of land use and climate could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
+Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
-Interannual fluctuations in carbon uptake by land ecosystems is the primary driver of year-to-year variation in atmosphereric CO~2~ growth rate
-Interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].
+The State of Hawai`i exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance.
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
+Mean annual temperature ranges from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180-9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate.
+Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970's [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
+These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017].
+Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@asner_high-resolution_2011].
+Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
+Beginning in the latter half of the 20th century, this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
 

### 2021-01-15 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 csl: iop.csl
-fontsize: 12pt
 geometry: margin=1in
@@
     latex_engine: xelatex
+mainfont: Times New Roman
+fontsize: 12pt
 urlcolor: blue
@@
 ```
+\centering
 Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,
+
 Clay Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^
+
+\raggedright
 
@@
 ^2^U.S. Geological Survey, Seattle, WA, USA    
-^3^University of Hawai`i at Mānoa, Honolulu, HI, USA  
+^3^University of Hawaiʻi at Mānoa, Honolulu, HI, USA  
 ^4^The East-West Center, Honolulu, HI, USA    
@@
 ## Abstract
-The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems.
+The State of Hawaiʻi passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems.
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
@@
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.
+Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.
 
@@
 
-The State of Hawai`i exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance.
+The State of Hawaiʻi exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance.
 The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
@@
 
-Hawaii was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement. Specifically, Hawaii passed Act
+Hawaiʻi was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement. Specifically, Hawaiʻi passed Act
 
@@
 The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019].
-We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
+We parameterized the Hawai"i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
 
@@
 
- Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawai`i [@kim_population_2018].
+ Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
 For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
@@
 We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016].
-We also calibrated gross photosynthesis in IBIS using a Hawai`i-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
+We also calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
 
@@
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of carbon dioxide fertilization in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing scenario and high land use scenario.
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of carbon dioxide fertilization in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing and high land use scenario.
 The carbon dioxide fertilization effect (CFE) is the percent change in net primary productivity (NPP) for every 100 ppm increase in atmospheric carbon dioxide.
@@
 Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software and to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting in Hawaii.
+for information on eucalyptus harvesting in Hawaiʻi.
 Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.

### 2021-01-15 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017].
-Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@asner_high-resolution_2011].
+Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@osher_effect_2003; @asner_high-resolution_2011].
 Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].

### 2021-01-21 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Despite this uncertainty, evidence indicates that both interannual and long-term variability in carbon uptake by land ecosystems is driven primarily by fluctuations in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017; @friedlingstein_global_2019].
-Spatially explicit future projections of ecosystem carbon balance incorporating the interactive effects of land use and climate could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
+Incorporating the interactive effects of land use and climate into spatially explicit future projections of ecosystem carbon balance could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
 Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
-The State of Hawaiʻi exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance.
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
-Mean annual temperature ranges from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180-9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate.
-Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970's [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
+The State of Hawaiʻi exemplifies the challenges of projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale.
+Hawaiʻi was the first U.S.
+state to enact legislation committing to full carbon neutrality, requiring the state to account for and offset all of its greenhouse gas emissions by 2045 (State of Hawaiʻi Acts 15 and 16).
+This legislation emphasizes the mitigation potential of natural ecosystems as a key component to emissions reduction, necessitating baseline estimates and future projections of land carbon sink strength.
+However, Hawaiʻi's challenging terrain complicates these assessment efforts.
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
+Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
 These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017].
@@
 Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
-Beginning in the latter half of the 20th century, this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
-
-Hawaiʻi was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement. Specifically, Hawaiʻi passed Act
+Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
 

### 2021-01-21 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
-The State of Hawaiʻi exemplifies the challenges of projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale.
+The State of Hawaiʻi exemplifies the challenges associated with projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale.
 Hawaiʻi was the first U.S.
@@
 Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
+These past trends surely inform the future impact of climate and land-use change on ecosystem carbon balance, but remain difficult to predict with any degree of certainty.
+To date only one study has attempted to integrate the major controlling processes of land-use, climate, and natural disturbances into projections of carbon balance for Hawaiian ecosystems, and that study was restricted to a single land-use change scenario, a single radiative forcing scenario (SRES A1B, equivalent to RCP 6), and a projection period extending only to the mid-21st century [@selmants_baseline_2017].
 

### 2021-01-22 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2021-01-21 Manuscript/HI_Model_ms.Rmd
@@
 Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
+Although these past trends surely inform the future impact of climate and land-use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts.
+To date only one study has attempted to integrate land-use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land-use change scenario and moderate radiative forcing [SRES A1B, equivalent to RCP 6; @selmants_baseline_2017].
 
-We used a fully coupled consider four unique future scenarios
+In this study, we consider four unique scenarios exploring a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions.
+We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands under each of these four scenarios on an annual basis for the period 2010–2100 [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
+The four unique scenarios represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
+In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary over time in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
+The goals of this study were to (a) estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, (b) quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and (c) assess the sensitivity of model estimates to the introduction of a CFE on NPP.
 
@@
 The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019].
-We parameterized the Hawai"i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
+We parameterized the Hawaiʻi LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.
 

### 2021-01-22 Manuscript/HI_Model_supp.Rmd
updated manuscript files

@@
 ```
+\centering
 Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
@@
 
+\raggedright
 **Affiliations:**  
@@
 ^2^U.S. Geological Survey, Seattle, WA, USA  
-^3^University of Hawai`i at Mānoa, Honolulu, HI, USA  
+^3^University of Hawaiʻi at Mānoa, Honolulu, HI, USA  
 ^4^The East-West Center, Honolulu, HI, USA  

### 2021-01-25 Manuscript/HI_Model_ms.Rmd
updated manuscript files, added MOD17 NPP data and R code, updated README file

@@
 However, Hawaiʻi's challenging terrain complicates these assessment efforts.
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances (Fig.
+S3), with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
 Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].

### 2021-01-25 MOD17_NPP/code/MOD17_processing.R

@@ -1,0 +1,22 @@
+## Calculate annual statewide NPP 
+## for the years 2010-2019
+## using MODIS MOD17 NPP estimates 
+## Paul C. Selmants
+## 2021-01-25 (ISO 8601)
+
+## R version 3.6.1
+
+# load required packages into R
+library(raster) #version 3.1-5
+library(rgdal) #version 1.4-8
+library(tidyr, warn.conflicts = FALSE) #version 1.0.2
+library(dplyr, warn.conflicts = FALSE) #version 0.8.5
+library(ggplot2) #version
+
+## Set working directory at highest level in HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model 
+
+# create list of annual MOD17 NPP .tif files
+rainlist <- list.files('./MOD17_NPP/data', full.names = TRUE)
+#create raster stack of annual NPP .tif files
+nppstack <- stack(npplist)

### 2021-01-25 Manuscript/HI_Model_ms.Rmd
updated manuscript file

@@
 Geological Survey Biological Carbon Sequestration Program.
-Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software and to Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting in Hawaiʻi.
+Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, Nicholas Koch of Forest Solutions, Inc.
+for information on eucalyptus harvesting in Hawaiʻi, and Caroline Conrad for help with Google Earth Engine scripting.
 Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.

### 2021-01-26 MOD17_NPP/code/MOD17_processing.R
updated MOD17 R script

@@ -17,6 +17,17 @@ library(ggplot2) #version
 ## GitHub repository, https://github.com/selmants/HI_Model 
 
 # create list of annual MOD17 NPP .tif files
-rainlist <- list.files('./MOD17_NPP/data', full.names = TRUE)
+npplist <- list.files('./MOD17_NPP/data', full.names = TRUE)
 #create raster stack of annual NPP .tif files
-nppstack <- stack(npplist)
+stack <- stack(npplist)
+# multiply all values by MODIS scale factor (0.0001)	
+# to get NPP values in kg C m-2 y-1
+npp <- stack * 1e-04
+
+# export pixel NPP data to dataframe
+# and sum values within each year to get 
+# statewide NPP estimates in Tg C y-1
+npp_data <- as.data.frame(npp, xy=TRUE, na.rm = TRUE) %>%
+
+
+

### 2021-02-01 MOD17_NPP/code/MOD17_processing.R
added fire emissions R  script and updated MOD17 R script

@@ -4,14 +4,14 @@
 ## Paul C. Selmants
 ## 2021-01-25 (ISO 8601)
 
-## R version 3.6.1
+## R version 4.0.2
 
 # load required packages into R
-library(raster) #version 3.1-5
-library(rgdal) #version 1.4-8
-library(tidyr, warn.conflicts = FALSE) #version 1.0.2
-library(dplyr, warn.conflicts = FALSE) #version 0.8.5
-library(ggplot2) #version
+library(raster) #version 3.3-13
+library(rgdal) #version 1.4-16
+library(tidyr, warn.conflicts = FALSE) #version 1.1.1
+library(dplyr, warn.conflicts = FALSE) #version 1.0.2
+library(ggplot2) #version 3.3.2
 
 ## Set working directory at highest level in HI_Model
 ## GitHub repository, https://github.com/selmants/HI_Model 
@@ -24,10 +24,16 @@ stack <- stack(npplist)
 # to get NPP values in kg C m-2 y-1
 npp <- stack * 1e-04
 
-# export pixel NPP data to dataframe
-# and sum values within each year to get 
+# export pixel NPP data to dataframe, 
+# pivot to long format, and sum values to get 
 # statewide NPP estimates in Tg C y-1
 npp_data <- as.data.frame(npp, xy=TRUE, na.rm = TRUE) %>%
+	pivot_longer(!x:y, names_to = "year", values_to = "npp") %>%
+	group_by(year) %>%
+	summarise(NPP_Tgy = (sum(npp)*2.5e5)*1e-9) %>%
+	mutate(Year = as.numeric(gsub("hawaii_npp_", "", year))) %>%
+	select(Year, NPP_Tgy) %>%
+	as.data.frame() 
 
 
 

### 2021-02-01 Manuscript/fig_code/HI_FireEmissions.R

@@ -1,0 +1,25 @@
+## Annual fire emissions estimates, 
+## State of Hawaii, 2011-2100
+## Paul C. Selmants
+## R version 3.6.1
+## 2021-02-01
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(cowplot)
+
+## Download LUCAS model output data from USGS ScienceBase: 
+## https://doi.org/10.5066/P9AWLFKZ
+
+# read HI LUCAS Transition/Harvest C Loss data into R and 
+# filter on fire emissions, summing by Island, Iteration, and Scenario
+fire <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
+	filter(TransitionGroupID == "FIRE", 
+		grepl("Emission", FlowGroupID)) %>%
+		group_by(Scenario, Timestep, Iteration, FromSecondaryStratumID) %>%
+		summarize(fire_emiss = sum(Amount)) %>%
+		as.data.frame() %>%
+		select(Scenario, Timestep, Iteration, Island = FromSecondaryStratumID,
+			fire_emiss)

### 2021-02-01 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Incorporating the interactive effects of land use and climate into spatially explicit future projections of ecosystem carbon balance could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
-Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
+Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future land-use and climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
@@
 The four unique scenarios represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
-In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary over time in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
+In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
 The goals of this study were to (a) estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, (b) quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and (c) assess the sensitivity of model estimates to the introduction of a CFE on NPP.
@@
 This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
-Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of the two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
+Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
 The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
@@
 
+Over the last decade, terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon.
+Net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios for the time period 2011-2019.
+
 ## Conclusion
@@
 
-## Data Availability
+## Data and Code Availability
 

### 2021-02-02 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
-Over the last decade, terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon.
+Over the last decade, we estimate that terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon (figure 4).
 Net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios for the time period 2011-2019.
+Based on these estimates, terrestrial ecosystems offset approximately 13% of 2015 CO~2~ emissions from energy production and transportation (5.04 TgC), the Stateʻs largest source of greenhouse gas emissions [@hawaii_inventory_2019].
+Our model mean NBP estimate for 2011-2019 agrees well with the State of Hawaiʻi's 2015 net carbon sink estimate of 0.66 TgC y^1^ from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
 

### 2021-02-09 Manuscript/HI_Model_ms.Rmd
updated manuscript files and added R script for fire emissions graph

--- 2021-02-01 Manuscript/HI_Model_ms.Rmd
@@
 
-**Running title:** Hawaii carbon balance  
-**Keywords:** land use, climate change, carbon balance, Hawaii, scenarios, disturbance, ecosystem model
+**Running title:** Hawaiʻi carbon balance  
+**Keywords:** land use, climate change, carbon balance, Hawaiʻi, scenarios, disturbance, ecosystem model
 
@@
 
-In this study, we consider four unique scenarios exploring a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions.
+Here we consider four unique scenarios exploring a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions.
 We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands under each of these four scenarios on an annual basis for the period 2010–2100 [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-The four unique scenarios represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
+The four unique scenarios represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
-The goals of this study were to (a) estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, (b) quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and (c) assess the sensitivity of model estimates to the introduction of a CFE on NPP.
+our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.
 
@@
 
-Over the last decade, terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon.
+We estimated that terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon over the last decade (figure 4).
 Net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios for the time period 2011-2019.
+Based on these LUCAS model estimates, Hawaiian terrestrial ecosystems offset approximately 13% of 2015 statewide CO~2~ emissions from energy production and transportation (5.04 TgC), the State of Hawaiʻi's largest source of greenhouse gas emissions [@hawaii_inventory_2019].
+Future projections indicate Hawaiian terrestrial ecosystems will continue to be a net sink for atmospheric carbon if global CO~2~ emissions peak around 2040 and then decline (RCP 4.5), and that reducing the intensity of future land use change in Hawaiʻi will further enhance carbon sink strength.
+If, however, global emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high intensity land-use change accelerating this transition and exacerbating net carbon loss.
+
+Our NBP estimates of ~ 0.64 TgC y^-1^ for the recent past (2011-2019) agree well with those from a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
 

### 2021-02-09 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 If, however, global emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high intensity land-use change accelerating this transition and exacerbating net carbon loss.
+Our model results also indicate that projections of ecosystem carbon balance are highly sensitive to the introduction of a CFE.
+Even a 5% increase in NPP per 100ppm increase in atmospheric CO~2~ was sufficient to maintain Hawaiian Island ecosystems as a net carbon sink under the high radiative forcing and high land-use scenario for the entire simulation period.
+Reconciling the high uncertainty surrounding the response of net photosynthesis to rising atmospheric CO~2~ is essential to more realistically constrain simulation model projections.
 

### 2021-02-10 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ### *Carbon stocks and fluxes*
-Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2a).
+Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010 (figure 2a), with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2b).
 Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land-use change.
 The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (figure 2a).
-In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 but declining to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (figure 2a).
+In contrast, high radiative forcing and high land use change resulted in the lowest ecosystem carbon gain, reaching a peak of ~332 Tg in 2063 and a decline to 327 Tg in 2100, resulting in a net increase of only 3% by the end of the simulation period (figure 2a).
 Ecosystem carbon accumulation was driven exclusively by increasing soil organic carbon across all four scenarios, all other stocks declined over time (figure 2b).
@@
 The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (figure S4) and sustained losses of forest and shrubland land area in the high land-use scenario (figure S5).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (figure S4) resulted in R~h~ being ~3% higher by 2100 than under RCP 4.5 (figure 3).
-Land-use change substantially reduced R~h~ in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (figure S5), similar to trends in NPP.
+In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (figure S4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
+Heterotrophic respiration declined substantially over time in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (figure S5), similar to trends in NPP.
 Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
@@
 We estimated that terrestrial ecosystems of the Hawaiian Islands have been a consistent net sink for atmospheric carbon over the last decade (figure 4).
-Net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios for the time period 2011-2019.
-Based on these LUCAS model estimates, Hawaiian terrestrial ecosystems offset approximately 13% of 2015 statewide CO~2~ emissions from energy production and transportation (5.04 TgC), the State of Hawaiʻi's largest source of greenhouse gas emissions [@hawaii_inventory_2019].
-Future projections indicate Hawaiian terrestrial ecosystems will continue to be a net sink for atmospheric carbon if global CO~2~ emissions peak around 2040 and then decline (RCP 4.5), and that reducing the intensity of future land use change in Hawaiʻi will further enhance carbon sink strength.
-If, however, global emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high intensity land-use change accelerating this transition and exacerbating net carbon loss.
+For the time period 2011-2019, net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios.
+Based on this mean annual NBP estimate, Hawaiian terrestrial ecosystems offset approximately 13% of 2015 statewide CO~2~ emissions from energy production and transportation (5.04 TgC), the State of Hawaiʻi's largest source of greenhouse gas emissions [@hawaii_inventory_2019].
+Future projections indicate Hawaiian terrestrial ecosystems will continue to be a net sink for atmospheric carbon if global CO~2~ emissions peak around 2040 and then decline (RCP 4.5), and that carbon sink strength can be further enhanced by reducing the intensity and extent of future land use change.
+If, however, global CO~2~ emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high levels of land-use change accelerating this transition and exacerbating net carbon loss.
 Our model results also indicate that projections of ecosystem carbon balance are highly sensitive to the introduction of a CFE.
-Even a 5% increase in NPP per 100ppm increase in atmospheric CO~2~ was sufficient to maintain Hawaiian Island ecosystems as a net carbon sink under the high radiative forcing and high land-use scenario for the entire simulation period.
+Even a 5% increase in NPP for every 100ppm increase in atmospheric CO~2~ was sufficient to completely offset all other carbon losses induced by the high radiative forcing and high land use scenario, maintaining Hawaiian Island ecosystems as a net carbon sink for the entire simulation period instead of transitioning to a net carbon source by mid-century.
 Reconciling the high uncertainty surrounding the response of net photosynthesis to rising atmospheric CO~2~ is essential to more realistically constrain simulation model projections.
 
-Our NBP estimates of ~ 0.64 TgC y^-1^ for the recent past (2011-2019) agree well with those from a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
+### *Drivers of ecosystem carbon balance*
+
+### *Comparison to other studies*
+There are very few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance.
+Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2019 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
+
+### *Limitations of this study*
 

### 2021-02-16 Manuscript/HI_Model_ms.Rmd
updated Fire R scripts and manuscript files

@@
 
-## Introduction
+## 1. Introduction
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019].
@@
 
-Here we consider four unique scenarios exploring a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions.
-We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands under each of these four scenarios on an annual basis for the period 2010–2100 [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-The four unique scenarios represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
+We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions  [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
+We explored four unique scenarios that represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
@@
 
-## Methods
+## 2. Methods
 We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
@@
 
-### *Study area*
+### *2.1 Study area*
 The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (figure 1), a total land area of 16,554 km^2^.
@@
 
-### *States and transitions*
+### *2.2 States and transitions*
 We developed two land-use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
@@
 
-### *Carbon stocks and flows*
+### *2.3 Carbon stocks and flows*
  The fate of carbon stocks was tracked for each simulation cell based on a suite of carbon flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
@@
 
-### *CO~2~ fertilization effect*
+### *2.4 CO~2~ fertilization effect*
 Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well [@Walker_integrating_2020].
@@
 
-### *Scenario simulations and analysis*
+### *2.5 Scenario simulations and analysis*
 Each of the four unique scenarios were run for 90 years at an annual timestep and repeated for 30 Monte Carlo realizations, using initial conditions corresponding to the year 2010.
@@
 
-## Results
+## 3. Results
 
-### *Carbon stocks and fluxes*
+### *3.1 Carbon stocks and fluxes*
 Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010 (figure 2a), with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2b).
@@
 
-### *Ecosystem carbon balance*
+### *3.2 Ecosystem carbon balance*
 Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined over time in all four scenarios (figure 4).
@@
 
-### *CO~2~ fertilization effect*
+### *3.3 CO~2~ fertilization effect*
 Projected estimates of both total ecosystem carbon storage and ecosystem carbon balance were highly sensitive to differing rates of a CFE on plant productivity.
 Under the high radiative forcing (RCP 8.5) and high land-use scenario, the inclusion of a CFE ranging from 5-15% led to ~33-98 Tg of additional carbon storage in ecosystems by the end of the century, a ~10-30% increase (figure 6a).
-Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian Island ecosystems from a net carbon source to the atmosphere for the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
+Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian Island ecosystems from a net carbon source to the atmosphere during the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
 Net carbon sink strength was further enhanced at higher CFE rates, with NBP increasing by an average of 0.07 Tg C y^-1^ for each 1% increase in CFE (figure 6b).
@@
 
-## Discussion
+## 4. Discussion
 
@@
 Even a 5% increase in NPP for every 100ppm increase in atmospheric CO~2~ was sufficient to completely offset all other carbon losses induced by the high radiative forcing and high land use scenario, maintaining Hawaiian Island ecosystems as a net carbon sink for the entire simulation period instead of transitioning to a net carbon source by mid-century.
-Reconciling the high uncertainty surrounding the response of net photosynthesis to rising atmospheric CO~2~ is essential to more realistically constrain simulation model projections.
+Reconciling the high uncertainty surrounding the response of net photosynthesis to rising atmospheric CO~2~ is essential to more realistically constrain model projections of ecosystem carbon balance.
 
-### *Drivers of ecosystem carbon balance*
+### *4.1 Drivers of ecosystem carbon balance*
 
-### *Comparison to other studies*
-There are very few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance.
+### *4.2 Comparison to other studies*
+There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.
 Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2019 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
 
-### *Limitations of this study*
+### *4.3 Limitations of this study*
 
-## Conclusion
+## 5. Conclusion
 

### 2021-02-16 Manuscript/HI_Model_supp.Rmd
updated manuscript files, added MOD17 NPP data and R code, updated README file
updated Fire R scripts and manuscript files

@@
 urlcolor: blue
-
+nocite: |
+  @jacobi_map_2017
 ---
@@
 \pagebreak
+
+## Moisture Zones
+
+State Classes for Agriculture, Forest, Grassland, Shrubland, and Tree Plantations were stratified into three Moisture Zones - Dry, Mesic, and Wet (Figure S1).
+These three zones were based on a moisture availability index (MAI), calculated as mean annual precipitation (MAP) minus potential evapotranspiration [PET; @price_mz_2012].
+Areas where the MAI was less than zero, i.e.
+where PET exceeded MAP, were classified in the Dry Moisture Zone.
+Areas with MAI values between zero and 1,661 were classified in the Mesic Moisture Zone.
+The MAI value of 1,661 is roughly equivalent to areas at 1,000m elevation that receive 2,500 mm of annual rainfall [@price_mz_2012].
+Areas with MAI values that exceeded 1,661 were classified in the Wet Mesic Zone.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi et al (2017)."}
+knitr::include_graphics("./fig_images/figS1_MZ.png")
+```
+
+## Wildland Fire
+
+This study used a new spatial database of wildland fire perimeters in the Hawaiian Islands for the years 1999-2019 compiled using data from the Monitoring Trends in Burn Severity (MTBS) program, the Hawaiʻi Wildfire Management Organization, the U.S.
+National Park Service, the Oʻahu Army Natural Resources Program, and independent remote sensing analyses using publicly available LandSat data.
+This new wildland fire spatial database was used to calculate annual area burned (Figure S2), wildland fire probabilities, and wildland fire size distributions.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
+The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
+knitr::include_graphics("./fig_images/figS2_AnnualAreaBurned.png")
+```
+
+## Climate
+
+Spatially explicit mean annual data on contemporary temperature and rainfall for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
+Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5 (Figure S5).
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands."}
+knitr::include_graphics("./fig_images/figS3_CurrentClimate.png")
+```
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by end of century under RCPs 4.5 and 8.5 (top panels) based on statistical downscaling of an ensemble of climate models (CMIP5)."}
+knitr::include_graphics("./fig_images/figS4_FutureClimate.png")
+```
+
+## Change in State Class Area
+
+Projections of total land area covered by each State Class were summed by year across the seven main Hawaiian Islands from LUCAS model output data.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100.
+Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
+knitr::include_graphics("./fig_images/figS5_StateClassArea.png")
+```
+
+## References

### 2021-02-16 Fire/code/FireSizeDistribution.R

@@ -10,7 +10,7 @@ library(raster)
 library(dplyr)
 
 # read in Fire Perimeters shapefile
-fires <- shapefile("./data/base/Hawaii Fire Perimeters 1999-2018.shp")
+fires <- shapefile("./Fire/data/base/2019_1999_Hawaii_Fire_Perimeters.shp")
 
 # calculate area of each polygon in km2
 fires$area_km2 <- area(fires)/1000000

### 2021-02-16 Manuscript/fig_code/HI_FireEmissions.R

@@ -23,3 +23,4 @@ fire <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
 		as.data.frame() %>%
 		select(Scenario, Timestep, Iteration, Island = FromSecondaryStratumID,
 			fire_emiss)
+

### 2021-02-19 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 However, Hawaiʻi's challenging terrain complicates these assessment efforts.
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances (Fig.
-S3), with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances (supplmental figure 3), with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
 Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
@@
 We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
-We also developed a separate set of scenarios to test model sensitivity to different levels of a CO~2~ fertilization effect (CFE).
+We also developed a separate set of scenarios to test model sensitivity to different levels of a CFE on NPP.
 The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
@@
 We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size.
-Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; figure S1), seven islands, and ten discrete land cover classes (figure 1).
+Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; supplemental figure 1), seven islands, and ten discrete land cover classes (figure 1).
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi et al (2017).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017).
 Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model.
@@
 Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
-There was no transition pathway out of an urbanized (developed) state.
+Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
+Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state.
 Water and wetland state types remained static throughout the simulation period.
@@
 
-The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (figure S2).
-We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (figure S1), island, and state type (figure 1) for the years 1999-2019.
+The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (supplemental figure 2).
+We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (supplemental figure 1), island, and state type (figure 1) for the years 1999-2019.
 Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
 Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
-The high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig.
-S2a).
+The high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a).
 
@@
 
- Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (figure S1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+ Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (supplmental figure 1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
@@
 ```
-
 Net primary production (NPP) for the seven main Hawaiian Islands declined across all four scenarios, driven primarily by climate change and to a lesser extent by land use change (Fig.
 3).
-The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (figure S4) and sustained losses of forest and shrubland land area in the high land-use scenario (figure S5).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (figure S4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
-Heterotrophic respiration declined substantially over time in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (figure S5), similar to trends in NPP.
+The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land-use scenario (supplmental figure 5).
+In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplmental figure 4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
+Heterotrophic respiration declined substantially over time in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 5), similar to trends in NPP.
 Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
@@
 ```
-
 ### *3.2 Ecosystem carbon balance*
@@
 ```
-
 ```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from the high to low land-use change scenario under each radiative forcing scenario (top panel) and when switching from the high (RCP 8.5) to low (RCP 4.5) radiative forcing scenario under each land-use scenario (bottom panel).
@@
 ```
-
 ### *3.3 CO~2~ fertilization effect*
@@
 ```
-
 ## 4. Discussion
@@
 
-### *4.1 Drivers of ecosystem carbon balance*
+### *4.1 Impact of different climate and land-use pathways* 
+By comparing model results of Hawaiian ecosystem carbon balance under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5).
+Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period.
+Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions.
+Switching from the ‘high’ to the ‘low’ land‐use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100.
+The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C).
+Notably,
 
@@
 There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.
-Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2019 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
+Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
+In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period {0.314 Tg C y^-1^; @selmants_baseline_2017].
+Our current model estimates of NPP and R~h~ being 9.5% and 15% lower, respectively [
 
@@
 
-## Data and Code Availability
+## Data and code availability
 

### 2021-02-19 Manuscript/HI_Model_supp.Rmd
updated manuscript files

@@
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi et al (2017)."}
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
 knitr::include_graphics("./fig_images/figS1_MZ.png")
 ```
-
 ## Wildland Fire
 
-This study used a new spatial database of wildland fire perimeters in the Hawaiian Islands for the years 1999-2019 compiled using data from the Monitoring Trends in Burn Severity (MTBS) program, the Hawaiʻi Wildfire Management Organization, the U.S.
+This study used a new spatial database of wildland fire perimeters in the Hawaiian Islands for the years 1999-2019 using data compiled from the Monitoring Trends in Burn Severity (MTBS) program, the Hawaiʻi Wildfire Management Organization, the U.S.
 National Park Service, the Oʻahu Army Natural Resources Program, and independent remote sensing analyses using publicly available LandSat data.
-This new wildland fire spatial database was used to calculate annual area burned (Figure S2), wildland fire probabilities, and wildland fire size distributions.
+This new wildland fire spatial database was used to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.
 
@@
 ```
-
 ## Climate
 
-Spatially explicit mean annual data on contemporary temperature and rainfall for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
+Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
 Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5 (Figure S5).
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands."}
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
+Data from Giambelluca \\textit{et al.} (2013) and Giambelluca \\textit{et al.} (2014)."}
 knitr::include_graphics("./fig_images/figS3_CurrentClimate.png")
 ```
-
 ```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by end of century under RCPs 4.5 and 8.5 (top panels) based on statistical downscaling of an ensemble of climate models (CMIP5)."}
@@
 ```
-
 ## Change in State Class Area
 
-Projections of total land area covered by each State Class were summed by year across the seven main Hawaiian Islands from LUCAS model output data.
+Projections of total land area covered by each State Class were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands from LUCAS model output data.
 
@@
 ```
-
 ## References

### 2021-02-19 Manuscript/fig_code/HI_CarbonBalance.R

@@ -175,7 +175,7 @@ draw_label(expression(Net~biome~productivity~(Tg~C~y^{-1})),
 # save two panel NBP fig as .png file
 ggsave("fig4_NBPpanels.png", width = 7, height = 6.5)
 
-# calculate cumulative ∆NBP when swticking land use & climate scenarios
+# calculate cumulative ∆NBP when swiching land use & climate scenarios
 dcumNBP <- cbal %>%
 	group_by(Scenario, Iteration) %>%
 	summarize(cumNBP = sum(NBP)/1000) %>%

### 2021-02-19 Manuscript/fig_code/HI_CarbonStocks.R

@@ -9,7 +9,7 @@ library(dplyr)
 library(ggplot2)
 library(cowplot)
 
-## Download LUCAS model output data from USGS ScienceBase: 
+## Access LUCAS model output data from USGS ScienceBase: 
 ## https://doi.org/10.5066/P9AWLFKZ
 
 # read Hawaii LUCAS model C stock output data into R

### 2021-02-19 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C).
-Notably,
+Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5).
+Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.
 

### 2021-02-23 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ### *4.1 Impact of different climate and land-use pathways* 
-By comparing model results of Hawaiian ecosystem carbon balance under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5).
+By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5).
 Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period.

### 2021-02-23 Manuscript/HI_Model_supp.Rmd
updated manuscript files

@@
 
-This study used a new spatial database of wildland fire perimeters in the Hawaiian Islands for the years 1999-2019 using data compiled from the Monitoring Trends in Burn Severity (MTBS) program, the Hawaiʻi Wildfire Management Organization, the U.S.
-National Park Service, the Oʻahu Army Natural Resources Program, and independent remote sensing analyses using publicly available LandSat data.
-This new wildland fire spatial database was used to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.
+This study used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.
+This new wildland fire spatial database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr.
+Clay Trauernicht in the Department of Natural Resources and Environmental Management at the University of Hawaiʻi at Mānoa.
+The goal of this effort was to locate and map all fires greater than or equal to 20 hectares, but some smaller fires were included as detected in imagery.
+Fire perimeters for the years 2002-2011 were primarily from the U.S.
+Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
+Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters primarily from the Kona and Kohala regions of Hawaiʻi Island.
+The U.S.
+National Park Service provided ground-based, GPS-mapped fires records from Hawaiʻi Volcanoes National Park.
+The Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fires on Oʻahu.
+All other fires were mapped directly by Dr.
+Trauernicht using LANDSAT and Sentinel-2 satellite imagery.
+For 2002-2011, USGS MTBS data were prioritized in the case of any duplicate record.
+Dr.
+Trauernicht mapped all fires from 2012 onwards in addition to filling missing records, where possible, using LANDSAT or Sentinel-2 imagery.
 

### 2021-03-03 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
+\pagebreak
+
 ```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of carbon dioxide fertilization in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing and high land use scenario.
@@
 ### *4.1 Impact of different climate and land-use pathways* 
-By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5).
-Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period.
+By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and wildland fire (figure 5).
+Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest impact, resulting in a median cumulative increase of 26 Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period.
 Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions.
-Switching from the ‘high’ to the ‘low’ land‐use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100.
-The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C).
+Switching from the high to the low land use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100.
+The combination of global climate mitigation and local reductions in land use conversion had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C).
 Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5).
 Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.
+These results demonstrate that reducing ecosystem carbon losses from land use change, harvest, and wildland fire can be an important component of greenhouse gas reduction efforts by sub-national jurisdictions like the State of Hawaiʻi, regardless of the global emissions trajectory.
+These results also highlight the utility of Hawaiʻi's multi-pronged approach of participating in global climate mitigation efforts by reducing emissions from the energy and transportation sectors while also reducing land-use emissions to minimize positive feedbacks to the climate system.
 
@@
 Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
-In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period {0.314 Tg C y^-1^; @selmants_baseline_2017].
-Our current model estimates of NPP and R~h~ being 9.5% and 15% lower, respectively [
+In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017], despite our current model estimates of contemporary NPP and R~h~ being 9.5% and 15% lower, respectively.
 

### 2021-03-09 Manuscript/HI_Model_ms.Rmd
updated manuscript files and deleted MOD17 files

@@
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land-use scenarios (low and high) over a 90-year timespan from 2010-2100.
-Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land-use change further enhancing carbon sink strength.
-In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land-use accelerating this transition and exacerbating net carbon loss.
+We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
+Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength.
+In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss.
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
@@
 Incorporating the interactive effects of land use and climate into spatially explicit future projections of ecosystem carbon balance could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
-Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future land-use and climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
+Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future land use and climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
@@
 These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017].
-Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@osher_effect_2003; @asner_high-resolution_2011].
+Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land use change [@osher_effect_2003; @asner_high-resolution_2011].
 Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
 Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
-Although these past trends surely inform the future impact of climate and land-use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts.
-To date only one study has attempted to integrate land-use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land-use change scenario and moderate radiative forcing [SRES A1B, equivalent to RCP 6; @selmants_baseline_2017].
+Although these past trends surely inform the future impact of climate and land use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts.
+To date only one study has attempted to integrate land use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land use change scenario and moderate radiative forcing [SRES A1B, equivalent to RCP 6; @selmants_baseline_2017].
 
 We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-We explored four unique scenarios that represent all combinations of two land-use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
+We explored four unique scenarios that represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
@@
 ## 2. Methods
-We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
+We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
 We also developed a separate set of scenarios to test model sensitivity to different levels of a CFE on NPP.
@@
 ### *2.2 States and transitions*
-We developed two land-use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
+We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
 Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
@@
  Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
-For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
-For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
-Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
+For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
+For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
+Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
 We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population.
-The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
-Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016].
-In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
+The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
+Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016].
+In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
 The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
@@
 Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
-Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
-The high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a).
+Wildfire in the low land use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019.
+The high land use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a).
 
@@
 Carbon stocks and fluxes for the seven main Hawaiian Islands were calculated for each scenario by summing within each Monte Carlo realization on an annual basis and then calculating annual means as well as the annual upper and lower limits of the 30 Monte Carlo realizations.
-Carbon balance for the seven main Hawaiian Islands was calculated on annual basis for each scenario and Monte Carlo realization as net biome productivity (NBP), which was equal to annual carbon input in the form of NPP minus the annual sum of all carbon losses from terrestrial ecosystems, including heterotrophic respiration (R~h~) from litter and soil, carbon fluxes to the atmosphere triggered by land use and land-use change, wildfire emissions, and aquatic carbon losses through leaching and overland flow.
+Carbon balance for the seven main Hawaiian Islands was calculated on annual basis for each scenario and Monte Carlo realization as net biome productivity (NBP), which was equal to annual carbon input in the form of NPP minus the annual sum of all carbon losses from terrestrial ecosystems, including heterotrophic respiration (R~h~) from litter and soil, carbon fluxes to the atmosphere triggered by land use and land use change, wildfire emissions, and aquatic carbon losses through leaching and overland flow.
 Positive NBP values indicated ecosystems of the seven main Hawaiian Islands were acting as a net sink for atmospheric CO~2~, while negative NBP values indicated that these ecosystems were acting as a net carbon source to the atmosphere [@chapin_reconciling_2006].
@@
 Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010 (figure 2a), with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; figure 2b).
-Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land-use change.
+Ecosystems accumulated carbon in all scenarios but at different rates, with trajectories shaped primarily by climate change and to a lesser extent by land use change.
 The highest and most consistent projected accumulation of ecosystem carbon occured under the combination of low radiative forcing and low land use change, yielding a ~15% increase in ecosystem carbon to an average of 363 Tg by 2100 (figure 2a).
@@
 3).
-The combination of high radiative forcing (RCP 8.5) and high land-use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land-use scenario (supplmental figure 5).
+The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land use scenario (supplmental figure 5).
 In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplmental figure 4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
-Heterotrophic respiration declined substantially over time in the high land-use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 5), similar to trends in NPP.
-Transition-triggered carbon fluxes to the atmosphere from land use, land-use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land-use scenario and ~0.2 Tg y^-1^ in the low land-use scenario (figure 3).
-Uncertainty around transition-triggered carbon fluxes were higher in the high land-use scenario, driven primarily by greater variability in wildland fire probabilities.
+Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 5), similar to trends in NPP.
+Transition-triggered carbon fluxes to the atmosphere from land use, land use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land use scenario and ~0.2 Tg y^-1^ in the low land use scenario (figure 3).
+Uncertainty around transition-triggered carbon fluxes were higher in the high land use scenario, driven primarily by greater variability in wildland fire probabilities.
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net primary production (NPP), heterotrophic respiration (Rh) and carbon fluxes induced by land use and land-use change for the seven main Hawaiian Islands.
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net primary production (NPP), heterotrophic respiration (Rh) and carbon fluxes induced by land use and land use change for the seven main Hawaiian Islands.
 Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
@@
 Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined over time in all four scenarios (figure 4).
-On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under the RCP 4.5 radiative forcing scenario, but carbon sink strength was ~40% lower in the high land-use scenario compared to the low land-use scenario by the end of the simulation period (figure 4).
-In contrast, ecosystems of the Hawaiian Islands acted as a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, with the transition from sink to source occuring 15 years earlier on average in the high land-use scenario than in the low land-use scenario (figure 4).
-The high land-use scenario under RCP 8.5 represented a ~40% larger net source of carbon to the atmosphere by the year 2100 than the low-land use scenario under the same radiative forcing.
+On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under the RCP 4.5 radiative forcing scenario, but carbon sink strength was ~40% lower in the high land use scenario compared to the low land use scenario by the end of the simulation period (figure 4).
+In contrast, ecosystems of the Hawaiian Islands acted as a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, with the transition from sink to source occuring 15 years earlier on average in the high land use scenario than in the low land use scenario (figure 4).
+The high land use scenario under RCP 8.5 represented a ~40% larger net source of carbon to the atmosphere by the year 2100 than the low-land use scenario under the same radiative forcing.
 Over the entire simulation period, both global emissions reductions and local avoided land conversion resulted in substantial increases in cumulative NBP (figure 5).
-However, switching from RCP 8.5 to RCP 4.5 increased cumulative NBP in the Hawaiian Islands more than twice as much as reducing emissions from local land-use change and wildfire disturbance (figure 5).
-Switching from RCP 8.5 to RCP 4.5 under the low land-use scenario yielded the greatest cumulative increase in NBP, resulting in a median gain of 26.5 Tg of carbon over the entire 90-year simulation period.
+However, switching from RCP 8.5 to RCP 4.5 increased cumulative NBP in the Hawaiian Islands more than twice as much as reducing emissions from local land use change and wildfire disturbance (figure 5).
+Switching from RCP 8.5 to RCP 4.5 under the low land use scenario yielded the greatest cumulative increase in NBP, resulting in a median gain of 26.5 Tg of carbon over the entire 90-year simulation period.
 
@@
 ```
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from the high to low land-use change scenario under each radiative forcing scenario (top panel) and when switching from the high (RCP 8.5) to low (RCP 4.5) radiative forcing scenario under each land-use scenario (bottom panel).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in cumulative net biome productivity (NBP) for the seven main Hawaiian Islands when switching from the high to low land use change scenario under each radiative forcing scenario (top panel) and when switching from the high (RCP 8.5) to low (RCP 4.5) radiative forcing scenario under each land use scenario (bottom panel).
 Box plots indicate the median (vertical black line), 25th and 75th percentiles (colored boxes), 10th and 90th percentiles (thin horizontal lines), and values outside of this range (black circles).
@@
 Projected estimates of both total ecosystem carbon storage and ecosystem carbon balance were highly sensitive to differing rates of a CFE on plant productivity.
-Under the high radiative forcing (RCP 8.5) and high land-use scenario, the inclusion of a CFE ranging from 5-15% led to ~33-98 Tg of additional carbon storage in ecosystems by the end of the century, a ~10-30% increase (figure 6a).
+Under the high radiative forcing (RCP 8.5) and high land use scenario, the inclusion of a CFE ranging from 5-15% led to ~33-98 Tg of additional carbon storage in ecosystems by the end of the century, a ~10-30% increase (figure 6a).
 Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian Island ecosystems from a net carbon source to the atmosphere during the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
 Net carbon sink strength was further enhanced at higher CFE rates, with NBP increasing by an average of 0.07 Tg C y^-1^ for each 1% increase in CFE (figure 6b).
-When compared to other scenarios, applying a 5% CFE to the high radiative forcing and high land-use scenario resulted in a mean annual NBP of 0.46 ± 0.3 Tg C y^-1^, roughly equivalent to mean annual NBP in the low radiative forcing and low land-use scenario with no CFE (0.52 ± 0.12).
-A 15% CFE applied to the high radiative forcing and high land-use scenario resulted in a mean annual NBP of 1.18 ± 0.29 Tg C y^-1^, more than double that of the low radiative forcing and low land-use scenario with no CFE.
+When compared to other scenarios, applying a 5% CFE to the high radiative forcing and high land use scenario resulted in a mean annual NBP of 0.46 ± 0.3 Tg C y^-1^, roughly equivalent to mean annual NBP in the low radiative forcing and low land use scenario with no CFE (0.52 ± 0.12).
+A 15% CFE applied to the high radiative forcing and high land use scenario resulted in a mean annual NBP of 1.18 ± 0.29 Tg C y^-1^, more than double that of the low radiative forcing and low land use scenario with no CFE.
 
@@
 Future projections indicate Hawaiian terrestrial ecosystems will continue to be a net sink for atmospheric carbon if global CO~2~ emissions peak around 2040 and then decline (RCP 4.5), and that carbon sink strength can be further enhanced by reducing the intensity and extent of future land use change.
-If, however, global CO~2~ emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high levels of land-use change accelerating this transition and exacerbating net carbon loss.
+If, however, global CO~2~ emissions continue to rise throughout the 21st century (RCP 8.5), our projections indicate Hawaiian ecosystems will transition from a net sink to a net source of CO~2~ to the atmosphere, with high levels of land use change accelerating this transition and exacerbating net carbon loss.
 Our model results also indicate that projections of ecosystem carbon balance are highly sensitive to the introduction of a CFE.
@@
 
-### *4.1 Impact of different climate and land-use pathways* 
+### *4.1 Impact of different climate and land use pathways* 
 By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and wildland fire (figure 5).
 Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest impact, resulting in a median cumulative increase of 26 Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period.
-Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions.
+Long-term reductions in the intensity of land use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions.
 Switching from the high to the low land use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100.
 The combination of global climate mitigation and local reductions in land use conversion had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C).
-Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5).
-Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.
+Notably, the relative impact of reducing emissions from land use change was much greater under the high radiative forcing pathway (RCP 8.5).
+Cumulative NBP increased by 130% when switching from the high to low land use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land use under RCP 4.5.
 These results demonstrate that reducing ecosystem carbon losses from land use change, harvest, and wildland fire can be an important component of greenhouse gas reduction efforts by sub-national jurisdictions like the State of Hawaiʻi, regardless of the global emissions trajectory.
-These results also highlight the utility of Hawaiʻi's multi-pronged approach of participating in global climate mitigation efforts by reducing emissions from the energy and transportation sectors while also reducing land-use emissions to minimize positive feedbacks to the climate system.
+These results also highlight the utility of Hawaiʻi's multi-pronged approach of participating in global climate mitigation efforts by reducing emissions from the energy and transportation sectors while also reducing land use emissions to minimize positive feedbacks to the climate system.
 
@@
 Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting in Hawaiʻi, and Caroline Conrad for help with Google Earth Engine scripting.
+for information on eucalyptus harvesting in Hawaiʻi.
 Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.

### 2021-03-09 MOD17_NPP/code/MOD17_processing.R

@@ -1,39 +1,0 @@
-## Calculate annual statewide NPP 
-## for the years 2010-2019
-## using MODIS MOD17 NPP estimates 
-## Paul C. Selmants
-## 2021-01-25 (ISO 8601)
-
-## R version 4.0.2
-
-# load required packages into R
-library(raster) #version 3.3-13
-library(rgdal) #version 1.4-16
-library(tidyr, warn.conflicts = FALSE) #version 1.1.1
-library(dplyr, warn.conflicts = FALSE) #version 1.0.2
-library(ggplot2) #version 3.3.2
-
-## Set working directory at highest level in HI_Model
-## GitHub repository, https://github.com/selmants/HI_Model 
-
-# create list of annual MOD17 NPP .tif files
-npplist <- list.files('./MOD17_NPP/data', full.names = TRUE)
-#create raster stack of annual NPP .tif files
-stack <- stack(npplist)
-# multiply all values by MODIS scale factor (0.0001)	
-# to get NPP values in kg C m-2 y-1
-npp <- stack * 1e-04
-
-# export pixel NPP data to dataframe, 
-# pivot to long format, and sum values to get 
-# statewide NPP estimates in Tg C y-1
-npp_data <- as.data.frame(npp, xy=TRUE, na.rm = TRUE) %>%
-	pivot_longer(!x:y, names_to = "year", values_to = "npp") %>%
-	group_by(year) %>%
-	summarise(NPP_Tgy = (sum(npp)*2.5e5)*1e-9) %>%
-	mutate(Year = as.numeric(gsub("hawaii_npp_", "", year))) %>%
-	select(Year, NPP_Tgy) %>%
-	as.data.frame() 
-
-
-

### 2021-03-09 CarbonStockFlow_IBIS/code/HI_FlowPathways_Baseline.R
updated FlowPathways R script to removeunused SOC_Input calculations

@@ -63,18 +63,6 @@ litterfall <- rawlitc %>%
 		FlowTypeID = 'Litterfall') %>%
 	filter(FromStateClassID != 'Agriculture:All') %>%
 	as.data.frame(.)
-#calculate multipliers for soil CO2 emission as proportion of 
-#current year litter decay (SOC Input)
-soilemiss <- flux %>%
-	filter(Flux == 'soc2co2') %>% 
-	mutate(turnover = Value/lit2soc$Value) %>%
-	group_by(FromStratumID, FromStateClassID) %>%
-	summarize(Multiplier = mean(turnover)) %>%
-	mutate(FromStockTypeID = 'Soil',
-		ToStockTypeID = 'Atmosphere',
-		StateAttributeTypeID = 'SOC Input',
-		FlowTypeID = 'Emission (soil)') %>%
-	as.data.frame(.)
 
 #calculate soil CO2 emission multipliers as 
 # a proportion of current year NPP
@@ -87,19 +75,6 @@ soilemissNPP <- flux %>%
 		ToStockTypeID = 'Atmosphere',
 		StateAttributeTypeID = 'NPP',
 		FlowTypeID = 'Emission (soil)') %>%
-	as.data.frame(.)
-
-#calculate multipliers for leaching as proportion 
-#of current year litter decay (SOC Input)
-leaching <- flux %>%
-	filter(Flux == 'yrleach') %>%
-	mutate(turnover = Value/(lit2soc$Value)) %>%
-	group_by(FromStratumID, FromStateClassID) %>%
-	summarize(Multiplier = mean(turnover)) %>%
-	mutate(FromStockTypeID = 'Soil',
-		ToStockTypeID = 'Aquatic',
-		StateAttributeTypeID = 'SOC Input',
-		FlowTypeID = 'Leaching') %>%
 	as.data.frame(.)
 
 # calculate leaching multipliers as

### 2021-03-10 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
- Carbon flow rates for each state type and moisture zone were estimated as the ratio of the IBIS-derived flux to the size of the originating carbon stock at each age [@sleeter_effects_2018].
-A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate.
-This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
+ For each combination of moisture zone and state type, IBIS-derived NPP values were adjusted using a spatially explicit stationary growth multiplier applied to each simulation cell to reflect local variations in NPP driven by microclimate.
+We used empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] to calculate this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and state type [@sleeter_effects_2019].
+Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell.
+All other carbon flow rates were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018].
 Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
@@
 Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
-In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017], despite our current model estimates of contemporary NPP and R~h~ being 9.5% and 15% lower, respectively.
+In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017].
+This discrepancy was likely driven by modifications in how we calculated per-pixel values of NPP, soil Rh, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well as the finer spatial resolution of this model run [250-m vs.
+1-km; @selmants_baseline_2017; @daniel_integrating_2018].
+
+our current model estimates of contemporary statewide NPP and R~h~ were 9.5% and 15% lower, respectively.
 

### 2021-03-11 Manuscript/HI_Model_ms.Rmd
updated manuscript files

--- 2021-03-09 Manuscript/HI_Model_ms.Rmd
@@
 
- Carbon flow rates for each state type and moisture zone were estimated as the ratio of the IBIS-derived flux to the size of the originating carbon stock at each age [@sleeter_effects_2018].
-A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate.
-This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008].
+ Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and state type adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [@sleeter_effects_2019].
+We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and state type based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
+Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell.
+All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
 Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
@@
 Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
-In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017], despite our current model estimates of contemporary NPP and R~h~ being 9.5% and 15% lower, respectively.
+In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.341 Tg C y^-1^; @selmants_baseline_2017].
+This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our modelʻs finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018].
+Previous versions of a Hawaiʻi LUCAS model were run at 1-km spatial resolution and simulation cells within each unique combination of moisture zone and state type all had the same mean IBIS-derived NPP value applied to them at the beginning of the simulation period.
+In contrast, our NPP estimates at 250-m spatial resolution were adjusted on a cell-by-cell basis using Hawaiʻi-specific climate data as described in section 2.3.
+As a result, our statewide NPP estimates from 2011-2020 were 9.5% lower on average than previous LUCAS model estimates for Hawaiʻi during the same time period [@selmants_baseline_2017], likely because of the greater influence of more arid simulation cells.
+Soil carbon losses via R~h~, leaching, and overland flow in previous versions of the LUCAS model were calculated as the ratio of the IBIS-derived flux to the size of the originating carbon stock, in this case soil organic carbon to 1-m depth [@daniel_integrating_2018].
+Here we calculated soil R~h~ and aquatic carbon losses as the ratio of the mean IBIS-derived flux to the microclimate-adjusted NPP value of each simulation cell, which is a more realistic driver than stock size [@jackson_ecology_2017].
+Compared to previous Hawaiʻi LUCAS model estimates [@selmants_baseline_2017], soil R~h~ and aquatic carbon losses from 2011-2019 were reduced by an average of 15% and 21%, respectively, which widened the gap between between carbon gain (NPP) and carbon losses and accounted for the overall increase in NBP estimates for this time period.
 
@@
 Geological Survey Biological Carbon Sequestration Program.
-Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, Nicholas Koch of Forest Solutions, Inc.
-for information on eucalyptus harvesting in Hawaiʻi.
-Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible.
+Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, and to Nicholas Koch of Forest Solutions, Inc.
+for information on _Eucalyptus_ spp.
+harvesting in Hawaiʻi.
+Thanks also to Christian Giardina and Zhiliang Zhu for providing the initial impetus for this research.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.

### 2021-03-12 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019].
-There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
-Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
+There is increasing recognition among policymakers that natural and agricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
+Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-efficient and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].
 However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.
@@
 We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-We explored four unique scenarios that represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and 8.5).
+We explored four unique scenarios that represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
@@
 ### *2.4 CO~2~ fertilization effect*
-Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well [@Walker_integrating_2020].
-However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans [@Walker_integrating_2020].
+Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well [@franks_sensitivity_2013].
+However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans [@walker_integrating_2020].
 Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model projections of ecosystem carbon balance to different rates of a CO~2~ fertilization effect (CFE).
@@
 There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.
-Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
+Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 Tg C in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019].
 In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.341 Tg C y^-1^; @selmants_baseline_2017].
@@
 Here we calculated soil R~h~ and aquatic carbon losses as the ratio of the mean IBIS-derived flux to the microclimate-adjusted NPP value of each simulation cell, which is a more realistic driver than stock size [@jackson_ecology_2017].
-Compared to previous Hawaiʻi LUCAS model estimates [@selmants_baseline_2017], soil R~h~ and aquatic carbon losses from 2011-2019 were reduced by an average of 15% and 21%, respectively, which widened the gap between between carbon gain (NPP) and carbon losses and accounted for the overall increase in NBP estimates for this time period.
+Compared to previous Hawaiʻi LUCAS model estimates [@selmants_baseline_2017], soil R~h~ and aquatic carbon losses from 2011-2020 were reduced by an average of 15% and 21%, respectively, which widened the gap between between carbon gain (NPP) and carbon losses and accounted for the overall increase in NBP estimates for this time period.
 
 ### *4.3 Limitations of this study* 
+There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect across climatic and nutrient availability gradients remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020].
+Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in the strength of a CFE.
+With no CFE, Hawaiian ecosystems became a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
+However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems collectively remaining a net carbon sink throughout the entire simulation period.
+A CFE of 15% applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.
 

### 2021-03-15 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 ### *4.3 Limitations of this study* 
-There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect across climatic and nutrient availability gradients remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020].
-Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in the strength of a CFE.
-With no CFE, Hawaiian ecosystems became a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
-However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems collectively remaining a net carbon sink throughout the entire simulation period.
-A CFE of 15% applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.
+There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020].
+Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength.
+With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
+However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period.
+A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.
+Despite this demonstrated sensitivity to a CFE, the combined effect of several potentially attenuating factors makes it difficult to determine the appropriate magnitude of a CFE with any degree of confidence [@walker_integrating_2020].
+Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020].
+Forest age may also be a factor, with young aggrading forests showing a strong positive growth response to CO~2~ fertilization [@walker_decadal_2019], while old-growth forests show little to no response [@jiang_fate_2020; @yang_low_2020].
+This evidence indicates that a CFE may be highly variable across space and time, suggesting that it may be unrealistic to apply a single CFE rate value across an entire region over several decades.
+Until mechanistic understanding is improved, the most conservative approach when projecting future ecosystem balance in the context of climate mitigation planning may be to assume no CFE, with the knowledge that any realized CFE will only enhance ecosystem carbon sequestration.
 

### 2021-03-15 Manuscript/HI_Model_ms.Rmd
updated manuscript and supplementary files

@@
 ### *4.3 Limitations of this study* 
-There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020].
+There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain [@franks_sensitivity_2013; @walker_integrating_2020].
 Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength.
 With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
-However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period.
+However, a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period.
 A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.

### 2021-03-15 Manuscript/HI_Model_supp.Rmd
updated manuscript and supplementary files

@@
 ```
+
+\pagebreak
+
 ## Climate
@@
 Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
-Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5 (Figure S5).
+Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 (Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5.
 
@@
 
-Projections of total land area covered by each State Class were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands from LUCAS model output data.
+Projections of total land area covered by each of five State Classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios  (low and high; Figure S5).
 
@@
 ```
+
+\pagebreak
+
 ## References

### 2021-03-15 Manuscript/fig_code/HI_FireEmissions.R

@@ -24,3 +24,11 @@ fire <- read.csv("TransitHarvFlows.csv", stringsAsFactors = FALSE) %>%
 		select(Scenario, Timestep, Iteration, Island = FromSecondaryStratumID,
 			fire_emiss)
 
+# Summarize mean annual fire emissions statewide 
+fire_state <- fire %>%
+	group_by(Scenario, Timestep, Iteration) %>%
+	summarize(emiss = sum(fire_emiss)) %>%
+	group_by(Scenario, Timestep) %>%
+	summarize(mean_emiss = mean(emiss),
+		min_emiss = min(emiss),
+		max_emiss = max(emiss))

### 2021-03-17 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 
+Our model does not currently differentiate between forests dominated by native versus non-native tree species, which could influence estimates of ecosystem carbon balance.
+Native forests in the Hawaiian Islands are dominated by ʻōhiʻa (_Metrosideros polymorpha_ Gauditch), an endemic foundational tree species found across a broad range of climatic and edaphic conditions [@ziegler_hawaiian_2002].
+Beginning in 2010, a fungal disease termed Rapid ʻŌhiʻa Death (ROD) caused by two _Ceratocystus_ spp.
+has emerged and caused widespread mortality to mature ʻōhiʻa trees across a range of size classes, primarily on Hawaiʻi Island [@mortenson_assessing_2016; @fortini_evolving_2019].
+The distribution and potential range of this emerging threat to Hawaiʻi's most widespread native tree species has only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
+ROD-affected forest canopies could also potentially see a replacement of long-lived, stress-tolerating ʻōhiʻa with non-native tree species that have faster growth rates, lower wood density, and faster tissue turnover, all of which could influence the long-term trajectory of forest carbon cycling in Hawaiʻi.
+New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.
+
 ## 5. Conclusion

### 2021-03-19 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength.
-With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
-However, a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period.
+With no CFE, Hawaiian ecosystems became a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario.
+However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period.
 A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.
-Despite this demonstrated sensitivity to a CFE, the combined effect of several potentially attenuating factors makes it difficult to determine the appropriate magnitude of a CFE with any degree of confidence [@walker_integrating_2020].
+Despite this demonstrated sensitivity to a CFE, several potentially attenuating factors make complicate selecting a realistic CFE value with any degree of confidence [@walker_integrating_2020].
 Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020].
@@
 has emerged and caused widespread mortality to mature ʻōhiʻa trees across a range of size classes, primarily on Hawaiʻi Island [@mortenson_assessing_2016; @fortini_evolving_2019].
-The distribution and potential range of this emerging threat to Hawaiʻi's most widespread native tree species has only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
-ROD-affected forest canopies could also potentially see a replacement of long-lived, stress-tolerating ʻōhiʻa with non-native tree species that have faster growth rates, lower wood density, and faster tissue turnover, all of which could influence the long-term trajectory of forest carbon cycling in Hawaiʻi.
+The distribution and potential range of this emerging threat to Hawaiʻi's dominant native tree species have only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
+ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, which could alter the long-term trajectory of forest carbon cycling in Hawaiʻi.
 New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.
+
+Interannual climate variability is a primary factor influencing spatial and temporal patterns in global wildland fire activity [@abatzoglou_global_2018], and climate warming is expected to increase wildland fire frequency and wildfire season length across a wide range of biomes [@sun_global_2019].
+Although our model projections capture the spatial and temporal variation in ignition probability and fire extent by sampling from previous fire years (1999-2019), we did not incorporate how projected future climate change under different RCPs will affect the frequency and extent of wildland fire.
+Recent fire probability modeling for the northwest portion of Hawaiʻi Island indicated that projected drying and warming trends under RCP 8.5 could increase maximum fire probability values more than three-fold and shift areas of peak flammability to higher elevation by mid-century [@trauernicht_vegetationrainfall_2019].
+Extending this probability fire modeling approach statewide would provide a quantitative, spatially explicit assessment of wildland fire probability for the main Hawaiian Islands as predicted by climate, land cover, and ignition density, which is highly correlated with population density [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019].
+This approach would provide future simulation model projections of ecosystem carbon balance with more realistic scenarios of expected annual area burned based on the integrated effects of future climate and land use change.
 
@@
 Government.
+
+\pagebreak
 

### 2021-03-22 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.
-Despite this demonstrated sensitivity to a CFE, several potentially attenuating factors make complicate selecting a realistic CFE value with any degree of confidence [@walker_integrating_2020].
+Despite this demonstrated sensitivity to a CFE, several potentially attenuating factors complicate the selection of a realistic CFE value with any degree of confidence [@walker_integrating_2020].
 Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020].
 Forest age may also be a factor, with young aggrading forests showing a strong positive growth response to CO~2~ fertilization [@walker_decadal_2019], while old-growth forests show little to no response [@jiang_fate_2020; @yang_low_2020].
-This evidence indicates that a CFE may be highly variable across space and time, suggesting that it may be unrealistic to apply a single CFE rate value across an entire region over several decades.
+This evidence indicates that a CFE may be highly variable across space and time, suggesting it may be unrealistic to apply a single CFE rate value across an entire region over several decades.
 Until mechanistic understanding is improved, the most conservative approach when projecting future ecosystem balance in the context of climate mitigation planning may be to assume no CFE, with the knowledge that any realized CFE will only enhance ecosystem carbon sequestration.
@@
 The distribution and potential range of this emerging threat to Hawaiʻi's dominant native tree species have only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
-ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, which could alter the long-term trajectory of forest carbon cycling in Hawaiʻi.
+ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, potentially alter the long-term trajectory of forest carbon cycling in Hawaiʻi.
 New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.
 
-Interannual climate variability is a primary factor influencing spatial and temporal patterns in global wildland fire activity [@abatzoglou_global_2018], and climate warming is expected to increase wildland fire frequency and wildfire season length across a wide range of biomes [@sun_global_2019].
-Although our model projections capture the spatial and temporal variation in ignition probability and fire extent by sampling from previous fire years (1999-2019), we did not incorporate how projected future climate change under different RCPs will affect the frequency and extent of wildland fire.
+Interannual climate variability is a primary factor influencing spatial and temporal patterns of global wildland fire activity [@abatzoglou_global_2018], with climate warming expected to increase wildland fire frequency and wildfire season length across a wide range of biomes [@sun_global_2019].
+Although our model projections capture the spatial and temporal variation in ignition probability and fire extent by sampling from previous fire years (1999-2019), we did not incorporate how projected future climate change will affect the frequency and extent of wildland fire.
 Recent fire probability modeling for the northwest portion of Hawaiʻi Island indicated that projected drying and warming trends under RCP 8.5 could increase maximum fire probability values more than three-fold and shift areas of peak flammability to higher elevation by mid-century [@trauernicht_vegetationrainfall_2019].
 Extending this probability fire modeling approach statewide would provide a quantitative, spatially explicit assessment of wildland fire probability for the main Hawaiian Islands as predicted by climate, land cover, and ignition density, which is highly correlated with population density [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019].
-This approach would provide future simulation model projections of ecosystem carbon balance with more realistic scenarios of expected annual area burned based on the integrated effects of future climate and land use change.
+This approach would provide future simulation model projections of Hawaiian ecosystem carbon balance with more realistic scenarios of expected annual area burned based on the integrated effects of future climate and land use change.
 
 ## 5. Conclusion
+
+Although terrestrial ecosystems are currently an important sink for atmospheric CO~2~, the future direction and magnitude of the land carbon sink are highly uncertain, especially at regional scales.
+Our simulation modeling results indicated that projected climate change, dictated by long-term trajectories in global greenhouse gas emissions, was the primary factor influencing terrestrial ecosystem carbon balance in the Hawaiian Islands.
+Long-term reductions in the intensity of land use change and wildland fire also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions.
+CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration.
+By incorporating the interactive effects of land use and climate change into future projections of ecosystem carbon balance, our model results could be used as a set of baseline projections for the State of Hawaiʻi to evaluate different ecosystem-based climate mitigation strategies.
+Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.
+These long-term projections will be critical to understanding how future land use and climate-biosphere feedbacks could influence the achievement of climate mitigation goals.
 
@@
 Government.
-
-\pagebreak
 
@@
 
+\pagebreak
+
 ## References

### 2021-03-23 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 \centering
-Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,
+Paul C. Selmants^1,7^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,
 
-Clay Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^
+Clay Trauernicht^3^, Abby G. Frazier^4,5^, Gregory P. Asner^6^
 
@@
 ^3^University of Hawaiʻi at Mānoa, Honolulu, HI, USA  
-^4^The East-West Center, Honolulu, HI, USA    
-^5^Arizona State University, Tempe, AZ, USA    
-^6^Author to whom correspondence should be addressed
+^4^East-West Center, Honolulu, HI, USA 
+^5^Clark University, Worcester, MA, USA   
+^6^Arizona State University, Tempe, AZ, USA    
+^7^Author to whom correspondence should be addressed
 
@@
 However, Hawaiʻi's challenging terrain complicates these assessment efforts.
-The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances (supplmental figure 3), with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013].
-Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
+The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~200-10,200 mm [@giambelluca_online_2013].
+Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008; @mckenzie_temperature_2019] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017].
 These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017].
 Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land use change [@osher_effect_2003; @asner_high-resolution_2011].
-Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
+Thousands of hectares were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
 Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
@@
 
- Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (supplmental figure 1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+ Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (supplemental figure 1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
@@
 All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
-Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
+Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on mid-century (2049-2069) and end-of-century (2070-2099) statistically downscaled CMIP5 temperature and rainfall projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
 The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
@@
 3).
-The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land use scenario (supplmental figure 5).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplmental figure 4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
+The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 5).
+In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
 Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 5), similar to trends in NPP.
@@
 This study was funded by the U.S.
-Geological Survey Biological Carbon Sequestration Program.
+Geological Survey Biological Carbon Sequestration Program and the Pacific Islands Climate Adaptation Science Center.
 Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, and to Nicholas Koch of Forest Solutions, Inc.

### 2021-03-23 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.
+Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.
 

### 2021-03-26 Manuscript/HI_Model_supp.Rmd
updated manuscript files
updated manuscript files, added journal cover letter

@@
 Areas with MAI values between zero and 1,661 were classified in the Mesic Moisture Zone.
-The MAI value of 1,661 is roughly equivalent to areas at 1,000m elevation that receive 2,500 mm of annual rainfall [@price_mz_2012].
+The MAI value of 1,661 is roughly equivalent to areas at 1,000 m elevation that receive 2,500 mm of annual rainfall [@price_mz_2012].
 Areas with MAI values that exceeded 1,661 were classified in the Wet Mesic Zone.
@@
 This study used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.
-This new wildland fire spatial database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr.
+This new database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr.
 Clay Trauernicht in the Department of Natural Resources and Environmental Management at the University of Hawaiʻi at Mānoa.
@@
 Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
-Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters primarily from the Kona and Kohala regions of Hawaiʻi Island.
+The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, primiarily in the Kona and Kohala regions.
 The U.S.
-National Park Service provided ground-based, GPS-mapped fires records from Hawaiʻi Volcanoes National Park.
-The Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fires on Oʻahu.
+National Park Service provided ground-based, GPS-mapped fire records from Hawaiʻi Volcanoes National Park.
+The Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fire records on Oʻahu.
 All other fires were mapped directly by Dr.
 Trauernicht using LANDSAT and Sentinel-2 satellite imagery.
-For 2002-2011, USGS MTBS data were prioritized in the case of any duplicate record.
-Dr.
-Trauernicht mapped all fires from 2012 onwards in addition to filling missing records, where possible, using LANDSAT or Sentinel-2 imagery.
+Data from USGS MTBS were prioritized for the years 2002-2011 in the case of duplicate records.
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
 The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
@@
 ```
-
 \pagebreak
@@
 
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
 Data from Giambelluca \\textit{et al.} (2013) and Giambelluca \\textit{et al.} (2014)."}
@@
 ```
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by end of century under RCPs 4.5 and 8.5 (top panels) based on statistical downscaling of an ensemble of climate models (CMIP5)."}
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by 2100 under RCP 4.5 and RCP 8.5 based on statistical downscaling of an ensemble of climate models (CMIP5)."}
 knitr::include_graphics("./fig_images/figS4_FutureClimate.png")
@@
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100.
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100.
 Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}

### 2021-03-26 Manuscript/coverletter/ERL_coverletter.Rmd
updated manuscript files, added journal cover letter

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
header-includes:
   - \linespread{1.05}
   - \newcommand{\blankline}{\quad\pagebreak[2]}
   - \usepackage{eso-pic}
   - \usepackage{graphicx}
   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_letterhead.pdf} \vfill }}}
geometry: top=2in, left=1.75in, bottom=1in, right=1in
mainfont: Times New Roman
fontsize: 12pt
urlcolor: blue
---

\AddToShipoutPicture*{\BackgroundPic}
\pagenumbering{gobble}

`r format(Sys.time(), '%B %d, %Y')`

\blankline

Dear Editors:

\blankline

Please consider our article, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. Hawaiʻi ellentesque non pretium turpis. Etiam nibh nibh, pharetra vel semper quis, rhoncus non sapien. Nunc gravida, elit nec malesuada elementum, purus orci consequat lacus, et gravida felis felis eu lacus. Aenean ac dapibus purus, eget consequat eros. Sed eros odio, rutrum vitae posuere id, facilisis eu elit. Duis sed lorem et tortor consectetur mattis. In nisi arcu, aliquet a purus in, sodales pulvinar sapien. Nulla auctor, orci id sodales consequat, augue odio egestas libero, tincidunt fringilla eros neque at orci. Duis accumsan tellus sapien, id cursus odio egestas in. Integer in nibh suscipit mauris auctor gravida.

Aenean dapibus placerat laoreet. Maecenas id arcu semper, feugiat urna eget, placerat ipsum. Maecenas porta sit amet ante vel finibus. Quisque porttitor urna gravida commodo ullamcorper. Nullam dapibus sodales lacus vitae facilisis. Praesent nec velit eget quam congue aliquet vitae sed urna. Duis venenatis ornare sagittis. Nullam dolor odio, porttitor pulvinar aliquet in, rutrum id dolor. Nullam vestibulum ligula diam, at fringilla tellus aliquet eu. In sit amet magna ac ante efficitur laoreet ut eget lectus. Aenean mauris quam, accumsan fermentum massa ac, blandit hendrerit arcu. Proin dignissim sapien quis maximus faucibus. Aenean lobortis tellus nec nulla fringilla, vitae sagittis felis pretium. Pellentesque nec luctus urna.

Cras velit neque, pulvinar vitae eros eget, commodo dignissim mauris. Mauris ac facilisis tortor. Mauris imperdiet diam quis justo posuere convallis non et felis. Donec vel justo orci. In et volutpat est. Integer lacinia faucibus tellus, non vehicula enim faucibus eget.

\blankline

Best regards,

```{r echo=FALSE, out.width='35%'}
knitr::include_graphics('pcs_signature.png')
```

\raggedright
Paul C. Selmants 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

### 2021-03-26 Manuscript/HI_Model_supp.Rmd
updated manuscript files

@@
 ---
-title: "Supplemental material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
 bibliography: HI_Model.bib
@@
 ```
-\centering
-Paul C. Selmants^1,6^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
-Parker C. Trauernicht^3^, Abby G. Frazier^4^, Gregory P. Asner^5^
+\raggedright
+## Supplementary Material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change
+`r format(Sys.time(), '%B %d, %Y')`
 
-\raggedright
-**Affiliations:**  
-^1^U.S. Geological Survey, Moffett Field, CA, USA   
-^2^U.S. Geological Survey, Seattle, WA, USA  
-^3^University of Hawaiʻi at Mānoa, Honolulu, HI, USA  
-^4^The East-West Center, Honolulu, HI, USA  
-^5^Arizona State University, Tempe, AZ, USA  
-^6^Author to whom correspondence should be addressed
-
-**Email:** pselmants@usgs.gov
-
-**Running title:** Hawaii carbon balance  
-**Keywords:** land use, climate change, carbon balance, Hawaii, scenarios, disturbance, ecosystem model
-
-**Date:** `r format(Sys.time(), '%B %d, %Y')`
-
-\pagebreak
-
-## Moisture Zones
+### Moisture Zones
 
@@
 These three zones were based on a moisture availability index (MAI), calculated as mean annual precipitation (MAP) minus potential evapotranspiration [PET; @price_mz_2012].
-Areas where the MAI was less than zero, i.e.
-where PET exceeded MAP, were classified in the Dry Moisture Zone.
+Areas where where MAI values were less than zero (i.e., where MAP < PET) were classified in the Dry Moisture Zone.
 Areas with MAI values between zero and 1,661 were classified in the Mesic Moisture Zone.
 The MAI value of 1,661 is roughly equivalent to areas at 1,000 m elevation that receive 2,500 mm of annual rainfall [@price_mz_2012].
-Areas with MAI values that exceeded 1,661 were classified in the Wet Mesic Zone.
+Areas with MAI values greater than 1,661 were classified in the Wet Moisture Zone.
 
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
 knitr::include_graphics("./fig_images/figS1_MZ.png")
 ```
-## Wildland Fire
+
+### Wildland Fire
 
@@
 This new database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr.
-Clay Trauernicht in the Department of Natural Resources and Environmental Management at the University of Hawaiʻi at Mānoa.
+Clay Trauernicht (Department of Natural Resources and Environmental Management, University of Hawaiʻi at Mānoa).
 The goal of this effort was to locate and map all fires greater than or equal to 20 hectares, but some smaller fires were included as detected in imagery.
@@
 Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
-The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, primiarily in the Kona and Kohala regions.
+The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, primarily in the Kona and Kohala regions.
 The U.S.
@@
 ```
-\pagebreak
 
-## Climate
+### Climate
 
 Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
-Spatially explicit projections of changes in annual temperature and rainfall by the year 2100 (Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCPs 4.5 and 8.5.
+Spatially explicit projections of change in annual temperature and change in annual rainfall by the end of century (2100; Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCP 4.5 and RCP 8.5.
 
@@
 ```
-## Change in State Class Area
+
+### Change in State Class Area
 

### 2021-03-30 Manuscript/coverletter/ERL_coverletter.Rmd
updated cover letter files

@@
 mainfont: Times New Roman
-fontsize: 12pt
+fontsize: 11pt
 urlcolor: blue
@@
 Please consider our article, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_.
-Hawaiʻi ellentesque non pretium turpis.
-Etiam nibh nibh, pharetra vel semper quis, rhoncus non sapien.
-Nunc gravida, elit nec malesuada elementum, purus orci consequat lacus, et gravida felis felis eu lacus.
-Aenean ac dapibus purus, eget consequat eros.
-Sed eros odio, rutrum vitae posuere id, facilisis eu elit.
-Duis sed lorem et tortor consectetur mattis.
-In nisi arcu, aliquet a purus in, sodales pulvinar sapien.
-Nulla auctor, orci id sodales consequat, augue odio egestas libero, tincidunt fringilla eros neque at orci.
-Duis accumsan tellus sapien, id cursus odio egestas in.
-Integer in nibh suscipit mauris auctor gravida.
+This article presents results from a stochastic, spatially explicit simulation model used to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands over a 90-year timespan from 2010-2100.
+Our results represent the first set of Hawaiʻi-specific projections to incorporate multiple land use and climate change scenarios, the first to  extend to the end of the 21st century, and the first to be conducted at sub-kilometer spatial resolution.
+This article also represents one of very few regional-scale efforts to examine how ecosystem carbon balance responds to different levels of a CO~2~ fertilization effect on net primary productivity.
 
-Aenean dapibus placerat laoreet.
-Maecenas id arcu semper, feugiat urna eget, placerat ipsum.
-Maecenas porta sit amet ante vel finibus.
-Quisque porttitor urna gravida commodo ullamcorper.
-Nullam dapibus sodales lacus vitae facilisis.
-Praesent nec velit eget quam congue aliquet vitae sed urna.
-Duis venenatis ornare sagittis.
-Nullam dolor odio, porttitor pulvinar aliquet in, rutrum id dolor.
-Nullam vestibulum ligula diam, at fringilla tellus aliquet eu.
-In sit amet magna ac ante efficitur laoreet ut eget lectus.
-Aenean mauris quam, accumsan fermentum massa ac, blandit hendrerit arcu.
-Proin dignissim sapien quis maximus faucibus.
-Aenean lobortis tellus nec nulla fringilla, vitae sagittis felis pretium.
-Pellentesque nec luctus urna.
+Our results are timely given the recent passage of legislation by the State of Hawaiʻi to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems.
+By incorporating the interactive effects of land use and climate change into future projections of ecosystem carbon balance, our model results could serve as baseline projections for the State of Hawaiʻi to evaluate different ecosystem-based climate mitigation strategies.
+Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.
 
-Cras velit neque, pulvinar vitae eros eget, commodo dignissim mauris.
-Mauris ac facilisis tortor.
-Mauris imperdiet diam quis justo posuere convallis non et felis.
-Donec vel justo orci.
-In et volutpat est.
-Integer lacinia faucibus tellus, non vehicula enim faucibus eget.
+We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions.
+We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration.
+We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.
 

### 2021-04-05 Manuscript/HI_Model_ms.Rmd
updated manuscript files, added journal cover letter
updated manuscript files

@@
 In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.341 Tg C y^-1^; @selmants_baseline_2017].
-This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our modelʻs finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018].
+This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our model's finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018].
 Previous versions of a Hawaiʻi LUCAS model were run at 1-km spatial resolution and simulation cells within each unique combination of moisture zone and state type all had the same mean IBIS-derived NPP value applied to them at the beginning of the simulation period.
@@
 The distribution and potential range of this emerging threat to Hawaiʻi's dominant native tree species have only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
-ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, potentially alter the long-term trajectory of forest carbon cycling in Hawaiʻi.
+ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, potentially altering the long-term trajectory of carbon cycling in Hawaiian forest ecosystems.
 New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.
@@
 harvesting in Hawaiʻi.
-Thanks also to Christian Giardina and Zhiliang Zhu for providing the initial impetus for this research.
+Thanks also to Christian Giardina and Zhiliang Zhu for providing the impetus for this research.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
@@
 
+Clay Trauernicht https://orcid.org/0000-0002-1509-8536
+
 Abby G. Frazier https://orcid.org/0000-0003-4076-4577

### 2021-04-26 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 \centering
-Paul C. Selmants^1,7^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,
+Paul C. Selmants^1^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,
 
@@
 ^6^Arizona State University, Tempe, AZ, USA
-^7^Author to whom correspondence should be addressed
 
@@
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
+We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
 Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength.

### 2021-05-04 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019].
-There is increasing recognition among policymakers that natural and agricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017].
+There is increasing recognition among policymakers that natural and agricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural climate solutions” [@cameron_ecosystem_2017].
 Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-efficient and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018].

### 2021-05-12 Manuscript/HI_Model_ms20210405.Rmd
updated manuscript files

--- 2021-04-05 Manuscript/HI_Model_ms.Rmd
@@
 In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.341 Tg C y^-1^; @selmants_baseline_2017].
-This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our model's finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018].
+This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our modelʻs finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018].
 Previous versions of a Hawaiʻi LUCAS model were run at 1-km spatial resolution and simulation cells within each unique combination of moisture zone and state type all had the same mean IBIS-derived NPP value applied to them at the beginning of the simulation period.
@@
 The distribution and potential range of this emerging threat to Hawaiʻi's dominant native tree species have only recently been mapped [@vaughn_approach_2018; @fortini_evolving_2019], but the pulse of newly dead organic matter and reduction in photosynthetic capacity induced by widespread tree mortality could significantly alter ecosystem carbon balance over the long-term [@sleeter_effects_2019].
-ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, potentially altering the long-term trajectory of carbon cycling in Hawaiian forest ecosystems.
+ROD-affected forests could also undergo a replacement of canopy dominant stress-tolerating ʻōhiʻa trees by non-native tree species with faster relative growth rates, lower wood density, and faster tissue turnover, potentially alter the long-term trajectory of forest carbon cycling in Hawaiʻi.
 New model projections for Hawaiʻi that incorporate ROD spread rates and forest restoration scenarios will therefore require differentiation among forest ecosystems dominated by native and non-native tree species.

### 2021-05-14 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
 Incorporating the interactive effects of land use and climate into spatially explicit future projections of ecosystem carbon balance could therefore provide a reference point to evaluate the effectiveness of land-based mitigation.
-Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future land use and climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
+Although a complex challenge, a growing number of sub-national jurisdictions plan to incorporate land-based mitigation strategies into their emissions reduction efforts.
+These jurisdictions would benefit from understanding how future land use and climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].
 
@@
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
-our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.
+Our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.
 
 ## 2. Methods
-We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5).
-We also developed a separate set of scenarios to test model sensitivity to different levels of a CFE on NPP.
-The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016].
+We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands.
+The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined land cover class  assigned to each simulation cell and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016].
+Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016].
 The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019].
@@
 We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
-Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
+Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
 Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
@@
 
- Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
+ Transition targets set the area to be transitioned over time.
+These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
 For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
@@
 All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
+
  Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on mid-century (2049-2069) and end-of-century (2070-2099) statistically downscaled CMIP5 temperature and rainfall projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
 The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
-The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario.
-We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014].
+The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q~10~ temperature coefficients based on climate model temperature projections for each radiative forcing scenario.
+The Q~10~ temperature coefficients represented the proportional change in detrital carbon turnover due to a 10° C change in mean annual temperature.
+We applied a Q~10~ of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q~10~ of 2.17 for litter decay flows [@bothwell_leaf_2014].
 Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.
@@
 For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further stimulation of NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
-This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
+This 600 ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
 
@@
 Interannual climate variability is a primary factor influencing spatial and temporal patterns of global wildland fire activity [@abatzoglou_global_2018], with climate warming expected to increase wildland fire frequency and wildfire season length across a wide range of biomes [@sun_global_2019].
-Although our model projections capture the spatial and temporal variation in ignition probability and fire extent by sampling from previous fire years (1999-2019), we did not incorporate how projected future climate change will affect the frequency and extent of wildland fire.
+Although our model projections capture the spatial and temporal variation in ignition probability and fire extent by sampling from previous fire years (1999-2019), we did not incorporate the effects of projected future climate change on the frequency and extent of wildland fire.
 Recent fire probability modeling for the northwest portion of Hawaiʻi Island indicated that projected drying and warming trends under RCP 8.5 could increase maximum fire probability values more than three-fold and shift areas of peak flammability to higher elevation by mid-century [@trauernicht_vegetationrainfall_2019].

### 2021-08-03 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated journal correspondence directory

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
header-includes:
   - \linespread{1.05}
   - \newcommand{\blankline}{\quad\pagebreak[2]}
   - \usepackage{eso-pic}
   - \usepackage{graphicx}
   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_response_letterhead.pdf} \vfill }}}
geometry: top=1in, left=1in, bottom=1in, right=1in
mainfont: Times New Roman
fontsize: 12pt
urlcolor: blue
---

\AddToShipoutPicture*{\BackgroundPic}
\pagenumbering{gobble}

`r format(Sys.time(), '%B %d, %Y')`

\blankline

Dear Editors:

\blankline

Please consider our article, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. This article presents results from a stochastic, spatially explicit simulation model used to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands over a 90-year timespan from 2010-2100. Our results represent the first set of Hawaiʻi-specific projections to incorporate multiple land use and climate change scenarios, the first to  extend to the end of the 21st century, and the first to be conducted at sub-kilometer spatial resolution. This article also represents one of very few regional-scale efforts to examine how ecosystem carbon balance responds to different levels of a CO~2~ fertilization effect on net primary productivity.

Our results are timely given the recent passage of legislation by the State of Hawaiʻi to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. By incorporating the interactive effects of land use and climate change into future projections of ecosystem carbon balance, our model results could serve as baseline projections for the State of Hawaiʻi to evaluate different ecosystem-based climate mitigation strategies. Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.

We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions. We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration. We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.

\blankline

Best regards,

\raggedright
Paul C. Selmants 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

Lots and lots more text, repeating last paragraph of letter here to see what happens on page two.

We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions. We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration. We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.

### 2021-08-03 Manuscript/HI_Model_ms.Rmd
updated journal correspondence directory

@@
 We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-We explored four unique scenarios that represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
+We developed four unique scenarios that explored different pathways, or future trajectories, of land use change and climate change.
+These four scenarios represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
@@
 We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands.
-The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined land cover class  assigned to each simulation cell and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016].
+The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined information specific to each simulation cell (e.g., land cover class, climate zone) and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016].
 Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016].
@@
 We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
+Transition pathways define which state types can be converted to any other state type.
 Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization).
+Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, i.e., they could each transition to multiple other state types.
+The barren state type could only transition to developed, so its sole transition pathway was urbanization.
 Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state.
@@
 In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
-The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
+The remaining 40% were rotation harvests replanted to *Eucalyptus* spp., which represents the vast majority of plantation forestry in the State of Hawaiʻi.
 

### 2021-08-03 Manuscript/journal_correspondence/ERL_responseletter.Rmd
moved NPP spatial multiplier and updated response letter

@@
    - \usepackage{graphicx}
-   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_response_letterhead.pdf} \vfill }}}
+   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_letterhead.pdf} \vfill }}}
 geometry: top=1in, left=1in, bottom=1in, right=1in
@@
 \AddToShipoutPicture*{\BackgroundPic}
-\pagenumbering{gobble}
+\pagenumbering{arabic}
 
@@
 
-Please consider our article, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_.
-This article presents results from a stochastic, spatially explicit simulation model used to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands over a 90-year timespan from 2010-2100.
-Our results represent the first set of Hawaiʻi-specific projections to incorporate multiple land use and climate change scenarios, the first to  extend to the end of the 21st century, and the first to be conducted at sub-kilometer spatial resolution.
-This article also represents one of very few regional-scale efforts to examine how ecosystem carbon balance responds to different levels of a CO~2~ fertilization effect on net primary productivity.
+Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_.
+We have fully addressed each of the referees’ comments in detail below.
+Our point-by-point replies immediately follow each referee comment, and we include line numbers of the revised manuscript at the end of each response to identify where we made changes to the text.
 
-Our results are timely given the recent passage of legislation by the State of Hawaiʻi to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems.
-By incorporating the interactive effects of land use and climate change into future projections of ecosystem carbon balance, our model results could serve as baseline projections for the State of Hawaiʻi to evaluate different ecosystem-based climate mitigation strategies.
-Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.
-
-We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions.
-We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration.
-We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.
+Long-term model projections such as ours will be critical to understanding how future land use and climate change may interact to influence the achievement of climate mitigation goals, and so we expect our paper to be of immediate interest to a wide audience of carbon cycle scientists, ecosystem modelers, land managers, and policy makers.
+We hope you find our revisions and responses to referees’ comments sufficient to merit publication of our manuscript.
 
@@
 \raggedright
-Paul C. Selmants 
+Paul C. Selmants, Ph.D. 
 \linebreak
@@
 
-Lots and lots more text, repeating last paragraph of letter here to see what happens on page two.
+## Referee 1, comments to the author:
 
-We found that local reductions in the intensity of land use change consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global climate mitigation through greenhouse gas emissions reductions.
-We also found that CO~2~ fertilization of NPP was the largest source of uncertainty in long-term projections of ecosystem carbon balance in the Hawaiian Islands, highlighting the need for greater mechanistic understanding of the cascading effects of rising atmospheric CO~2~ on ecosystem carbon sequestration.
-We believe long-term model projections such as ours will be critical to understanding how future land use and climate change could interact to influence the achievement of climate mitigation goals.
+**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios.
+This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration.
+The manuscript is well written and organized.
+However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled.
+And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed.
+I have provided detailed comments below.**
+
+**L109, why need to assign each cell a state type? What is consideration for it?**
+
+**L114, transitions between ‘state type’?
+Is it ‘land cover classes’?
+According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**
+
+**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median?
+What is the justification that low land use scenario will have rate lower than historical records?**
+
+**L126, missing description for urbanization rate for high land use scenario.**
+
+**L133, what transitions can increase forest area such as low land use scenario shown in figure S5?
+Does tree plantation refer to reforestation or afforestation?
+If afforestation, what is the source land cover class for this plantation?**
+
+**L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**
+
+**L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**
+
+**L140, is state type land cover classes?
+Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**
+
+**L158, moisture zone is included in state type as defined at line L110, isn’t it?**
+
+**L162, how litter and dead wood pools were initialized?**
+
+ **L161, what is 30-year climate normal come from?
+Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020?
+Or is there any bias between 30-year climate normal and RCP projections?**
+
+**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here.
+Where is the additional 32 Tg C from?
+What is ratio of aboveground biomass to total live biomass in IBIS?**
+
+**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
+
+**L166-168, is similar multiplier also applied to living biomass?
+If not, will this single NPP multiplier break the equilibrium between NPP and biomass?
+For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees.
+External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**
+
+**Line 173, why assume Rh as ratio of NPP?
+Is the ratio constant over time and across climate change scenarios?
+Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks?
+Because of lack of soil temperature in RCP 4.5 and 8.5?
+Need more justification for this flux ratio method.**
+
+**L177-179, are the growth and decay multipliers varied at annual basis?
+How are multipliers derived for time periods CMIP5 projections do not cover?
+Such as period between 2020-2049?
+It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?**
+
+**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance.
+It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall.
+For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**
+
+**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**
+
+**L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**
+
+**L242.
+It is misleading to state ‘climate change led to increased Rh over time’ as figure 3 shows decreasing trend of Rh under high land use scenario and two RCP climate change scenarios.
+It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
+
+**L244, why the increase in total soil carbon stock (figure 2) will not enhance Rh?
+Is it due to the assumption made by this work that Rh is only ratio of NPP and latter decrease over time?
+What are drivers of Rh decrease?**
+
+**L244, why Rh shows very sharp decrease before 2020 under all land use scenarios with all RCPs, such decrease rate seems to be higher than the rest period.
+Are these decreases supported by observations?**
+
+**Line 341, what is the source of ‘Hawai’i-specific climate data’?**
+
+## Referee 2, comments to the author:
+
+**The manuscript entitled “Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change” is interesting and suitable for publication but several mistakes in the manuscript should be corrected before acceptance.**
+
+**Specific comments**
+
+**Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**
+
+**All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**
+
+**I suggest that authors check the bibliographic references in the main text and revise the format of documents to meet the requirement of the Journal.**
+
+**In the abstract, please add more information about the methods of the experiment.**
+
+**In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**
+
+**Hypothesis should be given. How this work is different from the available literature?**
+
+**Objectives should be rewritten in detail and comprehensively.**
+
+**How did the authors perform the calibration and validation process?**
+
+**How did they select the sensitive parameters of the model that they used?**
+
+**Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?**

### 2021-08-04 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated response letter

@@
 \AddToShipoutPicture*{\BackgroundPic}
-\pagenumbering{arabic}
 
@@
 Long-term model projections such as ours will be critical to understanding how future land use and climate change may interact to influence the achievement of climate mitigation goals, and so we expect our paper to be of immediate interest to a wide audience of carbon cycle scientists, ecosystem modelers, land managers, and policy makers.
-We hope you find our revisions and responses to referees’ comments sufficient to merit publication of our manuscript.
+We hope you find the revisions and responses to referees’ comments sufficient to merit publication of our manuscript.
 
@@
 **L109, why need to assign each cell a state type? What is consideration for it?**  
+The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition.
+This set of information defines the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
+We have added text and references to this section and the next that better describes the state of each cell (lines ).
 
@@
 According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**
+State type in this model is the unique combination of moisture zone, island, and land cover class, as stated on line 110 of the original manuscript.
+This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic grassland on Oʻahu.
+We have added text that better defines how state type influences transition probabilities ().
 
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
+To quantify the impact of a low land use scenario, rates of land use change were sampled from the lower range of historical rates (median and below, bounded by zero).
+We have added text that justifies land use change rates sampled from a range lower than historical rates in the low land use scenario (lines xxx - xxx).
 
 **L126, missing description for urbanization rate for high land use scenario.**  
+As the referee themselves point this out in their previous comment, transition rates in the high land use scenario were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural expansion, agricultural contraction, and urbanization for each island.
+This description is on lines 121-123 of the original manuscript and lines xxx - xxxx of the revised manuscript.
+No changes were made in response to this comment.
 
@@
 **L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
+We have added text defining the historical harvest rate as (lines xxx - xxxx).
 
 **L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**  
+Wildland fire in Hawaiʻi takes place primarily in non-native grasslands, but can also occur in shrublands and forest.
+We have added text that describes where wildland fire can occur in the Hawaiian Islands (lines xxx - xxx).
 
@@
 Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**
+We agree that this was a confusing use of terminology. We have changed "state type" to "land cover class" (line xxx).
 
 **L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
+We agree this was, again, a confusing use of terminology. We have omitted "moisture zone" and now only refer to "state type" in this sentence (line xxx).
 
@@
 Or is there any bias between 30-year climate normal and RCP projections?**
+ The 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al. 2013, 2014).
 
@@
 What is ratio of aboveground biomass to total live biomass in IBIS?**
+Asner et al.
+(2016) only quantifies aboveground forest biomass.
+The additional 32 Tg C comes from belowground forest biomass, as well as total live biomass in grasslands, shrublands, and agriculture.
+We have added text describing how we estimated belowground forest biomass.
 
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
+We have added text listing the range and median rate of GPP from the calibration dataset (Kimball et al.
+2017) and added a table of mean NPP rates from IBIS for each moisture zone and land cover class (xxx).
 
@@
 External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**
+Biomass already accumulated will not be reduced if NPP declines.
+Rather, the rate of biomass increase will decline as NPP declines.
+This is true in our model, as it is in the real world.
+We have added text that better describes how and why we apply the NPP spatial multiplier (xxx).
 
@@
 For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**
+Projections of NPP change by scenario are presented in Figure 3.
 
 **L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
+Yes, climate projections only cover the period of mid-century (2049) and end-of-century (2070-2099) as described on line 179 of the original manuscript. We have added text that described how we
 

### 2021-08-06 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript file, supplementary file, and response letter file.

--- 2021-08-03 Manuscript/journal_correspondence/ERL_responseletter.Rmd
@@
 Long-term model projections such as ours will be critical to understanding how future land use and climate change may interact to influence the achievement of climate mitigation goals, and so we expect our paper to be of immediate interest to a wide audience of carbon cycle scientists, ecosystem modelers, land managers, and policy makers.
-We hope you find our revisions and responses to referees’ comments sufficient to merit publication of our manuscript.
+We hope you find our revisions and responses to referees’ comments sufficient to merit publication of this manuscript in _Environmental Research Letters_.
 
@@
 
-**L109, why need to assign each cell a state type? What is consideration for it?**
+**L109, why need to assign each cell a state type?
+What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition.
+This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
+We have added text and references to this section and the next that better describes the state of each cell (lines ).
 
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
+To quantify the impact of a low land use scenario, rates of agricultural contraction and agricultural expansion were sampled from the lower range of historical rates (median and below, bounded by zero).
+Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time, because no change in agricultural land area will occur during one or more timesteps within a given Monte Carlo iteration.
+We consider this to be a reasonable possibility under a low land use scenario, but the vast majority of timesteps and Monte Carlo iterations will experience change in agricultural land area within the historical range.
+We have added text that justifies bounding the lower end of agricultural land area change rates with zero rather than the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
 
 **L126, missing description for urbanization rate for high land use scenario.**  
+As the referee themselves point this out in their previous comment, transition rates in the high land use scenario, including urbanization, were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural expansion, agricultural contraction, and urbanization for each island.
+This description is on lines 121-123 of the original manuscript and lines xxx - xxxx of the revised manuscript.
+No changes were made in response to this comment.
 
@@
 If afforestation, what is the source land cover class for this plantation?**
+Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land).
+We have added text to the supplementary that explains the increase in both forest area and grassland area (supplementary lines xxx - xxx).
+Tree plantation does not refer to reforestation or afforestation.
+We did not not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
+Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (Fig.
+S5).
+We have added text to the Methods section that describes tree plantation harvest rates and transitions (lines xxx-xxx).
+The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al.
+(2017).
 
 **L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
+We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines xxx - xxx).
 
 **L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**  
+Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests.
+We have added text that describes where wildland fire occurs in the Hawaiian Islands (lines xxx - xxx).
 
@@
 Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**
+We agree that this was a confusing use of terminology. We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript.
 
 **L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
+Again, we agree this was a confusing use of terminology.
+We have omitted "moisture zone" and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class.
 
 **L162, how litter and dead wood pools were initialized?**  
+We have added text stating that litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).
 
@@
 Or is there any bias between 30-year climate normal and RCP projections?**
+The contemporary 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al.
+2013, 2014), as depicted in Supplementary figure 2.
+We have added text that better describes this (lines xxx-xxx).
 
@@
 What is ratio of aboveground biomass to total live biomass in IBIS?**
+Asner et al.
+(2016) only quantifies aboveground forest biomass.
+The additional 32 Tg C comes from the sum of belowground forest biomass and total live biomass in grasslands, shrublands, and agriculture.
+We have added text describing the equation we used to estimate belowground forest biomass, which was then used to calibrate the IBIS DGVM (lines xxx-xxx).
 
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
+We have added text (lines xxx-xxx) listing the range and median rate of GPP from the calibration dataset (Kimball et al.
+2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (Supplementary Table 1).
 
@@
 External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**
+We have added maps to the supplementary depicting the spatial distribution of stationary NPP multipliers as well as the spatial distribution of NPP that results from the product of IBIS mean NPP values and the NPP spatial multipliers based on the spatial variability of contemporary climate.
+In our model, as in the real world, a deline in NPP will result in a decline in biomass accumulation, but the biomass that has already accumulated persists and will not somehow be reduced just because growth rates have slowed.
+We have added text that better describes how and why we apply the NPP spatial multiplier (xxx).
 
@@
 Need more justification for this flux ratio method.**
+We have added text and references to this section justifying the fact that plant productivity and ecosystem respiration are highly correlated on annual timescales, and ecosystem respiration tends to scale with carbon inputs to the rhizosphere from photosynthesis (lines xxx).
+We agree it would be nice to also incorporate gridded annual maps of projected soil moisture and soil temperature, but these data products do not exist for the Hawaiian Islands.
 
@@
 It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?**
+Yes, the growth and decay multipliers vary annually.
+We have added text that describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx).
+The NPP trend over time under both RCPs is depicted in Figure 3.
+We believe this is a much more clear and succinct way to convey changes in NPP than in a series of maps of future projected NPP under each of the four scenarios.
+Presenting projected future NPP as we do in Figure 3 is consistent with the focus of this paper, which is on statewide trends of ecosystem carbon balance under different climate and land use scenarios.
 
@@
 For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**
+Again, projections of NPP change by land use and climate change scenario are presented in Figure 3, which we believe best conveys statewide trends of ecosystem carbon fluxes under different climate and land use scenarios.
+We have added text to the supplementary that better describes the empirical NPP model (lines xxx).
 
 **L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
+Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript.
+We have added text that describes how to the supplementary climate section that provides more detail on climate projections (lines xxx) No changes were made in response to this comment.
 
 **L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
+Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ to 7.95 Tg y^-1^), and the ranges of statewide NPP estimates for each of these years broadly overlapped (2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
+This modeled slight decline in NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
+We have added text that describes this change in NPP over the contemporary (2010-2020) time period (lines xxx-xxx).
 
@@
 It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
+We agree, this phrasing was misleading. We have altered the text
 
@@
 Are these decreases supported by observations?**
+The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of agricultural land (supplementary Figure 5).
+Agriculture has a higher rate of Rh per unit of NPP than other land cover classes, and so Rh declined more than NPP as agricultural land was converted to urban area.
+We have added text that describes these trends (lines xxx).
 
 **Line 341, what is the source of ‘Hawai’i-specific climate data’?**
+The Hawaiʻi specific climate data is from Giambelluca et al.
+(2013), Giambelluca (2014), Ellison Timm et al.
+(2015), and Ellison Timm et al.
+(2017) - all of which are cited several times throughout the manuscript and are depicted in supplementary figures 3 and 4.
+We have added citations to these data again here in response to this comment (lines xxx).
 
@@
 **Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
+We have carefully checked the manuscript for typographical errors and corrected them on lines x, y, and z of the revised manuscript.
+We have also confirmed that all acronyms are spelled out upon initial use.
 
 **All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
+We have checked the clarity of all figures by requesting multiple colleagues open the manuscript PDF file on their respective computers.
+None of these colleagues noted a lack of clarity in the figures.
+We would be happy to alter these figures if the ERL production team indicates they do not meet journal requirements.
 
 **I suggest that authors check the bibliographic references in the main text and revise the format of documents to meet the requirement of the Journal.**
+We are using the Institute of Physics (IOP) citation style language (csl), which meets the guidelines for this Journal. No changes were made in response to this comment.
 
-**In the abstract, please add more information about the methods of the experiment.**
+**In the abstract, please add more information about the methods of the experiment.**  We have added text to the abstract that provides more information about the methods of this project (lines xxx).
 
 **In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**  
+We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines xxx).
 
 **Hypothesis should be given. How this work is different from the available literature?**
+Since this study does not involve a controlled experiment, we believe a hypothesis statement is unwarranted and that a statement of objectives is sufficient.
+We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx), as stated in the Author guidelines for this Journal.
 
 **Objectives should be rewritten in detail and comprehensively.** 
+We believe we have stated the objectives of this study clearly and succinctly, both in the abstract (lines xxx) and at the end of the Introduction section (lines xxx).
+That said, we have altered the text to add more detail to our stated objectives (lines xxx-xxx).
 
 **How did the authors perform the calibration and validation process?**
+We describe calibration in the Methods section on lines xxx and lines xxx. Validation of ecosystem carbon balance estimates in the Hawaiian Islands is
 
 **How did they select the sensitive parameters of the model that they used?**
+We are unsure exactly what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
+There are a number of parameters in this model, and we describe in detail how they were selected in the Methods section and the Supplementary section.
+Without further clarification, we do not know how to better respond to this question.
+No changes were made in response to this comment.
 
 **Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?** 
+We disagree with this referee that our discussion section is unrelated to the results section.
+We followed the author guidelines for this journal by discussing the significance of the results and comparing them with previous work using relevant references.
+For example,

### 2021-08-06 Manuscript/HI_Model_ms.Rmd
updated manuscript file, supplementary file, and response letter file.

@@
 However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands.
+We used the Land Use and Carbon Scenario Simulator (LUCAS), a spatially explicit stochastic  simulation model that integrates landscape change and carbon gain-loss, to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
 We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
@@
 Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016].
+Transition pathways define which state types can be converted to any other state type.
 The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019].
@@
 We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016.
-Transition pathways define which state types can be converted to any other state type.
-Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, i.e., they could each transition to multiple other state types.
-The barren state type could only transition to developed, so its sole transition pathway was urbanization.
-Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state.
-Water and wetland state types remained static throughout the simulation period.
+Transition pathways were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
+Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into multiple other land cover classes.
+The barren land cover class could only transition to developed, so its sole transition pathway was urbanization.
+Although most land cover classes had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state.
+Water and wetland land cover classes remained static throughout the simulation period.
 
@@
 The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
+Tree plantation forestry in the State of Hawaiʻi is primarily short-rotation (5-7 years) *Eucalyptus* spp., harvested at a rate of approximately 2 km^2^ y^-1^ over the past decade.
 Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016].
 In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
-The remaining 40% were rotation harvests replanted to *Eucalyptus* spp., which represents the vast majority of plantation forestry in the State of Hawaiʻi.
+The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
+No land cover class could be converted to tree plantation in either land use scenario, consistent with recent historical trends of stable or declining area of tree plantations.
 
 The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (supplemental figure 2).
-We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (supplemental figure 1), island, and state type (figure 1) for the years 1999-2019.
+We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (supplemental figure 1), island, and land cover class (figure 1) for the years 1999-2019.
 Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017.
@@
 The high land use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a).
+The vast majority of wildland fire in Hawaiʻi is human-caused and initiates in non-native grasslands and shrublands [@trauernicht_contemporary_2015], but can spread into nearby forest areas.
+Moderate to high severity fires that spread into mesic and wet forests can convert these areas into grasslands about half the time.
+However, moderate to high severity fires in mesic and wet forests account for <5% of the total annual area burned, on average [@selmants_baseline_2017].
 
@@
 
- Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (supplemental figure 1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+ Initial carbon stocks and baseline carbon flows were estimated based on the state type and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
@@
 
- Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and state type adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [@sleeter_effects_2019].
-We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and state type based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
+ Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2019].
+We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and land cover class based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
 Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell.

### 2021-08-06 Manuscript/HI_Model_ms20210405.Rmd
updated manuscript file, supplementary file, and response letter file.

@@
 For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
+Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time, because one or more timesteps within a given Monte Carlo iteration may experience no change in agricultural land area.
+We consider this to be a reasonable possibility under a low land use change scenario, but the vast majority of timesteps and Monte Carlo iterations will experience changes in agricultural land area within the historical range.
 Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].

### 2021-08-06 Manuscript/HI_Model_supp.Rmd
updated manuscript file, supplementary file, and response letter file.

@@
 
-Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
-Spatially explicit projections of change in annual temperature and change in annual rainfall by the end of century (2100; Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCP 4.5 and RCP 8.5.
+Spatially explicit contemporary mean annual temperature and rainfall data (30-year means) for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
+Spatially explicit projections of change in annual temperature and change in annual rainfall for mid-century (2049-2069) and end of century (2070-2099; Figure S4) were developed using statistically downscaled CMIP5 climate projections under RCP 4.5 and RCP 8.5 for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017.
+To avoid biases, the modeled present-day (1975–2005) CMIP5 climatology was used to standardize the resulting predictor time series, and the simulated future changes in the predictors were measured relative to their present-day mean states.
+See @timm_statistical_2015 and @timm_future_2017 for a more detailed description of potential statistical downscaling biases.
 
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Contemporary 30-year climate normals for mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
 Data from Giambelluca \\textit{et al.} (2013) and Giambelluca \\textit{et al.} (2014)."}
@@
 
+### Initializing NPP
+
+```{r}
+library(kableExtra)
+
+ibisnpp <- read.csv("IBISnpp_wide.csv", stringsAsFactors = FALSE)
+kbl(ibisnpp, booktabs = T, caption = "Supplemental - Mean IBIS-derived NPP values by Land Cover and Moisture Zone, in kg of C per square meter per year.") %>%
+   add_header_above(c(" ", "NPP by Moisture Zone" = 3))
+```
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Stationary spatial NPP multipliers (top panel) and initial NPP based on the product of mean IBIS-derived NPP values (Table 1) and stationary spatial NPP multipliers (above).
+Developed and Barren land cover types are excluded from both maps."}
+knitr::include_graphics("./fig_images/figS4_initialNPP.png")
+```
 ### Change in State Class Area

### 2021-08-06 Transitions/code/HI_Transition_Targets.R

@@ -49,7 +49,7 @@ UrbanLow_early <- ddev %>%
 		SecondaryStratumID = island, TransitionGroupID, 
 		Amount, DistributionType, DistributionFrequencyID, 
 		DistributionSD, DistributionMin = min, DistributionMax = ddev_km2y) 
-# Low urbanization transition targes by island for 2045-2070
+# Low urbanization transition targets by island for 2045-2070
 # using zero to 25% of minimum annual population projection per island
 UrbanLow_mid <- ddev %>%
 	group_by(island) %>%

### 2021-08-09 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated references, manuscript file, and figure files

@@
 This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
-We have added text and references to this section and the next that better describes the state of each cell (lines ).
+We have added text and references to this section and the next that better describes the state of each cell (lines xxx).
 
@@
 State type in this model is the unique combination of moisture zone, island, and land cover class, as stated on line 110 of the original manuscript.
-This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic grassland on Oʻahu.
+This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic shrubland on Oʻahu.
 We have added text that better defines how state type influences transition probabilities ().
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
-To quantify the impact of a low land use scenario, rates of agricultural contraction and agricultural expansion were sampled from the lower range of historical rates (median and below, bounded by zero).
-Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time, because no change in agricultural land area will occur during one or more timesteps within a given Monte Carlo iteration.
-We consider this to be a reasonable possibility under a low land use scenario, but the vast majority of timesteps and Monte Carlo iterations will experience change in agricultural land area within the historical range.
-We have added text that justifies bounding the lower end of agricultural land area change rates with zero rather than the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
+To distinguish the impact of a low land use scenario from a high land use scenario on changes in agricultural land area, we sampled rates of agricultural contraction and agricultural expansion from a uniform distribution bounded by zero and the minimum historical rate.
+Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time.
+We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010).
+We have added text that justifies bounding  agricultural land area change rates between zero rather and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
 
@@
 What are drivers of Rh decrease?**
+Soil carbon stocks increase consistently only under the RCP 4.5 radiative forcing scenario.
+Under RCP 8.5, soil carbon stocks reach an asymptote by 2080 in the low land use scenario and begin to decline slightly in the high land use scenario.
+Also, Rh only declines consistently under the high land use scenario (Figure 3).
+In the low land use scenario Rh remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5.
+As stated previously, NPP is a direct driver of Rh because the vast majority of carbon respired heterotrphically from terrestrial ecosystems was fixed by photosynthesis that same year (Kuzyakov et al.
+2010; see lines xxx of revised manuscript).
+The drivers of Rh are a combination of climate change and land use change.
+High rates of urbanization remove plants and compress and pave over soils, reducing carbon input and lowering microbial activity.
+Climate change on the the other hand, has a ʻpush-pullʻ impact on Rh.
+On the one hand
 
@@
 Are these decreases supported by observations?**
+Land use change rates are the same in all scenarios from 2010-2020, which explains why Rh behaves consistently across these scenarios.
+We have added text to make that clear in the Methods section (lines xxx).
 The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of agricultural land (supplementary Figure 5).
-Agriculture has a higher rate of Rh per unit of NPP than other land cover classes, and so Rh declined more than NPP as agricultural land was converted to urban area.
+Agriculture has a higher rate of Rh per unit of NPP than other land cover classes, and so Rh declined more than overall NPP as agricultural land was converted to urban area.
 We have added text that describes these trends (lines xxx).
@@
 **Hypothesis should be given. How this work is different from the available literature?**
-Since this study does not involve a controlled experiment, we believe a hypothesis statement is unwarranted and that a statement of objectives is sufficient.
-We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx), as stated in the Author guidelines for this Journal.
+A hypothesis would be warranted if the purpose of this study was to test existing theory using a controlled experiment, which was not the case.
+The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient.
+We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx), as recommended by the author guidelines for this Journal.
+No changes were made in response to this comment.
 
@@
 **How did the authors perform the calibration and validation process?**
-We describe calibration in the Methods section on lines xxx and lines xxx. Validation of ecosystem carbon balance estimates in the Hawaiian Islands is
+We describe calibration of the model in the Methods section on lines xxx and lines xxx. Validation of ecosystem carbon balance estimates in the Hawaiian Islands is
 
@@
 We are unsure exactly what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
-There are a number of parameters in this model, and we describe in detail how they were selected in the Methods section and the Supplementary section.
+There are a number of parameters in this model, and we describe in detail how they were selected in both the Methods section and the Supplementary section.
 Without further clarification, we do not know how to better respond to this question.

### 2021-08-09 Manuscript/HI_Model_ms.Rmd
updated references, manuscript file, and figure files

@@
 The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined information specific to each simulation cell (e.g., land cover class, climate zone) and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016].
-Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016].
+Landscape change occurs by applying a Monte Carlo approach to track the state and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016].
 Transition pathways define which state types can be converted to any other state type.
@@
 Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; supplemental figure 1), seven islands, and ten discrete land cover classes (figure 1).
+We also tracked the age and time since transition for each simulation cell as continuous state variables [@daniel_state-and-transition_2016].
 
@@
 Transition pathways were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire.
-Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into multiple other land cover classes.
+Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into a variety of other land cover classes.
 The barren land cover class could only transition to developed, so its sole transition pathway was urbanization.
@@
 Transition targets set the area to be transitioned over time.
-These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
-For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
-For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island.
+These targets were based on recent historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
+For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum recent historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
+For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum recent historical rates for each island.
+Bounding agricultural change rates between zero and the recent historical minimum shifts the balance toward agricultural contraction over time in the low land use scenario, which reduces the land area under active cultivation and contrasts with the high variability in agricultural land change rates during the recent historical past [@suryanata_diversified_2009].
 Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
@@
 The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].
+Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016].
 Tree plantation forestry in the State of Hawaiʻi is primarily short-rotation (5-7 years) *Eucalyptus* spp., harvested at a rate of approximately 2 km^2^ y^-1^ over the past decade.
-Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016].
-In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture.
+In both the high and low land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to grassland or agriculture.
 The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
-No land cover class could be converted to tree plantation in either land use scenario, consistent with recent historical trends of stable or declining area of tree plantations.
+There was no transition pathway from any land cover class into tree plantation in either land use scenario, which is consistent with recent historical trends of stable or declining tree plantation land area.
 
@@
 Moderate to high severity fires that spread into mesic and wet forests can convert these areas into grasslands about half the time.
-However, moderate to high severity fires in mesic and wet forests account for <5% of the total annual area burned, on average [@selmants_baseline_2017].
+However, moderate to high severity fires in mesic and wet forests account for < 5% of the total annual area burned, on average [@selmants_baseline_2017; @daniel_integrating_2018].
 
@@
 
- Initial carbon stocks and baseline carbon flows were estimated based on the state type and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
-We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
-We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016].
-We also calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].
+Initial carbon stocks and baseline carbon flows were derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model.
+We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
+We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017.
+We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017].
+Mean annual GPP from @kimball_evaluating_2017 ranged from 0.123 - 3.88 kg C m^-2^ y^-1^ across the Hawaiian Islands, resulting in a NPP-to-GPP ratio ranging from 0.37 - 0.56.
 
- Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2019].
-We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and land cover class based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
-Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell.
-All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
+We initiated net primary production for each simulation cell in LUCAS as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019].
+The spatial NPP multipliers were the NPP anomaly for each simulation cell relative to mean empirically-derived NPP values for each state type (supplemental Figure 4).
+Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each state type to the microclimate-adjusted NPP value for each simulation cell.
+All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
 

### 2021-08-09 Manuscript/HI_Model_supp.Rmd
updated references, manuscript file, and figure files

--- 2021-03-26 Manuscript/HI_Model_supp.Rmd
@@
 ### Moisture Zones  
-
-State Classes for Agriculture, Forest, Grassland, Shrubland, and Tree Plantations were stratified into three Moisture Zones - Dry, Mesic, and Wet (Figure S1).
+Agriculture, Forest, Grassland, Shrubland, and Tree Plantation land cover classes were stratified into three Moisture Zones - Dry, Mesic, and Wet (supplemental Figure 1).
 These three zones were based on a moisture availability index (MAI), calculated as mean annual precipitation (MAP) minus potential evapotranspiration [PET; @price_mz_2012].
@@
 Areas with MAI values greater than 1,661 were classified in the Wet Moisture Zone.
-
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
 knitr::include_graphics("./fig_images/figS1_MZ.png")
 ```
-
 ### Wildland Fire  
-
-This study used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.
-This new database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr.
+We used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (supplemental Figure 2), wildland fire probabilities by state type, and wildland fire size distributions.
+This new database compiles prior mapping efforts and data collections with newly mapped fire perimeter data from Dr.
 Clay Trauernicht (Department of Natural Resources and Environmental Management, University of Hawaiʻi at Mānoa).
-The goal of this effort was to locate and map all fires greater than or equal to 20 hectares, but some smaller fires were included as detected in imagery.
+The goal was to locate and map all fires greater than or equal to 20 hectares, but some smaller fires were included as detected in imagery.
 Fire perimeters for the years 2002-2011 were primarily from the U.S.
 Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
-The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, primarily in the Kona and Kohala regions.
+The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, mostly from the Kona and Kohala regions.
 The U.S.
@@
 Data from USGS MTBS were prioritized for the years 2002-2011 in the case of duplicate records.
-
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
 The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
@@
 ```
-
 ### Climate  
-
-Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (Figure S3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
-Spatially explicit projections of change in annual temperature and change in annual rainfall by the end of century (2100; Figure S4) were developed using statistically downscaled CMIP5 climate projections for the Hawaiian Islands from @timm_statistical_2015 and @timm_future_2017 under RCP 4.5 and RCP 8.5.
-
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
+Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (supplemental Figure 3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
+Spatially explicit projections of change in annual temperature and annual rainfall for mid-century (2049-2069) and end of century (2070-2099; supplemental Figure 4) are  from statistically downscaled CMIP5 climate projections under RCP 4.5 and RCP 8.5 from @timm_statistical_2015 and @timm_future_2017.
+To avoid biases, the modeled present-day (1975–2005) CMIP5 climatology was used to standardize the resulting predictor time series, and the simulated future changes were measured relative to their present-day mean states.
+See @timm_statistical_2015 and @timm_future_2017 for a more detailed description of potential statistical downscaling biases.
+```{r fig.align = 'center', out.width = "80%", fig.cap = "Supplemental - Contemporary 30-year climate normals for mean annual temperature (top panel) and mean annual rainfall (bottom panel) for the seven main Hawaiian Islands.
 Data from Giambelluca \\textit{et al.} (2013) and Giambelluca \\textit{et al.} (2014)."}
@@
 ```
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by 2100 under RCP 4.5 and RCP 8.5 based on statistical downscaling of an ensemble of climate models (CMIP5)."}
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by 2100 under RCP 4.5 and RCP 8.5 based on statistical downscaling of an ensemble of climate models (CMIP5)."}
 knitr::include_graphics("./fig_images/figS4_FutureClimate.png")
 ```
-
+### Initializing NPP 
+Initial values for net primary production (NPP) in LUCAS were calculated using a lookup table derived from the Integrated Biosphere Simulator (IBIS; supplemental Table 1).
+These values were adjusted with spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate (supplemental Figure 4) as described in detail in @sleeter_effects_2018 and @sleeter_effects_2019.
+Briefly, we derived the set of NPP spatial multipliers by first calculating NPP independently of the IBIS values for each simulation cell using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] based on Hawaiʻi-specific climate data (supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014).
+We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each state type (i.e., the unique combination of moisture zone and land cover class; supplemental Figure 4).
+These NPP multipliers were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4).
+```{r}
+library(kableExtra)
+ibisnpp <- read.csv("IBISnpp_wide.csv", stringsAsFactors = FALSE)
+kbl(ibisnpp, booktabs = T, caption = "Supplemental - Mean IBIS-derived NPP values by Land Cover and Moisture Zone, in kg of C per square meter per year.") %>%
+   add_header_above(c(" ", "NPP by Moisture Zone" = 3))
+```
+```{r fig.align = 'center', out.width = "80%", fig.cap = "Supplemental - Stationary spatial NPP multipliers (top panel) and initial NPP based on the product of mean IBIS-derived NPP values (supplemental Table 1) and stationary spatial NPP multipliers (above).
+Developed and Barren land cover types are excluded from both maps."}
+knitr::include_graphics("./fig_images/figS5_initialNPP.png")
+```
 ### Change in State Class Area
-
-Projections of total land area covered by each of five State Classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high; Figure S5).
-
+Projections of total land area covered by each of five State Classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high; supplemental Figure 5).
+The overall decline in Agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland area over the same time period.
+Although Agricultural land area in the high land use scenario appears more stable over time, the trend actually reflects a balance between high rates of agricultural contraction in some areas and high rates of agricultural expansion in others.
 ```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100.
 Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
-knitr::include_graphics("./fig_images/figS5_StateClassArea.png")
+knitr::include_graphics("./fig_images/figS6_StateClassArea.png")
 ```
-
 \pagebreak

### 2021-08-09 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript files, increased resolution to 400 dpi for all figure files

@@
 What is the justification that low land use scenario will have rate lower than historical records?**
-To distinguish the impact of a low land use scenario from a high land use scenario on changes in agricultural land area, we sampled rates of agricultural contraction and agricultural expansion from a uniform distribution bounded by zero and the minimum historical rate.
-Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time.
+To distinguish the impact of a low land use scenario from a high land use scenario on changes in agricultural land area, we sampled rates of agricultural contraction and agricultural expansion in the low land use scenario from a uniform distribution bounded by zero and the minimum historical rate.
+Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus lowering the cropping intensity of land area over time.
 We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010).
-We have added text that justifies bounding  agricultural land area change rates between zero rather and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
+We have added text that justifies bounding agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
 
@@
 The contemporary 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al.
-2013, 2014), as depicted in Supplementary figure 2.
-We have added text that better describes this (lines xxx-xxx).
+2013, 2014), as depicted in Supplementary figure 3.
+We have added text that better describes this (supplementary lines xxx-xxx).
 
@@
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
-We have added text (lines xxx-xxx) listing the range and median rate of GPP from the calibration dataset (Kimball et al.
-2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (Supplementary Table 1).
+We have added text (lines xxx-xxx) listing the range GPP from the calibration dataset (Kimball et al.
+2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary Table 1).
 
@@
 Need more justification for this flux ratio method.**
-We have added text and references to this section justifying the fact that plant productivity and ecosystem respiration are highly correlated on annual timescales, and ecosystem respiration tends to scale with carbon inputs to the rhizosphere from photosynthesis (lines xxx).
-We agree it would be nice to also incorporate gridded annual maps of projected soil moisture and soil temperature, but these data products do not exist for the Hawaiian Islands.
+The ultimate source of CO~2~ respired from ecosystems is from a single process: photosynthesis.
+The vast majority of CO~2~ respired from soils was fixed by photosynthesis within a year of being respired, often within hours.
+A disproportionate fraction of annual variability in ecosystem carbon balance is associated with biophyscial factors (such as temperature and moisture) that modulate photosynthesis rather than respiration.
+Consequently, biophysical conditions that increase photosynthesis increase respiration over that same time period, and vice versa.
+We have added text and references to this section justifying the fact that plant productivity and ecosystem respiration are highly correlated on annual timescales, and ecosystem respiration tends to scale with carbon inputs to the soil from photosynthesis (lines xxx).
+Although we agree that incorporating gridded annual maps of projected future soil moisture and soil temperature could improve the predictive power our model, these data products do not exist for the Hawaiian Islands.
 
@@
 Presenting projected future NPP as we do in Figure 3 is consistent with the focus of this paper, which is on statewide trends of ecosystem carbon balance under different climate and land use scenarios.
+We have, however, added a map displaying the spatial variability in NPP used to initiate the LUCAS model (supplementary Figure 5).
+These spatial trends largely hold true throughout the model run.
 
@@
 For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**
-Again, projections of NPP change by land use and climate change scenario are presented in Figure 3, which we believe best conveys statewide trends of ecosystem carbon fluxes under different climate and land use scenarios.
-We have added text to the supplementary that better describes the empirical NPP model (lines xxx).
+Projections of NPP change by land use and climate change scenario are presented in Figure 3, which we believe best conveys statewide trends of ecosystem carbon fluxes under different climate and land use scenarios.
+We have added text and figures to the supplementary that better describes the empirical NPP model (supplementary lines xxx and supplementary Figure 5).
 
@@
 Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript.
-We have added text that describes how to the supplementary climate section that provides more detail on climate projections (lines xxx) No changes were made in response to this comment.
+We have added text to the supplementary climate section that provides more detail on climate projections (supplementary lines xxx).
 
 **L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
-Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ to 7.95 Tg y^-1^), and the ranges of statewide NPP estimates for each of these years broadly overlapped (2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
+Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ to 7.95 Tg y^-1^), and the ranges of statewide annual NPP estimates  broadly overlapped for each of these years(2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
 This modeled slight decline in NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
-We have added text that describes this change in NPP over the contemporary (2010-2020) time period (lines xxx-xxx).
+We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the discussion section (lines xxx-xxx).
 
@@
 It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
-We agree, this phrasing was misleading. We have altered the text
+We agree, this phrasing was misleading. We have altered the text as suggested by this referee (lines xxx).
 
@@
 2010; see lines xxx of revised manuscript).
-The drivers of Rh are a combination of climate change and land use change.
-High rates of urbanization remove plants and compress and pave over soils, reducing carbon input and lowering microbial activity.
-Climate change on the the other hand, has a ʻpush-pullʻ impact on Rh.
-On the one hand
 
@@
 Are these decreases supported by observations?**
-Land use change rates are the same in all scenarios from 2010-2020, which explains why Rh behaves consistently across these scenarios.
+Land use change rates are the same in all four scenarios from 2010-2020, which explains why Rh behaves consistently across these scenarios.
 We have added text to make that clear in the Methods section (lines xxx).
-The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of agricultural land (supplementary Figure 5).
-Agriculture has a higher rate of Rh per unit of NPP than other land cover classes, and so Rh declined more than overall NPP as agricultural land was converted to urban area.
+The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 5).
 We have added text that describes these trends (lines xxx).
@@
 **All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
-We have checked the clarity of all figures by requesting multiple colleagues open the manuscript PDF file on their respective computers.
-None of these colleagues noted a lack of clarity in the figures.
-We would be happy to alter these figures if the ERL production team indicates they do not meet journal requirements.
+We have increased the resolution of all figures to 400 dpi.
 

### 2021-08-09 Manuscript/HI_Model_ms.Rmd
updated manuscript files, increased resolution to 400 dpi for all figure files

@@
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
-We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017.
+We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] that was converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017.
 We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017].
@@
 
-We initiated net primary production for each simulation cell in LUCAS as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019].
+We initiated net primary production in LUCAS for each simulation cell as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019].
 The spatial NPP multipliers were the NPP anomaly for each simulation cell relative to mean empirically-derived NPP values for each state type (supplemental Figure 4).

### 2021-08-09 Manuscript/HI_Model_supp.Rmd
updated manuscript files, increased resolution to 400 dpi for all figure files

@@
 We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each state type (i.e., the unique combination of moisture zone and land cover class; supplemental Figure 4).
-These NPP multipliers were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4).
+These NPP multipliers (range = 0.27 - 5.2) were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4).
 ```{r}

### 2021-08-09 Manuscript/fig_code/HI_CFE_CarbonBalance.R

@@ -150,5 +150,5 @@ cfeNBPfig <- ggplot(nbp, aes(CFE_Rate, NBPmean)) +
 cfepanel <- plot_grid(cfeTECfig, cfeNBPfig, 
 	labels = c("(a)", "(b)"), rel_widths = c(1.6, 1))
 
-ggsave("fig6_cfe.png", cfepanel, width = 9, height = 3.75)
+ggsave("fig6_cfe.png", cfepanel, width = 9, height = 3.75, dpi = 400)
 

### 2021-08-09 Manuscript/fig_code/HI_CarbonBalance.R

@@ -94,7 +94,7 @@ cflowfig <- cflows %>%
 		theme(panel.grid.minor = element_blank())}
 
 # save cflowfig 
-ggsave("../fig_images/fig3_Cflows.png", width = 5, height = 5)
+ggsave("./fig_images/fig3_Cflows.png", width = 5, height = 5, dpi = 400)
 
 # calculate NEP and NBP from C flows
 cbal <- cflowjoin %>%
@@ -173,7 +173,7 @@ draw_label(expression(Net~biome~productivity~(Tg~C~y^{-1})),
 	x=0, y=0.55, vjust= 1.1, angle=90, size = 12) 
 
 # save two panel NBP fig as .png file
-ggsave("fig4_NBPpanels.png", width = 7, height = 6.5)
+ggsave("./fig_images/fig4_NBPpanels.png", width = 7, height = 6.5, dpi = 400)
 
 # calculate cumulative ∆NBP when swiching land use & climate scenarios
 dcumNBP <- cbal %>%
@@ -231,5 +231,5 @@ climdC <- dcumNBP %>%
 cumdNBP_panels <- plot_grid(ludC, climdC, ncol = 1, align = "v")
 
 # save two panel cumulative dNBP figure
-ggsave("fig5_cumdNBP.png", width = 6, height = 4)
+ggsave("./fig_images/fig5_cumdNBP.png", width = 6, height = 4, dpi = 400)
 

### 2021-08-09 Manuscript/fig_code/HI_CarbonStocks.R

@@ -113,4 +113,4 @@ cstockpanel <- plot_grid(TECfig, stocksfig,
 	labels = c("(a)", "(b)"), rel_widths = c(1.2, 1))
 
 # save C stocks fig as .png file
-ggsave("fig2_Cstocks.png", cstockpanel, width = 9, height = 3.5)
+ggsave("./fig_images/fig2_Cstocks.png", cstockpanel, width = 9, height = 3.5, dpi = 400)

### 2021-08-09 Manuscript/fig_code/HI_ClimateFigs.R

@@ -17,24 +17,24 @@ library(RColorBrewer)
 library(cowplot) 
 
 # create list of annual rainfall .tif files
-rainlist <- list.files('rainfall_tiff_annual', full.names = TRUE)
+rainlist <- list.files('./Climate/rainfall_tiff_annual', full.names = TRUE)
 #create raster stack of annual rainfall (mm) .tif files
 rainstack <- stack(rainlist)
 # make raster layer of 30-year mean annual rainfall and re-sample
 # to change projection to UTM (this takes ~30 seconds)
 rain30ymean <- calc(rainstack, mean)
 # read in rasters of late century rainfall anomolies for RCPs 4.5 & 8.5
-rcp85anom <- raster("sd_rfanom_rcp85_late.tif")
-rcp45anom <- raster("sd_rfanom_rcp45_late.tif")
+rcp85anom <- raster("./Climate/GCM_rfanom/sd_rfanom_rcp85_late.tif")
+rcp45anom <- raster("./Climate/GCM_rfanom/sd_rfanom_rcp45_late.tif")
 # make rasters of late century rainfall differences from present
 raindiff85 <- rain30ymean*(rcp85anom/100)
 raindiff45 <- rain30ymean*(rcp45anom/100)
 
 # read in current mean annual temperature raster
-MAT <- raster("MAT_annual.tif") 
+MAT <- raster("./Climate/MAT_annual.tif") 
 # read in rasters of late century temperature differences for RCPs 4.5 & 8.5
-Tdelta_rcp45 <- raster("sd_Tdeltas_rcp45_late.tif")
-Tdelta_rcp85 <- raster("sd_Tdeltas_rcp85_late.tif")
+Tdelta_rcp45 <- raster("./Climate/GCM_Tdelta/sd_Tdeltas_rcp45_late.tif")
+Tdelta_rcp85 <- raster("./Climate/GCM_tdelta/sd_Tdeltas_rcp85_late.tif")
 
 #prepare rain30ymean and raindiff data to plot with geom_tile in ggplot
 rain_df <- rasterToPoints(rain30ymean) %>%
@@ -90,7 +90,8 @@ airtem <- ggplot(data=MAT_df) +
 # create two panel current climate figure
 climatefig <- plot_grid(airtem, rain, ncol =1, align = "hv")
 # save current climate fig as .png file 
[…]

### 2021-08-09 Manuscript/fig_code/HI_FireArea_plot.R

@@ -55,4 +55,5 @@ firepanel <- plot_grid(stateplot, islandplot,
 	labels = c("A", "B"), rel_widths = c(1.15, 1))
 
 # save as .png file 
-ggsave("FigS2_AnnualAreaBurned.png", width = 7.5, height = 3)
+ggsave("./Manuscript/fig_images/FigS2_AnnualAreaBurned.png", 
+	width = 7.5, height = 3, dpi = 400)

### 2021-08-09 Manuscript/fig_code/HI_LandCoverFig.R

@@ -17,10 +17,11 @@ library(ggplot2)
 # 3 = Plantation	8 = Agriculture
 # 4 = WoodyCrop		9 = Wetland
 # 5 = Barren	   10 = Shrubland
-StateClass <- raster('./InputData/spatial_data/StateClassNew_250m.tif')
+StateClass <- raster(
+    './InputData/spatial_data/initial_conditions/StateClassNew_250m.tif')
 
 # read annual MAT Tiff into R
-MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
+MAT <- raster("./Climate/MAT_annual.tif")
 
 # make reclassification matrix for simplified land cover raster
 lc_rcl <- c(1, NA, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6,
@@ -77,4 +78,4 @@ lcmap <- ggplot(data=lc_df) +
         size = 3.5)
 
 # save Land Cover map as .png file in HI_Model/output_processing/output
-ggsave("fig1_HI_LC.png", height = 5, width = 7, dpi = 300)
+ggsave("./Manuscript/fig_images/fig1_HI_LC.png", height = 5, width = 7, dpi = 400)

### 2021-08-09 Manuscript/fig_code/HI_StateClassArea.R

@@ -17,7 +17,8 @@ library(tidyr)
 
 # read State Class model output data into R and summarize 
 # statewide by land use scenario
-sc <- read.csv("StateClassArea.csv", stringsAsFactors = FALSE) %>%
+sc <- read.csv("./HI_LandAreaTransitions/StateClassArea.csv", 
+	stringsAsFactors = FALSE) %>%
 	mutate(StateClassID = str_remove(.$StateClassID, ":All")) %>%
 	group_by(Scenario, Timestep, Iteration, StateClassID) %>%
 	summarize(area = sum(Area)) %>%
@@ -56,5 +57,5 @@ filter(StateClassID %in% c("Agriculture",
 		legend.text = element_text(size=10))}
 
 # save figure to .png file 
-ggsave("figS5_StateClassArea.png", height = 4, width = 7, dpi = 300)
+ggsave("figS6_StateClassArea.png", height = 4, width = 7, dpi = 400)
 	

### 2021-08-09 Manuscript/fig_code/MZfig.R

@@ -16,10 +16,10 @@ library(ggplot2)
 # 1 = Dry 		
 # 2 = Mesic 		
 # 3 = Wet	
-MZ <- raster('./InputData/spatial_data/MZ_3_250m.tif')
+MZ <- raster('./InputData/spatial_data/initial_conditions/MZ_3_250m.tif')
 
 # read annual MAT Tiff into R
-MAT <- raster("./FlowMultipliers/data/MAT_annual.tif")
+MAT <- raster("./Climate/MAT_annual.tif")
 
 # re-sample MZ projection to Lat. Long. using MAT
 MZdeg <- projectRaster(MZ, MAT, method = 'ngb')
@@ -52,4 +52,5 @@ mzmap <- ggplot(data=MZ_df) +
     	linetype = "solid", size = 0.5, color = "white"))
 
 # save Moisture Zone map as .png file 
-ggsave("figS1_MZ.png", height = 5, width = 7, dpi = 300)
+ggsave("./Manuscript/fig_images/figS1_MZ.png", 
+	height = 5, width = 7, dpi = 400)

### 2021-08-09 Manuscript/fig_code/NPPsuppfig.R

@@ -1,0 +1,109 @@
+# create NPP figures based on 
+# IBIS mean values per state type 
+# (MZ & land cover class) and NPP
+# stationary spatial multiplier
+# based on contemporary climate
+# Paul C. Selmants 
+# 2021-08-04
+# R version 4.1.0
+
+# load required packages into R
+library(raster) #version 3.1-5
+library(rgdal) #version 1.4-8
+library(tidyr, warn.conflicts = FALSE) #version 1.0.2
+library(dplyr, warn.conflicts = FALSE) #version 0.8.5
+library(ggplot2)
+library(viridis)
+library(RColorBrewer)
+library(cowplot) 
+
+
+## Set working directory at highest level in HI_Model
+## GitHub repository, https://github.com/selmants/HI_Model 
+
+# read in raster of combined Moisture Zones and land cover classes, where:
+# 1 = Dry Forest	 	 7 = Mesic Forest			13 = Wet Forest
+# 2 = Dry Grass		 	 8 = Mesic Grass			14 = Wet Grass
+# 3 = Dry Shrub		 	 9 = Mesic Shrub			15 = Wet Shrub 
+# 4 = Dry Plantation	10 = Mesic Plantation		16 = Wet Plantation
+# 5 = Dry Ag			11 = Mesic Ag				17 = Wet Ag
+# 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
+MZSC <- raster("./Fire/data/processed/MZSC.tif")
+
+# read in NCEAS NPP spatial multiplier raster, match crs projection to MZSC,
+# and mask with MZSC to remove barren and urban areas
+npp_sm <- raster(
+	"./InputData/spatial_data/spatial_multipliers/NCEAS_NPP_sm.tif") %>%
+	projectRaster(.,MZSC) %>%
+	mask(., MZSC)
+
+# read annual MAT raster as template for Lat. Long. conversion
+MAT <- raster("./Climate/MAT_annual.tif")
+
+# create reclassification matrix corresponding to IBIS mean NPP values
+# for each state type (MZ and land cover class)
+npp_rcl <- data.frame(sc = seq(1,18),
+	npp = c(0.29,0.21,0.31,0.29,0.34,0.21,0.84,0.47,0.55,0.84,0.32,0.67,
+		1.14,0.81,0.85,1.14,0.3,0.56)) %>%
+		as.matrix(.) %>%
[…]

### 2021-08-12 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript and supplementary files

@@
 Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_.
-We have fully addressed each of the referees’ comments in detail below.
-Our point-by-point replies immediately follow each referee comment, and we include line numbers of the revised manuscript at the end of each response to identify where we made changes to the text.
+We thank the two referees for their review of this manuscript, and have fully addressed each of their comments in detail below.
+Our point-by-point responses immediately follow each referee comment, and we include line numbers of the revised manuscript in each response to identify where we made changes to the text.
 
@@
 
+As suggested by this reviewer, we have added clarification and justification about how future NPP and land cover change were modeled, as well as new figures illustrating the spatial distribution of current and projected future carbon fluxes.
+Our responses to individual comments below indicate where we made these changes.
+
 **L109, why need to assign each cell a state type?
@@
 This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
-We have added text and references to this section and the next that better describes the state of each cell (lines xxx).
+We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines xxx).
 
@@
 According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**
-State type in this model is the unique combination of moisture zone, island, and land cover class, as stated on line 110 of the original manuscript.
-This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic shrubland on Oʻahu.
-We have added text that better defines how state type influences transition probabilities ().
+We agree this was unclear.
+We have added text to the first paragraph of the Methods section that defines "transition" (line xxx) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines xxx).
 
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
-To distinguish the impact of a low land use scenario from a high land use scenario on changes in agricultural land area, we sampled rates of agricultural contraction and agricultural expansion in the low land use scenario from a uniform distribution bounded by zero and the minimum historical rate.
-Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus lowering the cropping intensity of land area over time.
+Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus reducing over time the total land area under active cultivation in the Hawaiian Islands.
 We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010).
@@
 **L126, missing description for urbanization rate for high land use scenario.**  
-As the referee themselves point this out in their previous comment, transition rates in the high land use scenario, including urbanization, were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural expansion, agricultural contraction, and urbanization for each island.
-This description is on lines 121-123 of the original manuscript and lines xxx - xxxx of the revised manuscript.
-No changes were made in response to this comment.
+Transition rates for urbanization in the high land use scenario were sampled from a uniform distribution bounded by the median and maximum historical rate, as stated on lines 121-123 of the original manuscript.
+We have re-worded this description to make it more clear that it includes urbanization (lines xxx - xxx).
 
@@
 Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land).
-We have added text to the supplementary that explains the increase in both forest area and grassland area (supplementary lines xxx - xxx).
+We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines xxx - xxx).
 Tree plantation does not refer to reforestation or afforestation.
-We did not not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
-Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (Fig.
-S5).
-We have added text to the Methods section that describes tree plantation harvest rates and transitions (lines xxx-xxx).
+We did not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
+Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 6).
+We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines xxx-xxx).
 The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al.
@@
 Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests.
-We have added text that describes where wildland fire occurs in the Hawaiian Islands (lines xxx - xxx).
+We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines xxx - xxx).
 
@@
 Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**
-We agree that this was a confusing use of terminology. We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript.
+We agree this was a confusing use of terminology.
+We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript when it refers only to land cover classification.
+We now use the term "state type" only when referring to the unique combination of island, moisture zone, and land cover class.
 
 **L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
-Again, we agree this was a confusing use of terminology.
-We have omitted "moisture zone" and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class.
+This was a confusing use of terminology.
+We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class (line xxx).
 
 **L162, how litter and dead wood pools were initialized?**  
-We have added text stating that litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).
+We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).
 
@@
 Or is there any bias between 30-year climate normal and RCP projections?**
-The contemporary 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al.
-2013, 2014), as depicted in Supplementary figure 3.
-We have added text that better describes this (supplementary lines xxx-xxx).
+The contemporary 30-year climate normals for rainfall and temperature come from the references cited (Giambelluca et al.
+2013, 2014), as depicted in supplementary Figure 3.
+We have added text that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).
 
@@
 Asner et al.
-(2016) only quantifies aboveground forest biomass.
-The additional 32 Tg C comes from the sum of belowground forest biomass and total live biomass in grasslands, shrublands, and agriculture.
-We have added text describing the equation we used to estimate belowground forest biomass, which was then used to calibrate the IBIS DGVM (lines xxx-xxx).
+(2016) only quantifies aboveground forest live biomass.
+The additional 32 Tg C comes from the sum of belowground forest live biomass and total live biomass in grasslands, shrublands, and agriculture.
+We have added text and citations describing how we used a power function applied to the Asner et al.
+(2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines xxx-xxx).
 
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
-We have added text (lines xxx-xxx) listing the range GPP from the calibration dataset (Kimball et al.
+We have added text (lines xxx-xxx) listing the range of GPP values from the calibration dataset (Kimball et al.
 2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary Table 1).
@@
 External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**
-We have added maps to the supplementary depicting the spatial distribution of stationary NPP multipliers as well as the spatial distribution of NPP that results from the product of IBIS mean NPP values and the NPP spatial multipliers based on the spatial variability of contemporary climate.
-In our model, as in the real world, a deline in NPP will result in a decline in biomass accumulation, but the biomass that has already accumulated persists and will not somehow be reduced just because growth rates have slowed.
-We have added text that better describes how and why we apply the NPP spatial multiplier (xxx).
+No, there is no set of spatial multipliers applied to living biomass.
+We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines xxx - xxx, supplementary figures 5 & 6).
+We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines xxx - xxx).
 
@@
 Need more justification for this flux ratio method.**
-The ultimate source of CO~2~ respired from ecosystems is from a single process: photosynthesis.
-The vast majority of CO~2~ respired from soils was fixed by photosynthesis within a year of being respired, often within hours.
-A disproportionate fraction of annual variability in ecosystem carbon balance is associated with biophyscial factors (such as temperature and moisture) that modulate photosynthesis rather than respiration.
-Consequently, biophysical conditions that increase photosynthesis increase respiration over that same time period, and vice versa.
-We have added text and references to this section justifying the fact that plant productivity and ecosystem respiration are highly correlated on annual timescales, and ecosystem respiration tends to scale with carbon inputs to the soil from photosynthesis (lines xxx).
-Although we agree that incorporating gridded annual maps of projected future soil moisture and soil temperature could improve the predictive power our model, these data products do not exist for the Hawaiian Islands.
+We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines xxx - xxx).
+Although we agree that gridded annual maps of projected future soil moisture and soil temperature could potentially improve predictions of ecosystem carbon flux in our model, these data products do not exist for the Hawaiian Islands.
 
@@
 Yes, the growth and decay multipliers vary annually.
-We have added text that describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx).
-The NPP trend over time under both RCPs is depicted in Figure 3.
-We believe this is a much more clear and succinct way to convey changes in NPP than in a series of maps of future projected NPP under each of the four scenarios.
-Presenting projected future NPP as we do in Figure 3 is consistent with the focus of this paper, which is on statewide trends of ecosystem carbon balance under different climate and land use scenarios.
-We have, however, added a map displaying the spatial variability in NPP used to initiate the LUCAS model (supplementary Figure 5).
-These spatial trends largely hold true throughout the model run.
+We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx-xxx, supplementary lines xxx-xxx, supplementary Figures 5 & 6).
+The spatial distribution for the contemporary NPP spatial multipliers is now depicted in supplementary Figure 5, and the spatial distribution of projected future NPP multipliers under both RCP 4.5 and RCP 8.5 are now depicted in supplementary Figure 6.
 
@@
 For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**
-Projections of NPP change by land use and climate change scenario are presented in Figure 3, which we believe best conveys statewide trends of ecosystem carbon fluxes under different climate and land use scenarios.
-We have added text and figures to the supplementary that better describes the empirical NPP model (supplementary lines xxx and supplementary Figure 5).
+We have added a supplemental figure displaying the spatial distribution of projected future NPP multipliers based on future climate projections (supplemental Figure 6).
+We have also added text to the supplementary that better describes how we employ the empirical NPP equations to calculate spatial NPP anomalies for both contemporary and projected future NPP (supplemental lines xxx).
 
@@
 Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript.
-We have added text to the supplementary climate section that provides more detail on climate projections (supplementary lines xxx).
+We have added text to the supplementary Climate section that provides more detail on climate projections (supplementary lines xxx).
 
 **L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
-Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ to 7.95 Tg y^-1^), and the ranges of statewide annual NPP estimates  broadly overlapped for each of these years(2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
-This modeled slight decline in NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
+Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
+This slight decline in modeled NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
 We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the discussion section (lines xxx-xxx).
@@
 What are drivers of Rh decrease?**
-Soil carbon stocks increase consistently only under the RCP 4.5 radiative forcing scenario.
-Under RCP 8.5, soil carbon stocks reach an asymptote by 2080 in the low land use scenario and begin to decline slightly in the high land use scenario.
-Also, Rh only declines consistently under the high land use scenario (Figure 3).
+Rh only declines consistently under the high land use scenario in both RCPs (Figure 3).
 In the low land use scenario Rh remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5.
-As stated previously, NPP is a direct driver of Rh because the vast majority of carbon respired heterotrphically from terrestrial ecosystems was fixed by photosynthesis that same year (Kuzyakov et al.
-2010; see lines xxx of revised manuscript).
+As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (see lines xxx of revised manuscript).
+Also, soil carbon stocks do not increase consistently under all scenarios.
+Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline slightly in the high land use scenario.
+The drivers of R~h~ change over time are described on lines 241-245 of the original manuscript (lines xxx of the revised manuscript).
 
@@
 Are these decreases supported by observations?**
-Land use change rates are the same in all four scenarios from 2010-2020, which explains why Rh behaves consistently across these scenarios.
+Land use change rates are the same in all four scenarios from 2010-2020 and climate driven changes in R~h~ are still small during this initial time period, which explains why R~h~ behaves consistently across these four scenarios.
 We have added text to make that clear in the Methods section (lines xxx).
-The decrease in Rh before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 5).
-We have added text that describes these trends (lines xxx).
+The decrease in R~h~ before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 7).
+We have added text that describes these trends (supplementary lines xxx).
 
@@
 The Hawaiʻi specific climate data is from Giambelluca et al.
-(2013), Giambelluca (2014), Ellison Timm et al.
-(2015), and Ellison Timm et al.
+(2013), Giambelluca (2014), Elison Timm et al.
+(2015), and Elison Timm et al.
 (2017) - all of which are cited several times throughout the manuscript and are depicted in supplementary figures 3 and 4.
@@
 **Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
-We have carefully checked the manuscript for typographical errors and corrected them on lines x, y, and z of the revised manuscript.
-We have also confirmed that all acronyms are spelled out upon initial use.
+We have carefully checked the manuscript for typographical errors and changed "our" to "Our" on line 88, changed "600pp" to "600 ppm" on line 201, and corrected the spelling of "supplemental" on lines 61, 157, 240, and 241 of the revised manuscript.
+We have also confirmed that all acronyms are spelled out upon initial use in the main body of the manuscript.
 
 **All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
-We have increased the resolution of all figures to 400 dpi.
+We have increased the resolution of all figures to 400 dpi in response to this comment.
 
@@
 The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient.
-We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx), as recommended by the author guidelines for this Journal.
+We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx - xxx), as recommended by the author guidelines for this Journal.
 No changes were made in response to this comment.
@@
 **Objectives should be rewritten in detail and comprehensively.** 
-We believe we have stated the objectives of this study clearly and succinctly, both in the abstract (lines xxx) and at the end of the Introduction section (lines xxx).
+We believe we have stated the objectives of this study clearly and succinctly, both in the abstract and at the end of the Introduction section.
 That said, we have altered the text to add more detail to our stated objectives (lines xxx-xxx).
@@
 **How did the authors perform the calibration and validation process?**
-We describe calibration of the model in the Methods section on lines xxx and lines xxx. Validation of ecosystem carbon balance estimates in the Hawaiian Islands is
-
+We have added text describing the calibration of the model in the Methods section on lines xxx-xxx.
+Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data, as we describe in detail the Discussion section, however we were able to compare our results to those from previous studies (lines xxx-xxx).
 **How did they select the sensitive parameters of the model that they used?**
-We are unsure exactly what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
-There are a number of parameters in this model, and we describe in detail how they were selected in both the Methods section and the Supplementary section.
+It is unclear to us what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
+We assume they are referring to our test of model sensitivity to different rates of a CO~2~ fertilization effect (CFE).
+We describe how parameters were selected for this CFE sensitivity test in the Methods section on lines xxx-xxx of the revised manuscript.
 Without further clarification, we do not know how to better respond to this question.
@@
 **Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?** 
-We disagree with this referee that our discussion section is unrelated to the results section.
-We followed the author guidelines for this journal by discussing the significance of the results and comparing them with previous work using relevant references.
-For example,
+We followed the author guidelines for this journal by discussing the significance of the results and comparing them with previous work using relevant references, and so we disagree with this referee that our discussion section is unrelated to the results section.

### 2021-08-12 Manuscript/HI_Model_ms.Rmd
updated manuscript and supplementary files

--- 2021-08-09 Manuscript/HI_Model_ms.Rmd
@@
 We used the Land Use and Carbon Scenario Simulator (LUCAS), a spatially explicit stochastic simulation model that integrates landscape change and carbon gain-loss, to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
-We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100.
 Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength.
@@
 Thousands of hectares were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990].
-Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
+Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass and shrub species [@suryanata_diversified_2009; @perroy_evolving_2016].
 Although these past trends surely inform the future impact of climate and land use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts.
-To date only one study has attempted to integrate land use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land use change scenario and moderate radiative forcing [SRES A1B, equivalent to RCP 6; @selmants_baseline_2017].
+To date only one study has attempted to integrate land use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land use change scenario and a single moderate radiative forcing scenario [Special Report on Emission Scenarios [SRES] A1B, equivalent to Representative Concentration pathway [RCP] 6; @selmants_baseline_2017].
 
 We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019].
-We developed four unique scenarios that explored different pathways, or future trajectories, of land use change and climate change.
+We developed four unique scenarios that explored different pathways, or future trajectories, of land use and climate change.
 These four scenarios represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5).
 In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019].
-Our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.
+Our goals were to estimate statewide changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of modeled ecosystem carbon balance estimates to varying levels of a CFE on NPP.
 
@@
 We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size.
-Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; supplemental figure 1), seven islands, and ten discrete land cover classes (figure 1).
+Each simulation cell was assigned to one of 210 possible state types based on the unique combination of seven islands, three moisture zones (dry, mesic, and wet; supplemental figure 1), and ten discrete land cover classes (figure 1).
 We also tracked the age and time since transition for each simulation cell as continuous state variables [@daniel_state-and-transition_2016].
@@
 ```  
-
 ### *2.2 States and transitions*
@@
 Transition targets set the area to be transitioned over time.
-These targets were based on recent historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018].
-For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum recent historical rates of agricultural contraction, agricultural expansion, and urbanization for each island.
-For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum recent historical rates for each island.
-Bounding agricultural change rates between zero and the recent historical minimum shifts the balance toward agricultural contraction over time in the low land use scenario, which reduces the land area under active cultivation and contrasts with the high variability in agricultural land change rates during the recent historical past [@suryanata_diversified_2009].
+These targets were based on either recent historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] or on population projections for the State of Hawaiʻi [@kim_population_2018].
+For the high land use scenario, rates of agricultural contraction, agricultural expansion, and urbanization for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum recent historical rates for each island.
+For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum recent historical rates for each island, which shifts the balance toward agricultural contraction and leads to a reduction in land area under active cultivation over time.
 Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018].
@@
 The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
-There was no transition pathway from any land cover class into tree plantation in either land use scenario, which is consistent with recent historical trends of stable or declining tree plantation land area.
+There was no transition pathway from any land cover class into tree plantation in either land use scenario, which is consistent with recent historical trends of stable or declining tree plantation land area since the year 2000.
+For the contemporary period (2010-2020), transition targets in both the high and low land use scenarios were set at low land use scenario rates and diverged after 2020.
 
@@
 We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014].
-We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017.
+We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and total forest live biomass.
+The total forest live biomass data was derived from @asner_rapid_2016 estimates of aboveground forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017.
 We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017].
@@
 
-We initiated net primary production for each simulation cell in LUCAS as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2018; @sleeter_effects_2019].
-The spatial NPP multipliers were the NPP anomaly for each simulation cell relative to mean empirically-derived NPP values for each state type (supplemental Figure 4).
-Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each state type to the microclimate-adjusted NPP value for each simulation cell.
+We initiated net primary production in LUCAS for each simulation cell as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a set of spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate [supplemental figure 5; @sleeter_effects_2018; @sleeter_effects_2019].
+The stationary spatial NPP multipliers were the NPP anomaly for each simulation cell relative to the mean empirically-derived NPP values for each state type (supplemental Figure 5).
+Soil carbon flux to the atmosphere (heterotrophic respiration, R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each state type to the microclimate-adjusted NPP value for each simulation cell.
+We derived soil carbon loss as a fraction of NPP because photosynthesis is the dominant factor driving variation in ecosystem carbon fluxes, including R~h~, and the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year [@kuzyakov_time_2010; @baldocchi_inter-annual_2018].
 All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].
+IBIS-derived carbon stocks for each state type were allowed to equilibrate with spatially variable NPP (supplemental figure 5) via a 100-year LUCAS spinup model run with no fire or land cover change and initiated with SSURGO soil carbon [@ssurgo_2016] and IBIS values for live biomass, standing dead wood, down dead wood, and litter carbon stocks.
+Spatially variable carbon stocks from this spinup model run were used to initiate the main LUCAS model run.
 
-Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on mid-century (2049-2069) and end-of-century (2070-2099) statistically downscaled CMIP5 temperature and rainfall projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017].
-The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario.
-The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q~10~ temperature coefficients based on climate model temperature projections for each radiative forcing scenario.
+Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on mid-century (2049-2069) and end-of-century (2070-2099) statistically downscaled CMIP5 temperature and rainfall projections for the Hawaiian Islands under the RCP 4.5 and RCP 8.5 radiative forcing scenarios [supplemental figure 4; @timm_statistical_2015; @timm_future_2017].
+Annual increments in temporal growth and decay multipliers were calculated by dividing mid-century or end-of-century estimates by the number of intervening years.
+Temporal growth multipliers used to represent the impact of future changes in rainfall and temperature on NPP were calculated using empirical NPP equations [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario [supplemental figure 6; @sleeter_effects_2019].
+Temporal decay multipliers used to represent the effect of future warming on turnover rates of dead organic matter were calculated using Q~10~ temperature coefficients based on climate model temperature projections for each radiative forcing scenario [supplemental figure 4; @timm_future_2017].
 The Q~10~ temperature coefficients represented the proportional change in detrital carbon turnover due to a 10° C change in mean annual temperature.
-We applied a Q~10~ of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q~10~ of 2.17 for litter decay flows [@bothwell_leaf_2014].
+We applied a Q~10~ of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3_2009; @sleeter_effects_2019] and a Q~10~ of 2.17 for litter decay flows [@bothwell_leaf_2014].
 Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.
@@
 We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.
-For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further stimulation of NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
+For all CFE levels, we assumed a saturation point at an atmospheric CO~2~ concentration of 600 ppm, with no further stimulation of NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100.
 This 600 ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.
@@
 All simulations were performed within the SyncroSim (version 2.2.4) software framework with ST-Sim (version 3.2.13) and SF (version 3.2.10) add-on modules [@daniel_state-and-transition_2016; @daniel_integrating_2018].
-Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
+Model input data and output summaries were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.
 Carbon stocks and fluxes for the seven main Hawaiian Islands were calculated for each scenario by summing within each Monte Carlo realization on an annual basis and then calculating annual means as well as the annual upper and lower limits of the 30 Monte Carlo realizations.
@@
 ```
-Net primary production (NPP) for the seven main Hawaiian Islands declined across all four scenarios, driven primarily by climate change and to a lesser extent by land use change (Fig.
+Net primary production (NPP) for the seven main Hawaiian Islands was ~8.1 Tg y^-1^ at the beginning of the simulation period in 2010 (supplementary figure 5) with land use change driving an approximate 2% decline during the contemporary period (2010-2020).
+NPP continued to decline throughout the rest of the simulation period (2020-2100) across all four scenarios, but this long-term decline was driven primarily by climate change and to a lesser extent by land use change (Fig.
 3).
-The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figure 4) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 5).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being ~ 3% higher by 2100 than under RCP 4.5 (figure 3).
-Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 5), similar to trends in NPP.
-Transition-triggered carbon fluxes to the atmosphere from land use, land use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~0.4 Tg y^-1^ in the high land use scenario and ~0.2 Tg y^-1^ in the low land use scenario (figure 3).
+The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figures 4 and 6) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 7).
+In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being 3% higher on average by 2100 than under RCP 4.5 (figure 3).
+Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 7), similar to trends in NPP.
+Transition-triggered carbon fluxes to the atmosphere from land use, land use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~ 0.4 Tg y^-1^ in the high land use scenario and ~ 0.2 Tg y^-1^ in the low land use scenario (figure 3).
 Uncertainty around transition-triggered carbon fluxes were higher in the high land use scenario, driven primarily by greater variability in wildland fire probabilities.
@@
 Net biome productivity (NBP) averaged approximately 0.6 Tg C y^-1^ at the start of the simulation period and declined over time in all four scenarios (figure 4).
-On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink throughout the simulation period under the RCP 4.5 radiative forcing scenario, but carbon sink strength was ~40% lower in the high land use scenario compared to the low land use scenario by the end of the simulation period (figure 4).
+On average, terrestrial ecosystems of the seven main Hawaiian Islands collectively acted as a net carbon sink under the RCP 4.5 radiative forcing scenario throughout the simulation period, but carbon sink strength was ~40% lower in the high land use scenario compared to the low land use scenario by the end of the simulation period (figure 4).
 In contrast, ecosystems of the Hawaiian Islands acted as a net carbon source to the atmosphere toward the latter half of the simulation period under RCP 8.5, with the transition from sink to source occuring 15 years earlier on average in the high land use scenario than in the low land use scenario (figure 4).
@@
 Switching from RCP 8.5 to RCP 4.5 under the low land use scenario yielded the greatest cumulative increase in NBP, resulting in a median gain of 26.5 Tg of carbon over the entire 90-year simulation period.
-
 ```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in net biome productivity (NBP) for the seven main Hawaiian Islands.
-Values above zero indicate terrestrial ecosystems are acting as a net carbon sink for atmospheric carbon and values below zero indicate ecosystems are acting as a net carbon source to the atmosphere.
+Values above zero indicate terrestrial ecosystems are acting as a net sink for atmospheric carbon and values below zero indicate ecosystems are acting as a net source of carbon to the atmosphere.
 Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations.
@@
 Under the high radiative forcing (RCP 8.5) and high land use scenario, the inclusion of a CFE ranging from 5-15% led to ~33-98 Tg of additional carbon storage in ecosystems by the end of the century, a ~10-30% increase (figure 6a).
-Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian Island ecosystems from a net carbon source to the atmosphere during the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
+Compared to the reference scenario (0% CFE), a 5% CFE was sufficient to transform Hawaiian terrestrial ecosystems from a net carbon source to the atmosphere during the latter half of the 21st century (figure 4b) to a net carbon sink for the entire simulation period (figure 6b), completely offsetting all other carbon losses induced by high radiative forcing and high land use.
 Net carbon sink strength was further enhanced at higher CFE rates, with NBP increasing by an average of 0.07 Tg C y^-1^ for each 1% increase in CFE (figure 6b).
@@
 A 15% CFE applied to the high radiative forcing and high land use scenario resulted in a mean annual NBP of 1.18 ± 0.29 Tg C y^-1^, more than double that of the low radiative forcing and low land use scenario with no CFE.
-
-\pagebreak
-
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of carbon dioxide fertilization in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing and high land use scenario.
-The carbon dioxide fertilization effect (CFE) is the percent change in net primary productivity (NPP) for every 100 ppm increase in atmospheric carbon dioxide.
-The CFE for all rates is capped at 600 ppm, which is achieved around the year 2060.
-Solid lines in (a) indicate the mean total ecosystem carbon storage across 30 Monte Carlo realizations for each CFE rate, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations.
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Sensitivity of projected changes in total ecosystem carbon storage (a) and mean annual net biome productivity (b) to different rates of a carbon dioxide fertilization effect (CFE) in the seven main Hawaiian Islands under the RCP 8.5 radiative forcing and high land use scenario.
+The CFE is the percent change in net primary productivity (NPP) for every 100 ppm increase in atmospheric carbon dioxide.
+Solid lines in (a) indicate mean total ecosystem carbon storage across 30 Monte Carlo realizations for each CFE rate, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations.
 Solid circles in (b) represent mean annual net biome productivity averaged across all years and Monte Carlo realizations for each CFE rate, with vertical lines indicating the standard deviation of the mean.
-The dashed horizontal line in (b) represents the boundary between ecosystems acting as a net carbon sink (positive NBP values) and a net carbon source (negative NBP values)."}
+The dashed horizontal line in (b) represents the boundary between ecosystems acting as a net carbon sink (positive values) and a net carbon source (negative values)."}
 knitr::include_graphics("./fig_images/fig6_cfe.png") 
@@
 For the time period 2011-2019, net biome productivity (NBP) averaged 0.64 TgC y^-1^ and ranged from 0.46 to 0.88 TgC y^-1^ across all scenarios.
-Based on this mean annual NBP estimate, Hawaiian terrestrial ecosystems offset approximately 13% of 2015 statewide CO~2~ emissions from energy production and transportation (5.04 TgC), the State of Hawaiʻi's largest source of greenhouse gas emissions [@hawaii_inventory_2019].
+Based on this mean annual NBP estimate, Hawaiian terrestrial ecosystems offset approximately 13% of 2015 statewide CO~2~ emissions from energy production and transportation (5.04 Tg C), the State of Hawaiʻi's largest source of greenhouse gas emissions [@hawaii_inventory_2019].
 Future projections indicate Hawaiian terrestrial ecosystems will continue to be a net sink for atmospheric carbon if global CO~2~ emissions peak around 2040 and then decline (RCP 4.5), and that carbon sink strength can be further enhanced by reducing the intensity and extent of future land use change.
@@
 Our model results also indicate that projections of ecosystem carbon balance are highly sensitive to the introduction of a CFE.
-Even a 5% increase in NPP for every 100ppm increase in atmospheric CO~2~ was sufficient to completely offset all other carbon losses induced by the high radiative forcing and high land use scenario, maintaining Hawaiian Island ecosystems as a net carbon sink for the entire simulation period instead of transitioning to a net carbon source by mid-century.
+Even a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ was sufficient to completely offset all other carbon losses induced by the high radiative forcing and high land use scenario, maintaining Hawaiian Island ecosystems as a net carbon sink for the entire simulation period instead of transitioning to a net carbon source by mid-century.
 Reconciling the high uncertainty surrounding the response of net photosynthesis to rising atmospheric CO~2~ is essential to more realistically constrain model projections of ecosystem carbon balance.
@@
 Soil carbon losses via R~h~, leaching, and overland flow in previous versions of the LUCAS model were calculated as the ratio of the IBIS-derived flux to the size of the originating carbon stock, in this case soil organic carbon to 1-m depth [@daniel_integrating_2018].
-Here we calculated soil R~h~ and aquatic carbon losses as the ratio of the mean IBIS-derived flux to the microclimate-adjusted NPP value of each simulation cell, which is a more realistic driver than stock size [@jackson_ecology_2017].
+Here we calculated soil R~h~ and aquatic carbon losses as the ratio of the mean IBIS-derived flux to the microclimate-adjusted NPP value of each simulation cell, which is a more realistic driver than stock size [@kuzyakov_time_2010; @jackson_ecology_2017; @baldocchi_inter-annual_2018].
 Compared to previous Hawaiʻi LUCAS model estimates [@selmants_baseline_2017], soil R~h~ and aquatic carbon losses from 2011-2020 were reduced by an average of 15% and 21%, respectively, which widened the gap between between carbon gain (NPP) and carbon losses and accounted for the overall increase in NBP estimates for this time period.
@@
 Studies like ours that incorporate stochasticity into spatially explicit simulation models could also provide a framework for the growing number of sub-national jurisdictions that plan to incorporate ecosystem carbon sequestration into their emissions reduction efforts.
-These long-term projections will be critical to understanding how future land use and climate-biosphere feedbacks could influence the achievement of climate mitigation goals.
+These long-term projections will be critical to assessing the impact of future land use change and climate-biosphere feedbacks on meeting climate mitigation goals.
 
@@
 
-Tabular model output data and metadata are available in machine readable format from the USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
-Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository at https://github.com/selmants/HI_Model.
+The data that support the findings of this study are openly available.
+LUCAS model tabular output data and metadata are available in machine readable format from the USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
+LUCAS model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository at https://github.com/selmants/HI_Model.
 
@@
 
-\pagebreak
-
 ## References

### 2021-08-12 Manuscript/HI_Model_supp.Rmd
updated manuscript and supplementary files

--- 2021-08-09 Manuscript/HI_Model_supp.Rmd
@@
 ---
+title: "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change"
 bibliography: HI_Model.bib
@@
 ---
-
 ```{r setup, include=FALSE}
@@
 ```
+# Supplementary Material
+
 \raggedright
-## Supplementary Material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change
-`r format(Sys.time(), '%B %d, %Y')`
 
-### Moisture Zones  
-Agriculture, Forest, Grassland, Shrubland, and Tree Plantation land cover classes were stratified into three Moisture Zones - Dry, Mesic, and Wet (supplemental Figure 1).
+**Authors:** Paul C. Selmants^1^, Benjamin M. Sleeter^2^,  Jinxun Liu^1^, Tamara S. Wilson^1^,  
+Clay Trauernicht^3^, Abby G. Frazier^4,5^, Gregory P. Asner^6^
+
+**Affiliations:**  
+^1^U.S. Geological Survey, Moffett Field, CA, USA     
+^2^U.S. Geological Survey, Seattle, WA, USA    
+^3^University of Hawaiʻi at Mānoa, Honolulu, HI, USA  
+^4^East-West Center, Honolulu, HI, USA     
+^5^Clark University, Worcester, MA, USA   
+^6^Arizona State University, Tempe, AZ, USA
+
+**Running title:** Hawaiʻi carbon balance
+
+**Keywords:** land use, climate change, carbon balance, Hawaiʻi, scenarios, disturbance, ecosystem model
+
+**Date:** `r format(Sys.time(), '%B %d, %Y')`
+
+\pagebreak
+
+## Supplementary Methods
+### *Moisture Zones*  
+Agriculture, Forest, Grassland, Shrubland, and Tree Plantation land cover classes were stratified into three Moisture Zones - Dry, Mesic, and Wet (supplemental figure 1).
 These three zones were based on a moisture availability index (MAI), calculated as mean annual precipitation (MAP) minus potential evapotranspiration [PET; @price_mz_2012].
-Areas where where MAI values were less than zero (i.e., where MAP < PET) were classified in the Dry Moisture Zone.
+Areas where where MAI values were less than zero (i.e., MAP < PET) were classified in the Dry Moisture Zone.
 Areas with MAI values between zero and 1,661 were classified in the Mesic Moisture Zone.
@@
 ```
-### Wildland Fire  
-We used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (supplemental Figure 2), wildland fire probabilities by state type, and wildland fire size distributions.
-This new database compiles prior mapping efforts and data collections with newly mapped fire perimeter data from Dr.
+### *Wildland Fire*  
+We used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (supplemental figure 2), wildland fire probabilities by state type, and wildland fire size distributions.
+This new database compiles prior mapping efforts and data collections with new  fire perimeter data mapped by Dr.
 Clay Trauernicht (Department of Natural Resources and Environmental Management, University of Hawaiʻi at Mānoa).
@@
 Fire perimeters for the years 2002-2011 were primarily from the U.S.
-Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
+Geological Survey Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov).
 The Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters from Hawaiʻi Island, mostly from the Kona and Kohala regions.
 The U.S.
-National Park Service provided ground-based, GPS-mapped fire records from Hawaiʻi Volcanoes National Park.
-The Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fire records on Oʻahu.
+National Park Service provided ground-based, GPS-mapped fire records from Hawaiʻi Volcanoes National Park, and the Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fire records on Oʻahu.
 All other fires were mapped directly by Dr.
@@
 ```
-### Climate  
-Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (supplemental Figure 3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
-Spatially explicit projections of change in annual temperature and annual rainfall for mid-century (2049-2069) and end of century (2070-2099; supplemental Figure 4) are  from statistically downscaled CMIP5 climate projections under RCP 4.5 and RCP 8.5 from @timm_statistical_2015 and @timm_future_2017.
+### *Climate*  
+Spatially explicit contemporary mean annual temperature and rainfall data for the main Hawaiian Islands at 250-m resolution (supplemental figure 3) are from @giambelluca_online_2013 and @giambelluca_evapotranspiration_2014.
+Spatially explicit mid-century (2049-2069) and end of century (2070-2099) projections of change in annual temperature and annual rainfall (supplemental figure 4) are from statistically downscaled CMIP5 climate projections under RCP 4.5 and RCP 8.5 [@timm_statistical_2015; @timm_future_2017].
 To avoid biases, the modeled present-day (1975–2005) CMIP5 climatology was used to standardize the resulting predictor time series, and the simulated future changes were measured relative to their present-day mean states.
@@
 ```
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by 2100 under RCP 4.5 and RCP 8.5 based on statistical downscaling of an ensemble of climate models (CMIP5)."}
+```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Projected change in mean annual temperature (top panels) and mean annual rainfall (bottom panels) by 2100 under RCP 4.5 and RCP 8.5 based on CMIP5 statistical downscaling.
+Data from Elison Timm \\textit{et al.} (2015) and Elison Timm (2017)."}
 knitr::include_graphics("./fig_images/figS4_FutureClimate.png")
 ```
-### Initializing NPP 
-Initial values for net primary production (NPP) in LUCAS were calculated using a lookup table derived from the Integrated Biosphere Simulator (IBIS; supplemental Table 1).
-These values were adjusted with spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate (supplemental Figure 4) as described in detail in @sleeter_effects_2018 and @sleeter_effects_2019.
-Briefly, we derived the set of NPP spatial multipliers by first calculating NPP independently of the IBIS values for each simulation cell using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] based on Hawaiʻi-specific climate data (supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014).
-We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each state type (i.e., the unique combination of moisture zone and land cover class; supplemental Figure 4).
-These NPP multipliers were then applied to the IBIS-derived values to create a spatial distribution of NPP used to initiate the LUCAS model (supplemental Figure 4).
+### *Scaling NPP*  
+Initial values for net primary production (NPP) in LUCAS were calculated using values from the Integrated Biosphere Simulator (IBIS; supplemental table 1) adjusted with spatially explicit stationary NPP multipliers to reflect local variation driven by microclimate (supplemental Figure 5, top panel).
+The resulting NPP estimates (supplemental figure 5, bottom panel) preserve the IBIS-derived means for each combination of land cover class and moisture zone [supplemental table 1; @sleeter_effects_2018; @sleeter_effects_2019].
+We derived the set of NPP stationary spatial multipliers by first calculating NPP independently of the IBIS values for each simulation cell using empirical relationships between annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] based on Hawaiʻi-specific contemporary climate data (supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014).
+We then calculated the NPP spatial multipliers as the NPP anomaly for each simulation cell relative to the mean of these empirically-derived NPP values for each unique combination of moisture zone and land cover class (supplemental Figure 5, top panel).
+LUCAS model NPP was initiated using the product of stationary spatial NPP multipliers (supplemental figure 5, top panel) and  IBIS-derived NPP (supplemental table 1).
+
+Changes in NPP due to projected changes in climate were estimated using a set of temporal NPP multipliers representing the proportion of contemporary (2010) NPP (supplemental Figure 5) occuring in each simulation at mid-century and end-of-century under RCPs 4.5 and 8.5.
+We calculated temporal NPP multipliers by first estimating NPP for each simulation cell based on contemporary climate (supplemental figure 3) and projected future climates (supplemental figure 4) using empirical relationships between NPP and climate [@schuur_productivity_2003; @del_grosso_global_2008].
+We then divided contemporary NPP estimates by future NPP estimates in each simulation cell to derive mid-century and end-of-century temporal multipliers (supplemental figure 6).
+Annual increments were calculated by dividing mid-century and end-of-century temporal multipliers by the number of intervening years.
 ```{r}
@@
 ```
-### Change in State Class Area
-Projections of total land area covered by each of five State Classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high; supplemental Figure 5).
-The overall decline in Agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland area over the same time period.
-Although Agricultural land area in the high land use scenario appears more stable over time, the trend actually reflects a balance between high rates of agricultural contraction in some areas and high rates of agricultural expansion in others.
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projections of total land area by year for each of five State Classes in the seven main Hawaiian Islands (Agriculture, Developed, Forest, Grassland, and Shrubland) under low and high land use change scenarios for the period 2010-2100.
+```{r fig.align = 'center', out.width = "80%", fig.cap = "Supplemental - Temporal NPP multipliers for the year 2100 under RCP 4.5 (top panel) and RCP 8.5 (bottom panel).
+Temporal NPP multipliers represent the relative amount of initial (2010) NPP that occurs in each pixel in the year 2100 under each of two representative concentration pathways (RCPs) for atmospheric carbon dioxide.
+Developed and Barren land cover types are excluded from both maps."}
+knitr::include_graphics("./fig_images/figS6_tempNPPmulti.png")
+```
+### *Land Cover Change*
+Projections of total land area covered by each of five land cover classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high; supplemental Figure 7).
+The decline in Agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland area over the same time period.
+Changes in agricultural land area over time in the high land use scenario reflect a balance between high rates of agricultural contraction and high rates of agricultural expansion.
+```{r fig.align = 'center', out.width = "90%", fig.cap = "Supplemental - Projections of total land area by year for Agriculture, Developed, Forest, Grassland and Shrubland land cover classes in the Hawaiian Islands under low and high land use change scenarios for the period 2010-2100.
 Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
-knitr::include_graphics("./fig_images/figS6_StateClassArea.png")
+knitr::include_graphics("./fig_images/figS7_StateClassArea.png")
 ```
-\pagebreak
-
 ## References

### 2021-08-12 FlowMultipliers/code/NPPmultipliers.R

@@ -243,6 +243,14 @@ growthmultipliers <- overlay(futureNPP, presentNPP,
 # re-name layers in raster stack
 names(growthmultipliers) <- rcpnames
 
+# Write end-of-century NPP multiplier rasters for RCP 4.5 
+writeRaster(growthmultipliers[[1]], "./FlowMultipliers/data/RCP45_NPPmulti_end.tif", 
+	overwrite = TRUE)
+
+# Write end-of-century NPP multiplier rasters for RCP 8.5
+writeRaster(growthmultipliers[[3]], "./FlowMultipliers/data/RCP85_NPPmulti_end.tif", 
+	overwrite = TRUE)
+
 # create function to mask growthmultipliers stack with individual rasters
 # from islandstack
 islandmask <- function(x) {

### 2021-08-12 Manuscript/fig_code/HI_NPPscalingFigs.R

@@ -1,8 +1,10 @@
 # create NPP figures based on 
 # IBIS mean values per state type 
-# (MZ & land cover class) and NPP
+# (MZ & land cover class), NPP
 # stationary spatial multiplier
-# based on contemporary climate
+# based on contemporary climate,
+# and NPP temporal multipliers based on
+# CMIP5 climate projections
 # Paul C. Selmants 
 # 2021-08-04
 # R version 4.1.0
@@ -30,11 +32,26 @@ library(cowplot)
 # 6 = Dry WoodyCrop		12 = Mesic WoodyCrop		18 = Wet WoodyCrop
 MZSC <- raster("./Fire/data/processed/MZSC.tif")
 
-# read in NCEAS NPP spatial multiplier raster, match crs projection to MZSC,
-# and mask with MZSC to remove barren and urban areas
+# read in NCEAS NPP stationary spatial multiplier raster, 
+# match crs projection to MZSC, and mask with MZSC 
+# to remove barren and urban areas
 npp_sm <- raster(
 	"./InputData/spatial_data/spatial_multipliers/NCEAS_NPP_sm.tif") %>%
 	projectRaster(.,MZSC) %>%
+	mask(., MZSC)
+
+# read in RCP 4.5 NPP temporal spatial multiplier raster, 
+# match crs projection to MZSC, and mask with MZSC 
+npp_rcp45 <- raster(
+	"./InputData/spatial_data/spatial_multipliers/RCP45_NPPmulti_end.tif") %>%
+	projectRaster(., MZSC) %>%
+	mask(., MZSC)
+
+# read in RCP 8.5 NPP temporal spatial multiplier raster, 
+# match crs projection to MZSC, and mask with MZSC 
+npp_rcp85 <- raster(
+	"./InputData/spatial_data/spatial_multipliers/RCP85_NPPmulti_end.tif") %>%
+	projectRaster(., MZSC) %>%
 	mask(., MZSC)
 
 # read annual MAT raster as template for Lat. Long. conversion
@@ -54,18 +71,21 @@ IBIS_NPP <- reclassify(MZSC, npp_rcl)
 # multiply IBIS_NPP raster by npp spatial multiplier raster
 npp_initial <- IBIS_NPP * npp_sm 
 
-# re-sample npp_sm projection to Lat. Long. using MAT raster, then
[…]

### 2021-08-12 Manuscript/fig_code/HI_StateClassArea.R

@@ -57,5 +57,5 @@ filter(StateClassID %in% c("Agriculture",
 		legend.text = element_text(size=10))}
 
 # save figure to .png file 
-ggsave("figS6_StateClassArea.png", height = 4, width = 7, dpi = 400)
+ggsave("figS7_StateClassArea.png", height = 4, width = 7, dpi = 400)
 	

### 2021-08-12 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript files and response letter

@@
 Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_.
-We thank the two referees for their review of this manuscript, and have fully addressed each of their comments in detail below.
+We thank the two referees for their reviews of this manuscript, and have fully addressed each of their comments in detail below.
 Our point-by-point responses immediately follow each referee comment, and we include line numbers of the revised manuscript in each response to identify where we made changes to the text.
@@
 This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
-We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines xxx).
+We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines 101-102).
 
@@
 We agree this was unclear.
-We have added text to the first paragraph of the Methods section that defines "transition" (line xxx) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines xxx).
+We have added text to the first paragraph of the Methods section that defines "transition" (line 103) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines 121-128).
 
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
-Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus reducing over time the total land area under active cultivation in the Hawaiian Islands.
+Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, reducing over time the total land area under active cultivation in the Hawaiian Islands.
 We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010).
-We have added text that justifies bounding agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).
+We have added text that explains why we bounded agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines 134 - 137).
 
@@
 Transition rates for urbanization in the high land use scenario were sampled from a uniform distribution bounded by the median and maximum historical rate, as stated on lines 121-123 of the original manuscript.
-We have re-worded this description to make it more clear that it includes urbanization (lines xxx - xxx).
+We have re-worded this description to make it more clear that it includes urbanization (lines 131 - 134).
 
@@
 We did not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
-Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 6).
-We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines xxx-xxx).
+Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 7).
+We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines 143-152).
 The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al.
@@
 **L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
-We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines xxx - xxx).
+We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines 143 - 152).
 
@@
 Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests.
-We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines xxx - xxx).
+We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines 163 - 168).
 
@@
 This was a confusing use of terminology.
-We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class (line xxx).
+We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition as the unique combination of island, moisture zone, and land cover class (lines 115-117).
 
@@
 2013, 2014), as depicted in supplementary Figure 3.
-We have added text that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).
+We have added text to the supplementary that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).
 
@@
 We have added text and citations describing how we used a power function applied to the Asner et al.
-(2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines xxx-xxx).
+(2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines 184-186).
 
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
-We have added text (lines xxx-xxx) listing the range of GPP values from the calibration dataset (Kimball et al.
-2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary Table 1).
+We have added text (lines 188-189) listing the range of GPP values from the Kimball et al.
+(2017) calibration dataset and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary table 1).
 
@@
 We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines xxx - xxx, supplementary figures 5 & 6).
-We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines xxx - xxx).
+We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines 204-209).
 
@@
 Need more justification for this flux ratio method.**
-We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines xxx - xxx).
+We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines 199-202).
 Although we agree that gridded annual maps of projected future soil moisture and soil temperature could potentially improve predictions of ecosystem carbon flux in our model, these data products do not exist for the Hawaiian Islands.
@@
 Yes, the growth and decay multipliers vary annually.
-We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx-xxx, supplementary lines xxx-xxx, supplementary Figures 5 & 6).
-The spatial distribution for the contemporary NPP spatial multipliers is now depicted in supplementary Figure 5, and the spatial distribution of projected future NPP multipliers under both RCP 4.5 and RCP 8.5 are now depicted in supplementary Figure 6.
+We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 210-215, supplementary lines xxx-xxx, supplementary Figures 5 & 6).
 
@@
 **L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
-Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
+Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.1 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2011 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
 This slight decline in modeled NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
-We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the discussion section (lines xxx-xxx).
+We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the Results section (lines 274-276).
 

### 2021-08-12 Manuscript/HI_Model_ms.Rmd
updated manuscript files and response letter

@@
 The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figures 4 and 6) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 7).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being 3% higher on average by 2100 than under RCP 4.5 (figure 3).
+In contrast, more intense warming under RCP 8.5 (supplemental figure 4) led to an increase in heterotrophic respiration (R~h~) in the latter half of the 21st century, such that R~h~ was 3% higher on average by 2100 in the RCP 8.5 radiative forcing scenario than in the RCP 4.5 radiative forcing scenario (figure 3).
 Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 7), similar to trends in NPP.

### 2021-08-12 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript and response letter files

@@
 Best regards,
+
+```{r echo=FALSE, out.width='35%'}
+knitr::include_graphics('pcs_signature.png')
+```
 
@@
 Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land).
-We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines xxx - xxx).
+We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines 81-83).
 Tree plantation does not refer to reforestation or afforestation.
@@
 We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines 143-152).
-The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al.
-(2017).
+The "source land cover class" for tree plantations was abandoned sugar cane land as described in the Introduction section (lines 73-77).
 
@@
 We agree this was a confusing use of terminology.
-We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript when it refers only to land cover classification.
-We now use the term "state type" only when referring to the unique combination of island, moisture zone, and land cover class.
+We have altered the text here to read "...
+unique combination of moisture zone (supplemental figure 1), island, and land cover class (figure 1) for the years 1999-2019." (line 157-158).
+We have also changed the text to read  "land cover class" elsewhere throughout the manuscript when it refers only to land cover classification.
 
@@
 This was a confusing use of terminology.
-We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition as the unique combination of island, moisture zone, and land cover class (lines 115-117).
+We have completely re-written this paragraph based on this and other comments from this reviewer (lines 179-190), such that it no longer uses the term "state type" at all.
 
 **L162, how litter and dead wood pools were initialized?**  
-We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).
+We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines 179-180) and allowed to equilibrate with spatially variable NPP using a spinup model (lines 204-209).
 
@@
 Or is there any bias between 30-year climate normal and RCP projections?**
-The contemporary 30-year climate normals for rainfall and temperature come from the references cited (Giambelluca et al.
-2013, 2014), as depicted in supplementary Figure 3.
-We have added text to the supplementary that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).
+The contemporary 30-year climate normals for rainfall and temperature come from the references cited on line 182 (Giambelluca et al.
+2013, 2014) as depicted in supplementary Figure 3.
+We have added text to the supplementary that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines 47-53).
 
@@
 No, there is no set of spatial multipliers applied to living biomass.
-We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines xxx - xxx, supplementary figures 5 & 6).
+We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines 55-68, supplementary figures 5 & 6).
 We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines 204-209).
@@
 Yes, the growth and decay multipliers vary annually.
-We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 210-215, supplementary lines xxx-xxx, supplementary Figures 5 & 6).
+We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 210-215, supplementary lines 55-68, supplemental figures 5 & 6).
 
@@
 We have added a supplemental figure displaying the spatial distribution of projected future NPP multipliers based on future climate projections (supplemental Figure 6).
-We have also added text to the supplementary that better describes how we employ the empirical NPP equations to calculate spatial NPP anomalies for both contemporary and projected future NPP (supplemental lines xxx).
+We have also added text to the supplementary that better describes how we employ the empirical NPP equations to calculate spatial NPP anomalies for both contemporary and projected future NPP (supplementary lines 55-68).
 
@@
 Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript.
-We have added text to the supplementary Climate section that provides more detail on climate projections (supplementary lines xxx).
+We have added text to the supplementary Climate section that provides more detail on climate projections (supplementary lines 45-53).
 
@@
 Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.1 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2011 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^).
-This slight decline in modeled NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
+This slight decline in modeled NPP over 10 years, which is well within the error range of moderate resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
 We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the Results section (lines 274-276).
@@
 It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
-We agree, this phrasing was misleading. We have altered the text as suggested by this referee (lines xxx).
+We agree, this phrasing was misleading. We have altered the text to clarify that R~h~ increased in the low land use scenario in the latter half of the 21st century under RCP 8.5 (lines 282-285).
 
@@
 What are drivers of Rh decrease?**
-Rh only declines consistently under the high land use scenario in both RCPs (Figure 3).
-In the low land use scenario Rh remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5.
-As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (see lines xxx of revised manuscript).
+R~h~ only declines consistently under the high land use scenario in both RCPs (Figure 3).
+In the low land use scenario, R~h~ remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5.
+As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (lines 199-202).
 Also, soil carbon stocks do not increase consistently under all scenarios.
-Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline slightly in the high land use scenario.
-The drivers of R~h~ change over time are described on lines 241-245 of the original manuscript (lines xxx of the revised manuscript).
+Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline slightly in the high land use scenario (figure 2).
+We have revised this section of text to better describe the drivers of R~h~ change over time (lines 282-282).
 
@@
 Are these decreases supported by observations?**
-Land use change rates are the same in all four scenarios from 2010-2020 and climate driven changes in R~h~ are still small during this initial time period, which explains why R~h~ behaves consistently across these four scenarios.
-We have added text to make that clear in the Methods section (lines xxx).
+Land use change rates are set to be the same in both land use scenarios during the contemporary period from 2010-2020 (supplementary figure 7) and climate driven changes in R~h~ are still small during this initial time period.
+This explains why R~h~ behaves consistently across all four scenarios from 2010-2020 and diverges afterward.
 The decrease in R~h~ before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 7).
-We have added text that describes these trends (supplementary lines xxx).
 
 **Line 341, what is the source of ‘Hawai’i-specific climate data’?**
-The Hawaiʻi specific climate data is from Giambelluca et al.
-(2013), Giambelluca (2014), Elison Timm et al.
-(2015), and Elison Timm et al.
-(2017) - all of which are cited several times throughout the manuscript and are depicted in supplementary figures 3 and 4.
-We have added citations to these data again here in response to this comment (lines xxx).
+The Hawaiʻi specific climate data are from Giambelluca et al.
+(2013) and Giambelluca (2014), which are cited several times throughout the manuscript and are depicted in supplementary figures 3.
+We have added these citations and a reference to the supplemental figure again here (lines 382-383).
 
@@
 **Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
-We have carefully checked the manuscript for typographical errors and changed "our" to "Our" on line 88, changed "600pp" to "600 ppm" on line 201, and corrected the spelling of "supplemental" on lines 61, 157, 240, and 241 of the revised manuscript.
+We have carefully checked the manuscript for typographical errors and changed "600ppm" to "600 ppm" on line 240 as well as correcting several misspellings of the word "supplemental" in the revised manuscript.
 We have also confirmed that all acronyms are spelled out upon initial use in the main body of the manuscript.
@@
 **All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
-We have increased the resolution of all figures to 400 dpi in response to this comment.
+We have re-made all the figures to increase their resolution to 400 dpi in response to this comment.
 
@@
 
-**In the abstract, please add more information about the methods of the experiment.**  We have added text to the abstract that provides more information about the methods of this project (lines xxx).
+**In the abstract, please add more information about the methods of the experiment.**  We have added text to the abstract that provides more information about the methods of this project (lines 22-26).
 
 **In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**  
-We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines xxx).
+We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines 33-35 and lines 449-465).
 
 **Hypothesis should be given. How this work is different from the available literature?**
-A hypothesis would be warranted if the purpose of this study was to test existing theory using a controlled experiment, which was not the case.
-The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient.
-We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx - xxx), as recommended by the author guidelines for this Journal.
+We agree that a hypothesis would be warranted if the purpose of this study was to test existing theory using a controlled experiment, but that was not the case.
+The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient (see lines 93-97).
+We describe in detail how this work compares to the available literature in section 4.2 "Comparison other studies" (lines 369-394), as recommended by the author guidelines for this Journal.
 No changes were made in response to this comment.
@@
 **Objectives should be rewritten in detail and comprehensively.** 
-We believe we have stated the objectives of this study clearly and succinctly, both in the abstract and at the end of the Introduction section.
-That said, we have altered the text to add more detail to our stated objectives (lines xxx-xxx).
+We have altered the text to add more detail to our stated objectives (lines 93-97). 
+**How did the authors perform the calibration and validation process?**
+We have added text describing the calibration of the model to the Methods section (lines 182-190).
+Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data (described in section 4.2 "Comparison to other studies"), as well as the fact that we used much of the available statewide data to calibrate our model (see lines 182-190).
+However, we were able to compare our results to those from previous studies and provide details on factors driving differences in specific results (lines 369-394).
 
-**How did the authors perform the calibration and validation process?**
-We have added text describing the calibration of the model in the Methods section on lines xxx-xxx.
-Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data, as we describe in detail the Discussion section, however we were able to compare our results to those from previous studies (lines xxx-xxx).
 **How did they select the sensitive parameters of the model that they used?**
-It is unclear to us what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
+We are unclear what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
 We assume they are referring to our test of model sensitivity to different rates of a CO~2~ fertilization effect (CFE).
-We describe how parameters were selected for this CFE sensitivity test in the Methods section on lines xxx-xxx of the revised manuscript.
+We provide information and citations about how we selected parameters for the CFE sensitivity test in the Methods section (lines 229-241).
 Without further clarification, we do not know how to better respond to this question.
@@
 **Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?** 
-We followed the author guidelines for this journal by discussing the significance of the results and comparing them with previous work using relevant references, and so we disagree with this referee that our discussion section is unrelated to the results section.
+We followed the author guidelines for this journal by discussing the significance of our results and comparing them with previous work using relevant references.
+We therefore disagree with this referee that our Discussion section is unrelated to the Results section.
+Throughout the discussion, we put our results in the context of emissions mitigation for the State of Hawaiʻi, discuss the relative impact of future global emissions reductions versus local land use decisions on statewide ecosystem carbon balance, compare our results to those of previous studies, highlight how our model could be improved, and describe the significance of our results.
+We hope the editor will agree that the text of our Discussion section provides context to our results and highlights their significance without simply reiterating what we wrote in the Results section.
+No changes were made in response to this comment.

### 2021-08-12 Manuscript/HI_Model_ms.Rmd
updated manuscript and response letter files

--- 2021-08-12 Manuscript/HI_Model_ms.Rmd
@@
 The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figures 4 and 6) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 7).
-In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being 3% higher on average by 2100 than under RCP 4.5 (figure 3).
+In contrast, intense warming under RCP 8.5 (supplemental figure 4) led to an increase in heterotrophic respiration (R~h~) in the latter half of the 21st century under the low land use scenario (figure 3), and R~h~ was 3% higher on average by 2100 in the RCP 8.5 radiative forcing scenario than in the RCP 4.5 radiative forcing scenario.
 Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 7), similar to trends in NPP.
@@
 Previous versions of a Hawaiʻi LUCAS model were run at 1-km spatial resolution and simulation cells within each unique combination of moisture zone and state type all had the same mean IBIS-derived NPP value applied to them at the beginning of the simulation period.
-In contrast, our NPP estimates at 250-m spatial resolution were adjusted on a cell-by-cell basis using Hawaiʻi-specific climate data as described in section 2.3.
+In contrast, our NPP estimates at 250-m spatial resolution were adjusted on a cell-by-cell basis using Hawaiʻi-specific climate data [supplementary figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014] as described in section 2.3.
 As a result, our statewide NPP estimates from 2011-2020 were 9.5% lower on average than previous LUCAS model estimates for Hawaiʻi during the same time period [@selmants_baseline_2017], likely because of the greater influence of more arid simulation cells.

### 2021-08-13 Manuscript/journal_correspondence/ERL_responseletter.Rmd
updated manuscript, supplementary, and response letter files

@@
 
-As suggested by this reviewer, we have added clarification and justification about how future NPP and land cover change were modeled, as well as new figures illustrating the spatial distribution of current and projected future carbon fluxes.
+As suggested by this referee, we have added clarification and justification about how future NPP and land cover change were modeled, as well as new figures illustrating the spatial distribution of current and projected future carbon fluxes.
 Our responses to individual comments below indicate where we made these changes.
@@
 **L109, why need to assign each cell a state type?
-What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition.
+What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island and time since transition.
 This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur.
-We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines 101-102).
+We have added text and references to the first paragraph in the Methods section that better defines the terms 'state' and 'transition' and how they relate to each other (lines 102-106).
 
@@
 We agree this was unclear.
-We have added text to the first paragraph of the Methods section that defines "transition" (line 103) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines 121-128).
+We have added text to the first paragraph of the Methods section that defines the terms 'transition' and 'transition pathway' (lines 104-106) and we have re-written this section to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines 122-129).
 
@@
 What is the justification that low land use scenario will have rate lower than historical records?**
-Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, reducing over time the total land area under active cultivation in the Hawaiian Islands.
+Sampling from this range in the low land use scenario shifts the balance of agricultural land area change toward agricultural contraction, reducing over time the total land area under active cultivation in the Hawaiian Islands.
 We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010).
-We have added text that explains why we bounded agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines 134 - 137).
+We have added text that explains why we bounded agricultural land area change rates between zero and the minimum recent historical rates in the low land use scenario (lines 135 - 138).
 
@@
 Transition rates for urbanization in the high land use scenario were sampled from a uniform distribution bounded by the median and maximum historical rate, as stated on lines 121-123 of the original manuscript.
-We have re-worded this description to make it more clear that it includes urbanization (lines 131 - 134).
+We have re-worded this description to make it more clear that it includes urbanization (lines 132 - 135).
 
@@
 Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land).
-We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines 81-83).
+We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is primarily due to agricultural contraction over the same time period (supplementary lines 82-83; supplemental figure 7).
 Tree plantation does not refer to reforestation or afforestation.
-We did not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
-Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 7).
-We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines 143-152).
-The "source land cover class" for tree plantations was abandoned sugar cane land as described in the Introduction section (lines 73-77).
+We did not include expansion of tree plantations as a transition pathway, and so there is no land cover class that can be converted to tree plantation in either land use scenario.
+Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture.
+We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines 144-153).
+The "source land cover class" for tree plantations was abandoned sugar cane land as described in the Introduction section (lines 74-78).
 
 **L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
-We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines 143 - 152).
+We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines 144 - 153).
 
@@
 Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests.
-We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines 163 - 168).
+We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines 164 - 169).
 
@@
 We have altered the text here to read "...
-unique combination of moisture zone (supplemental figure 1), island, and land cover class (figure 1) for the years 1999-2019." (line 157-158).
-We have also changed the text to read  "land cover class" elsewhere throughout the manuscript when it refers only to land cover classification.
+each unique combination of moisture zone (supplemental figure 1), island, and land cover class (figure 1) for the years 1999-2019." (line 158-159).
+We have also changed the text to read 'land cover class' rather than 'state type' elsewhere throughout the manuscript when referring only to the land cover classification of simulation cells.
 
 **L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
-This was a confusing use of terminology.
-We have completely re-written this paragraph based on this and other comments from this reviewer (lines 179-190), such that it no longer uses the term "state type" at all.
+Yes, this was a confusing use of terminology.
+We have completely re-written this paragraph based on this and other comments from this referee (lines 180-191), such that we no longer uses the term 'state type' at all.
 
 **L162, how litter and dead wood pools were initialized?**  
-We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines 179-180) and allowed to equilibrate with spatially variable NPP using a spinup model (lines 204-209).
+We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines 180-181) and allowed to equilibrate with spatially variable NPP using a spinup model (lines 205-210).
 
@@
 Or is there any bias between 30-year climate normal and RCP projections?**
-The contemporary 30-year climate normals for rainfall and temperature come from the references cited on line 182 (Giambelluca et al.
-2013, 2014) as depicted in supplementary Figure 3.
+The contemporary 30-year climate normals for rainfall and temperature come from the references Giambelluca et al.
+2013 and Giambelluca et al.
+2014, as cited on line 183 and depicted in supplementary Figure 3.
 We have added text to the supplementary that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines 47-53).
@@
 We have added text and citations describing how we used a power function applied to the Asner et al.
-(2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines 184-186).
+(2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines 185-187).
 
 **L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
-We have added text (lines 188-189) listing the range of GPP values from the Kimball et al.
+We have added text (lines 189-190) listing the range of GPP values from the Kimball et al.
 (2017) calibration dataset and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary table 1).
@@
 No, there is no set of spatial multipliers applied to living biomass.
-We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines 55-68, supplementary figures 5 & 6).
-We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines 204-209).
+We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines 55-68, supplemental figures 5 & 6).
+We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks allowed for these stocks to reach a dynamic equilibrium with NPP across the landscape of the Hawaiian Islands (lines 205-210).
 
@@
 Need more justification for this flux ratio method.**
-We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines 199-202).
+We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales, and that the vast majority of carbon lost from soils on an annual basis was fixed by photosynthesis within that year (lines 200-203).
 Although we agree that gridded annual maps of projected future soil moisture and soil temperature could potentially improve predictions of ecosystem carbon flux in our model, these data products do not exist for the Hawaiian Islands.
@@
 Yes, the growth and decay multipliers vary annually.
-We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 210-215, supplementary lines 55-68, supplemental figures 5 & 6).
+We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 211-216, supplementary lines 55-68, supplemental figures 5 & 6).
 
@@
 This slight decline in modeled NPP over 10 years, which is well within the error range of moderate resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change.
-We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the Results section (lines 274-276).
+We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the Results section (lines 275-277).
 
@@
 It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
-We agree, this phrasing was misleading. We have altered the text to clarify that R~h~ increased in the low land use scenario in the latter half of the 21st century under RCP 8.5 (lines 282-285).
+We agree, this phrasing was misleading. We have altered the text to clarify that R~h~ increased in the low land use scenario in the latter half of the 21st century under RCP 8.5 (lines 283-286).
 
@@
 What are drivers of Rh decrease?**
-R~h~ only declines consistently under the high land use scenario in both RCPs (Figure 3).
-In the low land use scenario, R~h~ remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5.
-As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (lines 199-202).
-Also, soil carbon stocks do not increase consistently under all scenarios.
-Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline slightly in the high land use scenario (figure 2).
-We have revised this section of text to better describe the drivers of R~h~ change over time (lines 282-282).
+R~h~ only declines consistently under the high land use scenario in both RCPs (figure 3).
+In the low land use scenario, R~h~ remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5 (figure 3).
+As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (lines 200-203).
+Also, soil carbon stocks do not increase consistently under all scenarios (figure 2).
+Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline in the high land use scenario (figure 2).
+We have revised this section of text to better describe the drivers of R~h~ change over time (lines 283-289).
 
@@
 Are these decreases supported by observations?**
-Land use change rates are set to be the same in both land use scenarios during the contemporary period from 2010-2020 (supplementary figure 7) and climate driven changes in R~h~ are still small during this initial time period.
+Land use change rates are set to be the same in both land use scenarios during the contemporary period from 2010-2020 (lines 153-154; supplementary figure 7) and climate driven changes in R~h~ are still small during this initial time period.
 This explains why R~h~ behaves consistently across all four scenarios from 2010-2020 and diverges afterward.
@@
 (2013) and Giambelluca (2014), which are cited several times throughout the manuscript and are depicted in supplementary figures 3.
-We have added these citations and a reference to the supplemental figure again here (lines 382-383).
+We have added these citations again here, as well as a reference to the supplemental figure (lines 383-384).
 
@@
 **Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
-We have carefully checked the manuscript for typographical errors and changed "600ppm" to "600 ppm" on line 240 as well as correcting several misspellings of the word "supplemental" in the revised manuscript.
+We have carefully checked the manuscript for typographical errors and have changed "600ppm" to "600 ppm" on line 241 as well as corrected several misspellings of the word "supplemental" in the revised manuscript.
 We have also confirmed that all acronyms are spelled out upon initial use in the main body of the manuscript.
@@
 **In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**  
-We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines 33-35 and lines 449-465).
+We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines 34-36).
 
 **Hypothesis should be given. How this work is different from the available literature?**
-We agree that a hypothesis would be warranted if the purpose of this study was to test existing theory using a controlled experiment, but that was not the case.
-The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient (see lines 93-97).
-We describe in detail how this work compares to the available literature in section 4.2 "Comparison other studies" (lines 369-394), as recommended by the author guidelines for this Journal.
+A hypothesis statement would be warranted if the purpose of this study was to test existing theory using a controlled experiment, but that was not the case.
+The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient (see lines 94-98).
+We describe in detail how this work compares to the available literature in section 4.2 "Comparison other studies" (lines 370-395), as recommended by the author guidelines for this Journal.
 No changes were made in response to this comment.
@@
 **Objectives should be rewritten in detail and comprehensively.** 
-We have altered the text to add more detail to our stated objectives (lines 93-97). 
+We have altered the text to add more detail to our stated objectives (lines 94-98).
+
 **How did the authors perform the calibration and validation process?**
-We have added text describing the calibration of the model to the Methods section (lines 182-190).
-Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data (described in section 4.2 "Comparison to other studies"), as well as the fact that we used much of the available statewide data to calibrate our model (see lines 182-190).
-However, we were able to compare our results to those from previous studies and provide details on factors driving differences in specific results (lines 369-394).
+We have added text describing the calibration of the model to the Methods section (lines 182-191).
+Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data (described in section 4.2 "Comparison to other studies", lines 370-372), as well as the fact that we used much of the available statewide data to calibrate our model (see lines 182-191).
+However, we were able to compare our results to those from previous studies and provide details on factors driving differences in specific results (lines 372-395).
 
 **How did they select the sensitive parameters of the model that they used?**
-We are unclear what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
+We are unsure what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript.
 We assume they are referring to our test of model sensitivity to different rates of a CO~2~ fertilization effect (CFE).
-We provide information and citations about how we selected parameters for the CFE sensitivity test in the Methods section (lines 229-241).
-Without further clarification, we do not know how to better respond to this question.
+We provide information and citations about how we selected parameters for the CFE sensitivity test in the Methods section (lines 230-242).
 No changes were made in response to this comment.

### 2021-08-13 Manuscript/HI_Model_ms.Rmd
updated manuscript, supplementary, and response letter files

@@
 In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss.
-A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance.
-Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.
+A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance, highlighting the need for greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~.
+Long-term model projections such as ours that incorporate the interactive effects of land use and climate change on regional ecosystem carbon balance will be critical to evaluating the potential of ecosystem-based climate mitigation strategies.
 

### 2021-08-13 Manuscript/HI_Model_supp.Rmd
updated manuscript, supplementary, and response letter files

@@
 ### *Land Cover Change*
-Projections of total land area covered by each of five land cover classes (Agriculture, Developed, Forest, Grassland and Shrubland) were summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high; supplemental Figure 7).
-The decline in Agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland area over the same time period.
+Projections of total land area covered by each of five land cover classes (Agriculture, Developed, Forest, Grassland and Shrubland) summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high) are displayed in supplemental Figure 7.
+The decline in agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland land area over the same time period.
 Changes in agricultural land area over time in the high land use scenario reflect a balance between high rates of agricultural contraction and high rates of agricultural expansion.

### 2021-08-13 Manuscript/fig_code/HI_MZfig.R
updated R figure scripts

@@ -51,6 +51,6 @@ mzmap <- ggplot(data=MZ_df) +
     legend.background = element_rect(fill="white", 
     	linetype = "solid", size = 0.5, color = "white"))
 
-# save Moisture Zone map as .png file 
+# save Moisture Zone map as .png file
 ggsave("./Manuscript/fig_images/figS1_MZ.png", 
 	height = 5, width = 7, dpi = 400)

### 2021-08-13 Manuscript/HI_Model_ms.Rmd
updated ReadMe and manuscript files

@@
 The data that support the findings of this study are openly available.
-LUCAS model tabular output data and metadata are available in machine readable format from the USGS ScienceBase data repository at https://doi.org/10.5066/P9AWLFKZ.
-LUCAS model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository at https://github.com/selmants/HI_Model.
+LUCAS model tabular output data and metadata are available in machine readable format from  USGS ScienceBase at https://doi.org/10.5066/P9AWLFKZ.
+LUCAS model input data and R scripts used to format input data, summarize output data, and compile this manuscript are available at https://doi.org/10.5281/zenodo.5198072.
 

### 2021-08-13 Manuscript/HI_Model_ms.Rmd
updated manuscript and response letter files

@@
 The data that support the findings of this study are openly available.
-LUCAS model tabular output data and metadata are available in machine readable format from  USGS ScienceBase at https://doi.org/10.5066/P9AWLFKZ.
-LUCAS model input data and R scripts used to format input data, summarize output data, and compile this manuscript are available at https://doi.org/10.5281/zenodo.5198072.
+LUCAS model tabular output data and metadata are available in machine readable format from  the USGS ScienceBase repository at https://doi.org/10.5066/P9AWLFKZ.
+LUCAS model input data and R scripts used to format input data, summarize output data and compile this manuscript are available from a GitHub repository (https://github.com/selmants/HI_Model) archived at https://doi.org/10.5281/zenodo.5198072.
 

### 2021-08-13 Manuscript/HI_Model_ms.Rmd
updated manuscript files

@@
   pdf_document: 
+    keep_tex: yes
     latex_engine: xelatex

### 2021-08-13 Manuscript/HI_Model_ms20210405.Rmd
updated manuscript files

@@
 The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario.
-We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014].
+We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014].
 Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.

### 2021-09-01 Manuscript/HI_Model_supp.Rmd
updated supplementary files

@@
 Areas with MAI values greater than 1,661 were classified in the Wet Moisture Zone.
-```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Moisture zones of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017)."}
+```{r fig.align = 'center', out.width = "90%", fig.cap = "Supplemental - Moisture zones of the seven main Hawaiian Islands, reproduced from Jacobi \\textit{et al.} (2017).
+Credit: U.S.
+Geological Survey."}
 knitr::include_graphics("./fig_images/figS1_MZ.png")
@@
 Data from USGS MTBS were prioritized for the years 2002-2011 in the case of duplicate records.
-```{r fig.align = 'center', out.width = "95%", fig.cap = "Supplemental - Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Annual area burned by wildland fire in the State of Hawaiʻi from 1999-2019, summed by year across the seven main Hawaiian Islands (A) and within each of the four largest islands (B).
 The dashed horizontal line in (A) represents the median area burned from 1999-2019."}
@@
 ### *Land Cover Change*
-Projections of total land area covered by each of five land cover classes (Agriculture, Developed, Forest, Grassland and Shrubland) summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high) are displayed in supplemental Figure 7.
+Projections of land area covered by each of five land cover classes (Agriculture, Developed, Forest, Grassland and Shrubland) summed by year and Monte Carlo iteration across the seven main Hawaiian Islands for each of two land use scenarios (low and high) are displayed in supplemental Figure 7.
 The decline in agricultural land between 2020 and 2040 in the low land use scenario led directly to the increase in Forest, Grassland, and Shrubland land area over the same time period.
 Changes in agricultural land area over time in the high land use scenario reflect a balance between high rates of agricultural contraction and high rates of agricultural expansion.
-```{r fig.align = 'center', out.width = "90%", fig.cap = "Supplemental - Projections of total land area by year for Agriculture, Developed, Forest, Grassland and Shrubland land cover classes in the Hawaiian Islands under low and high land use change scenarios for the period 2010-2100.
+
+```{r fig.align = 'center', out.width = "99%", fig.cap = "Supplemental - Projections of total land area by year for Agriculture, Developed, Forest, Grassland and Shrubland land cover classes in the Hawaiian Islands under low and high land use change scenarios for the period 2010-2100.
 Solid lines represent the mean of 30 Monte Carlo realizations and shaded areas represent minimum and maximum Monte Carlo values."}
